### 2014-04-07 thesispropbody_d1.tex
Uploading thesis materials
Checklists, policy, and stuff for the proposal. Thesis itself will be
at least a different file. Likely will be a different directory
entirely.

\title{AGB Stars in the IR: Physical Characteristics, Galactic Spatial Distribution, and Chemical Impact of the Milky Way AGB Population}
\author{Nicholas Hunt-Walker}
\maketitle

\section{Physical/Observational Characteristics}
What do I want to talk about here? Absolutely must learn and talk about {\tt DUSTY} here, as well as in section 3.  Estimating mass, temperature, Mbol, mass loss rate, main pulsation period, amount/extent/temperature of circumstellar shell, age, chemical class if possible.  Want some color-color and color-mag diagrams here too.  Probably also talk about candidate selection.

Look into the OGLE and MACHO samples to see if there's some relation between pulsation period and WISE color.  Probably is.

Figure out some way to tell chemical type of AGB star for rest of un-typed candidate sample, which is MOST of the sample.

\section{Spatial Distribution}
I want to talk about my maps here.  Want XYZ 2-axis maps, but also want a 3-D map, as well as a 2D gal l, gal b map.  Also talk about using {\tt TRILEGAL} and using GLIMPSE. Somehow compare my actual distribution of AGB stars to models of galactic structure.  Note of course that my extinction calculations are based entirely on LSST code, so there would be bias.

Consider creating comparable XYZ maps for untyped AGB candidates using both the O-rich and C-rich color-mag relations.

\section{Chemical Impact}
Something about ISM enrichment and chemical output over time. Link the mass loss rate to the chemical content of the circumstellar shell.  Perhaps calculate the chemical enrichment as a function of distance from the star?

\section{Time Frame}
2 years from the end of the general exam.
Major points to hit:
\begin{itemize}
\item Selection of AGB candidates from WISE data
\item Split into high likelihood O-rich and C-rich populations.
\item Figure out someway to relax selection criteria to potentially get more stars.  Consider looking at candidates from both WISE All-Sky and ALLWISE (confusing).
\end{itemize}

### 2014-10-31 thesispropbody_d1.tex
updating
Uploading more!

@@
-\title{AGB Stars in the IR: Physical Characteristics, Galactic Spatial Distribution, and Chemical Impact of the Milky Way AGB Population}
-\author{Nicholas Hunt-Walker}
-\maketitle
+\section{Introduction}
+The formation of galaxies like the Milky Way was long thought to be a steady process that created a smooth distribution of stars.
+This standard view was exemplified by the \cite{1980ApJS...44...73B} and \cite{1989ARA&A..27..555G} models, and described in detail by, e.g., \cite{1993ARA&A..31..575M}.
+In these, the Milky Way is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
+Recent discoveries of complex substructure in the distribution of the Milky Way's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V,2008ApJ...673..864J} have deeply shaken this standard view with indications of signs of rich substructure within the Milky Way galaxy.
+This was due to the ability to map the Milky Way using the distances and positions of its own stellar residents.
+I propose to continue the mapping of the Milky Way in further detail, using the ubiquitous stars of the Asymptotic Giant Branch (hereon AGB).
 
-\section{Physical/Observational Characteristics}
-What do I want to talk about here?
-Absolutely must learn and talk about {\tt DUSTY} here, as well as in section 3.
-Estimating mass, temperature, Mbol, mass loss rate, main pulsation period, amount/extent/temperature of circumstellar shell, age, chemical class if possible.
-Want some color-color and color-mag diagrams here too.
-Probably also talk about candidate selection.
+\subsection*{\centerline{\ub{First Steps Toward Milky Way Tomography}}}
+Until recently, our knowledge of the basic structural components of the Milky Way was limited to indirect inferences based on stellar population models motivated by other galaxies, and what little information was available from the High Precision Parallax Collecting Satellite \cite[HIPPARCOS;][]{1984SSRv...39....1K} catalog and smaller pencil-beam surveys.
+The advent of the Sloan Digital Sky Survey \citep[SDSS;][]{2000AJ....120.1579Y} alleviated this limitation, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
+SDSS photometry enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
+In turn, this led to the large-scale ``tomography" of the Milky Way via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
+The resulting maps provided a quantitative basis for analysis of the main structural components of the Galaxy, enabled efficient searches for complex substructure, and a robust comparison with various model predictions \citep{2012ApJ...757..166B,2012ARA&A..50..251I}.
+An example of these results is reproduced in Figure 1, showing a panoramic view of the variation in the median $[Fe/H]$ over an unprecedentedly-large volume of the Galaxy.
+%Get this figure and import it here
 
-Look into the OGLE and MACHO samples to see if there's some relation between pulsation period and WISE color.  Probably is.
+%-----------
+\vspace{-0.41in}
+\begin{minipage}{0.9\textwidth}
+\vbox{
+      \begin{minipage}[l]{3in}
+      \hrule \smallskip
+        {\footnotesize\baselineskip0.4cm{{\cap Figure 1:} {
+        The variation 
+        of the median photometric stellar metallicity in cylindrical Galactic coordinates 
+        $R$ and $|Z|$  for $\sim$2.5 million stars from SDSS with $14.5 < r <20$,
+        $0.2 < g-r < 0.4$ (blue turn-off stars), and photometric distance in the 0.8-9 kpc 
+        range %\citep[details in][]{2008ApJ...684..287I}. 
+        Each bin 
+        (50 pc by 50 pc) contained in this map is colored according
+        to the legend in the top left. The gradient of the median metallicity is essentially 
+        parallel to the $|Z|$ axis, except in the region of the Monoceros stream, as marked. 
+        The gray scale background is the best-fit model for the stellar number density distribution 
+        from \cite{2008ApJ...673..864J}. The inset in the top right illustrates the extent of the data 
+        volume relative to the rest of Galaxy, and the background image is the Andromeda 
+        galaxy. The three squares outline the regions used to measure the 3-dimensional
+        velocity distribution for halo stars. The arrows illustrate the variation of the velocity
+        ellipsoid orientation, which always points toward the Galactic center. Adapted from 
+        \cite{2012ARA&A..50..251I}.
+\smallskip \hrule}}}
+     %
+      \end{minipage}\ \hfill \
+      \begin{minipage}[r]{3in}
+        \epsfig{file=figs/IBJ_fig5.pdf, height=4.5in}
+      \end{minipage} \ \hfill \
+}
+\end{minipage}
+\vspace{-28pt}
+%------------
 
-Figure out some way to tell chemical type of AGB star for rest of un-typed candidate sample, which is MOST of the sample.
+%These recent developments, based on a host of accurate large-area surveys such as the Two-Micron All-Sky Survey \citep[2MASS;][]{2006AJ....131.1163S} and the Sloan Digital Sky Survey \citep[SDSS;][]{2000AJ....120.1579Y}, have made it abundantly clear that the Milky Way is a complex and dynamical structure that is still being shaped by the merging of smaller galaxies.
+Numerical simulations suggest that this merger process plays a crucial role in setting the structure and motions of stars within galaxies.
+Since the individual stars that make up the stellar populations in the Milky Way can be studied in great detail, their characterization proposed here will provide clues about galaxy evolution that cannot be extracted from observations of distant galaxies.
+%
+%To that end, I propose a two-year study to:
+%\begin{itemize}
+%\item mine the extensive data set provided by the Wide-Field Infrared Survey Explorer \citep[WISE;][]{2010AJ....140.1868W}, relating observables to the \emph{Spitzer} Galactic Legacy Infrared Mid-Plane Survey Extraordinaire \citep[GLIMPSE;][]{2003PASP..115..953B,2009PASP..121..213C}, SDSS, 2MASS, and other surveys.
+%\item reliably identify and track a unique yet ubiquitous population of high-luminosity, IR-bright stars: Asymptotic Giant Branch (hereon AGB) stars.
+%\item map the structure of the Milky Way in unprecedented detail, with AGB stars as our indicators of distance and age.
+%\item use the distribution of AGB stars as observational constraints on models of the formation and evolution of the Milky Way disk.
+%\end{itemize}
 
-\section{Spatial Distribution}
-I want to talk about my maps here.
-Want XYZ 2-axis maps, but also want a 3-D map, as well as a 2D gal l, gal b map.
-Also talk about using {\tt TRILEGAL} and using GLIMPSE.
-Somehow compare my actual distribution of AGB stars to models of galactic structure.
-Note of course that my extinction calculations are based entirely on LSST code, so there would be bias.
+\section{Work Completed/In Progress}
+\subsection{Development of Broad Color-Color Criteria}
+%Using positional matches between SIMBAD, MACHO, and OGLE to ALLWISE-2MASS, create simple color-color, color-magnitude, and photometric quality criteria to select AGB stars in the ALLWISE-2MASS data set.
+We begin by selecting a known sample of AGB stars and matching to the ALLWISE-2MASS data release.
+The MACHO project \citep{2008AJ....136.1242F} and the Optical Gravitational Lens Experiment (OGLE) III Online Catalog of Variable Stars\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}} both contain catalogs of AGB stars, confirmed from their variability.
+Additionally, the SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} contains objects classified as AGB stars from small surveys and individual studies.
+We positionally match these catalogs to objects from the ALLWISE-2MASS data using a circle of uncertainty with radius 1", and necessitate that every match have a 2MASS association.
 
-Consider creating comparable XYZ maps for untyped AGB candidates using both the O-rich and C-rich color-mag relations.
+The region of color-color space occupied by these stars is also occupied by other objects that happen to have dusty environments, as well as naked stars on the main sequence and extragalactic sources.
+As such, we need a reliable sample of contaminants that we can exclude in color-color space to increase the likelihood that our population of candidates represents the objects we wish to study.
+We gather AGN, QSOs, and star forming galaxies from the Sloan Digital Sky Survey \citep{2000AJ....120.1579Y} data release (DR) 7, specifically the NYU Value-Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}, and the LRGs from the SDSS Luminous Red Galaxy Survey \citep{2010ApJ...710.1444K}.
+Data for stars in the SDSS stellar locus \citep{2014MNRAS.440.3430D} were drawn from the DR 9 SEGUE Stellar Parameters Pipeline (SSPP).
 
-\section{Chemical Impact}
-Something about ISM enrichment and chemical output over time.
-Link the mass loss rate to the chemical content of the circumstellar shell.
-Perhaps calculate the chemical enrichment as a function of distance from the star?
+\begin{figure}
+\includegraphics[width=3in]{figs/sampleboundaries.jpg}
+\includegraphics[width=3in]{figs/contamboundaries.jpg}
+% Include a third figure that's the end result
+\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\label{fig:boundaries}
+\end{figure}
 
-\section{Time Frame}
-2 years from the end of the general exam.
-Major points to hit:
-\begin{itemize}
-\item Selection of AGB candidates from WISE data
-\item Split into high likelihood O-rich and C-rich populations.
-\item Figure out someway to relax selection criteria to potentially get more stars.  Consider looking at candidates from both WISE All-Sky and ALLWISE (confusing).
-\end{itemize}
+The color-color criteria set are fairly simple (Figure~\ref{fig:boundaries}), expanding upon regions in color-color space encountered in the literature.
+While we wish to maximize the completeness of our AGB star sample, we emphasize reliability of the sample and so minimize contamination below the 1\% level for each contaminant population and in total.
+We note that in our selection criteria we exclude a substantial population of AGB stars with small or non-existent circumstellar shells.
+This is necessary in order to also prevent stars from the stellar locus, a far more populous sample, from contaminating our candidates.
+
+\subsection{Estimation of Distances to Candidates with {\tt GALFAST}}
+%Create a sample of likely AGB star candidates and use GALFAST to estimate the dust column and distance to each candidate, using the color-color criteria of \cite{2014MNRAS.442.3361N} to differentiate between AGB stars with O-rich and C-rich atmospheres.
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/bulge_lmc_galcoords.pdf}
+\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\label{fig:galplot}
+\end{figure}
+
+As a first step toward creating a 3D map of our galaxy using candidate AGB stars, we set out to find reliable color-magnitude relationships for our stars.
+From our known sample, we select objects that fit within the narrow color-color criteria for O-rich and C-rich AGB stars set by \cite{2014MNRAS.442.3361N}.
+We then choose stars in the direction of regions of a) known distance and b) high star concentration in order to derive our color-magnitude relationships for these stars: the Milky Way bulge and the Large Magellanic Cloud (Figure~\ref{fig:galplot}).
+For the LMC, we minimize foreground contamination by selecting stars with $b < -20^\circ$ and also meeting the following criteria:
+
+$$\displaystyle\left(\frac{\text{RA} - \text{RA}_\text{LMC}}{3.5}\right)^2 + \left(\frac{\text{Dec} - \text{Dec}_\text{LMC}}{1.5} \right)^2 < 5 \text{ deg}^2.$$
+
+\noindent With this selection of stars, we demonstrate that AGB stars possess an intrinsically-narrow magnitude distribution, as seen in \cite{1985A&A...152L...1H} and \cite{2002MNRAS.337..749J}.
+We however do not use this sample further to define the color-magnitude relationships for AGB stars in the IR as our {\tt GALFAST} model used for estimating the dust columns for stars doesn't include dust in or the shape of the LMC.
+We neglect to use the Small Magellanic Cloud entirely due to the low count of recovered objects from our initial color-color criteria.
+We instead solely use the sample in the direction of the Milky Way bulge for the necessary color-magnitude calibration (Figure~\ref{fig:colormag}).
+We minimize the contamination from foreground stars in the direction of the Galactic bulge by selecting stars with $\lvert l\rvert < 10^\circ$, $1^\circ < \lvert b \rvert < 5^\circ$.
+
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/bulge_lmc_colormag.pdf}
+\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\label{fig:colormag}
+\end{figure}
+
+Color-magnitude relations are then found for each IR color with respect to $M_{w1}$ as derived from the Milky Way bulge using linear regression.
+With these we estimate distances to our candidate sample to within a small degree of uncertainty ($\sim$0.5 mag).
+
+Once a candidate sample has been retrieved from the ALLWISE-2MASS database matching our broad color criteria, we narrow them into high-photometric-quality O-rich and C-rich populations using the same criteria from \cite{2014MNRAS.442.3361N}.
+Assuming absolute magnitudes based upon the aforementioned color-magnitude relationships, we simultaneously estimate the distance and coefficient for dust extinction along the line of sight for each star using a convolution of the {\tt GALFAST} model and the dust maps from \cite{1998ApJ...500..525S}.
+
+\subsection{The Milky Way in 3D and Disk Measurements}
+%Create a 3D map of the galaxy using the aforementioned distances.  Measure the disk scale height as a function of radius, and the disk scale length as a function of height (if possible).
+With distance information and position on the sky, we construct an X-Y-Z map of the Milky Way (Figure~\ref{fig:xyz_candidates}).
+We note that the overwhelming majority of AGB candidates in the Milky Way are O-rich, with the population C-rich AGB candidates of high photometric quality being essentially nonexistent, so we move forward solely with an analysis of candidates classified as such.
+We also note, as is clear in Figure~\ref{fig:xyz_candidates}, that this is by no means a complete picture of the Milky Way.
+The sample is heavily affected by dust extinction and source confusion within 0.5 kpc of the galactic disk, and entirely beyond the bulge within $|b| < 1^\circ$ of the disk.
+Additionally, because we exclude sources that are beyond the saturation limits of WISE, we lose a large fraction of AGB candidates near the Sun (more toward the galactic anticenter due to a lack of interstellar dust to attenuate the source's intrinsic brightness).
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=5in]{figs/orich_candidates_xyz.pdf}
+\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\label{fig:xyz_candidates}
+\end{figure}
+
+Despite the loss of sources beyond the bulge and near the Sun, we are still able to make measurements of some common disk properties.
+We focus on the +X half of the galaxy due to the higher completeness of sources, and exclude the wedge of the Milky Way containing that solar circle.
+We then examine the radial and vertical distribution of AGB candidates, and derive measurements of the scale height as a function of radius, as well as the scale length as a function of height (Figure~\ref{fig:horvert_profile}).
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=6.5in]{figs/orich_radial_vertical_profile.pdf}
+\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\label{fig:horvert_profile}
+\end{figure}
+
+\section{Work to be Completed}
+\subsection{Comparison with {\tt TRILEGAL}}
+%Same as above, but use TRILEGAL and compare
+{\tt TRILEGAL} \citep{2005A&A...436..895G, 2007ASPC..378...20G} provides us with the possibility of testing several quantities produced from the work done thus far.
+{\tt GALFAST} provided us with an estimation of the dust column along the line of sight to our candidate AGB stars, but with a set galactic model.
+{\tt TRILEGAL}  allows us to test the validity of that model, alongside several other models of the Milky Way, with respect to our now-derived 3D distribution of AGB stars by providing stellar number densities along lines of sight.
+
+With the simulated photometry output by {\tt TRILEGAL}, along with various choices of dust chemistry used as input for these simulations, we can use our color-color information as proxies for measurements of photospheric temperature and mass-loss rate, measurements unobtainable from the photometry of ALLWISE.
+{\tt TRILEGAL} also has the added benefit of producing photometry for any given photometric set.
+Thus, we would be able to simulate photometry from the GLIMPSE survey, correlate that with photometry from ALLWISE, and use GLIMPSE photometry to fill in the population of AGB stars missing from our current sample due to source confusion and dust extinction.
+
+Finally, {\tt TRILEGAL} also produces estimations of stellar age, metallicity, [C/O], and periodicity.
+Convolved with our existing ALLWISE results and potential GLIMPSE results, we may be able to produce a 3D picture of chemical enrichment and age of the Milky Way, while also being able to see the chemical contribution of AGB stars to their local galactic neighborhoods, giving us an even fuller picture of the evolution of our galaxy.
+
+\subsection{Derive Relations Between SDSS Stellar Properties and Position}
+%Match candidate sample to APOGEE red giant survey to get actual stellar parameters as a function of position.
+Use above relationships to estimate circumstellar shell size, age, mass-loss rate, metallicity, anything really.
+Correlate that with X-Y-Z position in the galaxy (or maybe even just R-Z).
+We have matched the ALLWISE AGB candidate sample to the latest results from the SDSS APOGEE survey.
+This provides us with actual measurements of stellar parameters that we can use to characterize our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
+
+\subsection{Conclusion About Galactic Evolution w.r.t. SDSS and TRILEGAL}
+%Tie the above to galactic evolution. Try to pick out large structures in the Milky Way such as the Sagittarius over density and the Monoceros Overdensities. Should at least stand out in metallicity.
+
+\section{Timeline}
+Time it takes for each bit

### 2014-11-05 thesispropbody_d1.tex
Updating the thesis

@@
 %-----------
+%Fig 1
 \vspace{-0.41in}
@@
       \hrule \smallskip
-        {\footnotesize\baselineskip0.4cm{{\cap Figure 1:} {
+        {\footnotesize\baselineskip0.4cm{{\cap Figure 0:} {
         The variation 
@@
 \section{Work Completed/In Progress}
+Not since the series of SDSS tomography papers have we been able to obtain information about large scale physical and chemical characteristics of the Milky Way galaxy.
+Even then, the coverage of SDSS in the plane of the galaxy, and even in the galactic halo, was limited to the stripes of observation of SDSS.
+ALLWISE gives us the advantage of being able to observe the whole sky.
+And while the NIR and MIR are not immune to dust extinction along the line of sight, the influence of dust on observations is minimal as compared to optical wavelengths.
+
+If we're working in the NIR and Mid-IR, AGB stars are the perfect objects to serve as a proxy for structure of the Milky Way.
+AGB stars with dusty circumstellar shells shine brightly in the IR, with that brightness only increasing with the size of the shell.
+They're bright enough to be observed not only at the galactic center (even with dust extinction), but beyond out to the edge of the galactic disk.
+They're also ubiquitous throughout the galaxy, being the evolutionary products of low- and intermediate-mass Main Sequence stars.
+They have massive stellar winds that enrich their local galactic neighborhood, and their chemical content can represent the metallicity of the region in which they reside.
+As such, our work here is first to obtain a clean, reliable sample of these stars within the galaxy.
+Once obtained, we can then use the sample of ALLWISE AGB candidates in conjunction with high-confidence models of the galaxy and of the consequences of AGB star evolution to describe the Milky Way in detail, parameterized by the physical characteristics of these AGB stars.
+
 \subsection{Development of Broad Color-Color Criteria}
-%Using positional matches between SIMBAD, MACHO, and OGLE to ALLWISE-2MASS, create simple color-color, color-magnitude, and photometric quality criteria to select AGB stars in the ALLWISE-2MASS data set.
 We begin by selecting a known sample of AGB stars and matching to the ALLWISE-2MASS data release.
@@
 \begin{figure}
-\includegraphics[width=3in]{figs/sampleboundaries.jpg}
-\includegraphics[width=3in]{figs/contamboundaries.jpg}
-% Include a third figure that's the end result
-\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+% Fig 2
+\includegraphics[width=2in]{figs/sampleboundaries.jpg}
+\includegraphics[width=2in]{figs/contamboundaries.jpg}
+\includegraphics[width=2in]{figs/sample_results.jpg}
+\caption{\emph{Left}: SIMBAD, OGLE-III, and MACHO AGB stars (black, red, blue) matched to ALLWISE within 1".
+\emph{Middle}: LRGs, starforming galaxies, AGN, QSO, and Locus Stars (green, red, orange, blue, black).
+\emph{Right}: AGB stars passing ALLWISE color criteria.}
 \label{fig:boundaries}
@@
 \begin{figure}[h]
+% Fig 3
 \includegraphics[width=6in]{figs/bulge_lmc_galcoords.pdf}
-\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\caption{Galactic plot of all AGB stars passing color criteria, highlighting stars in the direction of the bulge (red) and stars in the LMC (blue).}
 \label{fig:galplot}
@@
 \begin{figure}[h]
+% Fig 4
 \includegraphics[width=6in]{figs/bulge_lmc_colormag.pdf}
-\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\caption{Absolute W1 magnitude vs 2MASS-ALLWISE colors for stars in the Milky Way bulge (top) and LMC (bottom). Stars are separated into O-rich (green) and C-rich (black).}
 \label{fig:colormag}
@@
 \begin{figure}[h]
+% Fig 5
 \centering
 \includegraphics[width=5in]{figs/orich_candidates_xyz.pdf}
-\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\caption{X-Y-Z distributions of the ALLWISE AGB candidate sample's O-rich stars.
+The Sun is set as the yellow dot with (X,Y,Z) = (8.0 kpc, 0 kpc, 0 kpc).
+The gray dashed lines mark (0,0) in each plot, and the contours are spaced by 0.75 logarithmically, starting at 0.}
 \label{fig:xyz_candidates}
@@
 \begin{figure}[h]
+% Fig 6
 \centering
@@
 \section{Work to be Completed}
+With the initial groundwork completed for the clean selection of the AGB candidate sample, dust extinction along the line of sight, and distance measurements throughout the Milky Way, we can proceed to convolve that data with high quality models of AGB star evolution and the distribution of stars throughout the Milky Way.
 \subsection{Comparison with {\tt TRILEGAL}}

### 2014-11-07 thesispropbody_d1.tex
Update 11/6/2014

@@
 In these, the Milky Way is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
-Recent discoveries of complex substructure in the distribution of the Milky Way's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V,2008ApJ...673..864J} have deeply shaken this standard view with indications of signs of rich substructure within the Milky Way galaxy.
+Recent discoveries of complex substructure in the distribution of the Milky Way's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V,2008ApJ...673..864J} have deeply shaken this standard view with indications of signs of rich substructure within the Milky Way Galaxy.
 This was due to the ability to map the Milky Way using the distances and positions of its own stellar residents.
@@
 \subsection*{\centerline{\ub{First Steps Toward Milky Way Tomography}}}
-Until recently, our knowledge of the basic structural components of the Milky Way was limited to indirect inferences based on stellar population models motivated by other galaxies, and what little information was available from the High Precision Parallax Collecting Satellite \cite[HIPPARCOS;][]{1984SSRv...39....1K} catalog and smaller pencil-beam surveys.
-The advent of the Sloan Digital Sky Survey \citep[SDSS;][]{2000AJ....120.1579Y} alleviated this limitation, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
+Until recently, our knowledge of the basic structural components of the Milky Way was limited to indirect inferences based on stellar population models motivated by other galaxies, and what little information was available from the \emph{High Precision Parallax Collecting Satellite} \cite[HIPPARCOS;][]{1984SSRv...39....1K} catalog and smaller pencil-beam surveys.
+The advent of the \emph{Sloan Digital Sky Survey} \citep[SDSS;][]{2000AJ....120.1579Y} alleviated this limitation, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
 SDSS photometry enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
@@
 %Get this figure and import it here
+
+\begin{center}\textbf{\textcolor{red}{THIS STILL NEEDS WORK AND RESTRUCTURING}}\end{center}
 
@@
         volume relative to the rest of Galaxy, and the background image is the Andromeda 
-        galaxy. The three squares outline the regions used to measure the 3-dimensional
+        Galaxy. The three squares outline the regions used to measure the 3-dimensional
         velocity distribution for halo stars. The arrows illustrate the variation of the velocity
@@
 
-%These recent developments, based on a host of accurate large-area surveys such as the Two-Micron All-Sky Survey \citep[2MASS;][]{2006AJ....131.1163S} and the Sloan Digital Sky Survey \citep[SDSS;][]{2000AJ....120.1579Y}, have made it abundantly clear that the Milky Way is a complex and dynamical structure that is still being shaped by the merging of smaller galaxies.
-Numerical simulations suggest that this merger process plays a crucial role in setting the structure and motions of stars within galaxies.
-Since the individual stars that make up the stellar populations in the Milky Way can be studied in great detail, their characterization proposed here will provide clues about galaxy evolution that cannot be extracted from observations of distant galaxies.
-%
-%To that end, I propose a two-year study to:
-%\begin{itemize}
-%\item mine the extensive data set provided by the Wide-Field Infrared Survey Explorer \citep[WISE;][]{2010AJ....140.1868W}, relating observables to the \emph{Spitzer} Galactic Legacy Infrared Mid-Plane Survey Extraordinaire \citep[GLIMPSE;][]{2003PASP..115..953B,2009PASP..121..213C}, SDSS, 2MASS, and other surveys.
-%\item reliably identify and track a unique yet ubiquitous population of high-luminosity, IR-bright stars: Asymptotic Giant Branch (hereon AGB) stars.
-%\item map the structure of the Milky Way in unprecedented detail, with AGB stars as our indicators of distance and age.
-%\item use the distribution of AGB stars as observational constraints on models of the formation and evolution of the Milky Way disk.
-%\end{itemize}
+\section{Work Completed/In Progress}
+\label{sec:completedWork}
+The coverage of SDSS in the plane of the Galaxy, and even in the Galactic halo, is thus far limited to the regions targeted by the \emph{Sloan Extension for Galactic Understanding and Exploration} \citep[SEGUE, ][]{2012ApJ...757..166B}.
+More detailed studies of Galactic structure need to rely on the extensive observational capabilities of surveys of the entire sky, preferably with instruments that either display weak sensitivity to dust extinction along the line of sight or are immune.
+The all-sky survey conducted by the \emph{Wide-field Infrared Survey Explorer} \citep[WISE,][]{2012wise.rept....1C} gives us the advantage of being able to observe the whole sky in the infrared (IR) at 3.4, 4.6, 12, and 22$\mu$m with unprecedented depth and resolution.
+While the near- and mid-infrared (hereon NIR and MIR) are not immune to dust extinction along the line of sight, the influence of dust on observations is minimal as compared to optical wavelengths \citep{2012ApJ...757..166B}.
 
-\section{Work Completed/In Progress}
-Not since the series of SDSS tomography papers have we been able to obtain information about large scale physical and chemical characteristics of the Milky Way galaxy.
-Even then, the coverage of SDSS in the plane of the galaxy, and even in the galactic halo, was limited to the stripes of observation of SDSS.
-ALLWISE gives us the advantage of being able to observe the whole sky.
-And while the NIR and MIR are not immune to dust extinction along the line of sight, the influence of dust on observations is minimal as compared to optical wavelengths.
+While observations in the NIR and MIR allow us to minimize the influence of dust extinction, and using WISE (particularly the latest data products released as ALLWISE) provides access to the entire sky, a proper survey of the Galaxy requires reliable objects that can be observed throughout the entirety of the Milky Way.
+AGB stars are the perfect objects to serve as those probes of the Galaxy.
+AGB stars with dusty circumstellar shells shine brightly in the IR, particularly with the SiO present in the shells of O-rich stars and the AmC in shells of C-rich stars \citep{2011A&A...534A..79I}.
+They are bright enough to be detected by ALLWISE not only at the Galactic center (even with dust extinction), but beyond out to the edge of the Galactic disk \citep{2013RAA....13..323T}.
+They're also ubiquitous throughout the Galaxy, being the evolutionary products of low- and intermediate-mass Main Sequence stars \citep{1983ARA&A..21..271I}.
+They have massive stellar winds that enrich their local Galactic neighborhood, and their chemical content can represent the metallicity of the region in which they reside.
+As such, our work here is first to obtain a clean, reliable sample of these stars within the Galaxy.
+Once obtained, we can then use the sample of ALLWISE AGB candidates in conjunction with high-confidence models of the Galaxy and of the consequences of AGB star evolution to describe the Milky Way in detail, parameterized by the physical characteristics of these AGB stars.
 
-If we're working in the NIR and Mid-IR, AGB stars are the perfect objects to serve as a proxy for structure of the Milky Way.
-AGB stars with dusty circumstellar shells shine brightly in the IR, with that brightness only increasing with the size of the shell.
-They're bright enough to be observed not only at the galactic center (even with dust extinction), but beyond out to the edge of the galactic disk.
-They're also ubiquitous throughout the galaxy, being the evolutionary products of low- and intermediate-mass Main Sequence stars.
-They have massive stellar winds that enrich their local galactic neighborhood, and their chemical content can represent the metallicity of the region in which they reside.
-As such, our work here is first to obtain a clean, reliable sample of these stars within the galaxy.
-Once obtained, we can then use the sample of ALLWISE AGB candidates in conjunction with high-confidence models of the galaxy and of the consequences of AGB star evolution to describe the Milky Way in detail, parameterized by the physical characteristics of these AGB stars.
-
-\subsection{Development of Broad Color-Color Criteria}
+\subsection{Development of Initial Color Criteria for Selecting AGB Stars}
 We begin by selecting a known sample of AGB stars and matching to the ALLWISE-2MASS data release.
@@
 The region of color-color space occupied by these stars is also occupied by other objects that happen to have dusty environments, as well as naked stars on the main sequence and extragalactic sources.
-As such, we need a reliable sample of contaminants that we can exclude in color-color space to increase the likelihood that our population of candidates represents the objects we wish to study.
+Thus we need a reliable sample of contaminants that we can exclude in color-color space to increase the likelihood that our candidates represent the objects we wish to study.
 We gather AGN, QSOs, and star forming galaxies from the Sloan Digital Sky Survey \citep{2000AJ....120.1579Y} data release (DR) 7, specifically the NYU Value-Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}, and the LRGs from the SDSS Luminous Red Galaxy Survey \citep{2010ApJ...710.1444K}.
@@
 
-The color-color criteria set are fairly simple (Figure~\ref{fig:boundaries}), expanding upon regions in color-color space encountered in the literature.
+The color-color criteria set are fairly simple (Figure~\ref{fig:boundaries}): $J-K_s > 1.1$, $W2-W3 > 0.3$, and $W3-W4 < -0.83(W2-W3) + 3.37$.
 While we wish to maximize the completeness of our AGB star sample, we emphasize reliability of the sample and so minimize contamination below the 1\% level for each contaminant population and in total.
@@
 
-\subsection{Estimation of Distances to Candidates with {\tt GALFAST}}
+\subsection{Estimation of Distances to Candidates with {\tt GALFASTDUST}}
 %Create a sample of likely AGB star candidates and use GALFAST to estimate the dust column and distance to each candidate, using the color-color criteria of \cite{2014MNRAS.442.3361N} to differentiate between AGB stars with O-rich and C-rich atmospheres.
@@
 
-As a first step toward creating a 3D map of our galaxy using candidate AGB stars, we set out to find reliable color-magnitude relationships for our stars.
+As a first step toward creating a 3D map of our Galaxy using candidate AGB stars, we set out to find reliable color-magnitude relationships for our stars.
 From our known sample, we select objects that fit within the narrow color-color criteria for O-rich and C-rich AGB stars set by \cite{2014MNRAS.442.3361N}.
@@
 \noindent With this selection of stars, we demonstrate that AGB stars possess an intrinsically-narrow magnitude distribution, as seen in \cite{1985A&A...152L...1H} and \cite{2002MNRAS.337..749J}.
-We however do not use this sample further to define the color-magnitude relationships for AGB stars in the IR as our {\tt GALFAST} model used for estimating the dust columns for stars doesn't include dust in or the shape of the LMC.
+We however do not use this sample further to define the color-magnitude relationships for AGB stars in the IR.
+The LSST dust extinction code that we use to estimate the dust extinction along the line of sight (hereon {\tt GALFASTDUST}) doesn't include dust in or the shape of the LMC.
 We neglect to use the Small Magellanic Cloud entirely due to the low count of recovered objects from our initial color-color criteria.
@@
 \includegraphics[width=6in]{figs/bulge_lmc_colormag.pdf}
-\caption{Absolute W1 magnitude vs 2MASS-ALLWISE colors for stars in the Milky Way bulge (top) and LMC (bottom). Stars are separated into O-rich (green) and C-rich (black).}
+\caption{Absolute W1 magnitude vs 2MASS-ALLWISE colors for stars in the Milky Way bulge (top) and LMC (bottom). Stars are separated into O-rich (blue) and C-rich (red).}
 \label{fig:colormag}
@@
 Color-magnitude relations are then found for each IR color with respect to $M_{w1}$ as derived from the Milky Way bulge using linear regression.
-With these we estimate distances to our candidate sample to within a small degree of uncertainty ($\sim$0.5 mag).
+With $W1$ directly obtained from ALLWISE and $M_\text{w1}$ calculated with respect to the Milky Way bulge, we use {\tt GALFASTDUST} to simultaneously estimate distance moduli ($DM$) and extinction coefficients ($A_\lambda$) for each object.
+{\tt GALFASTDUST} returns the color excess $E(B-V)$ for a given ($l$, $b$) and distance $D$ in kpc by placing the object within the axisymmetric galaxy model used by \cite{2005AJ....130..659A}, normalizing $E(B-V)$ to the dust maps of \cite{1998ApJ...500..525S}, and interpolating the dust column along the line of sight.
+We convert $E(B-V)$ to $A_\lambda$ using the relations in the appendix of \cite{2014MNRAS.440.3430D}.
+These estimated $DM$ lie within a small degree of uncertainty ($\sim$0.5 mag), due to scatter about the initial color-magnitude relationship seen in Figure~\ref{fig:colormag}.
 
 Once a candidate sample has been retrieved from the ALLWISE-2MASS database matching our broad color criteria, we narrow them into high-photometric-quality O-rich and C-rich populations using the same criteria from \cite{2014MNRAS.442.3361N}.
-Assuming absolute magnitudes based upon the aforementioned color-magnitude relationships, we simultaneously estimate the distance and coefficient for dust extinction along the line of sight for each star using a convolution of the {\tt GALFAST} model and the dust maps from \cite{1998ApJ...500..525S}.
+Assuming absolute magnitudes based upon the aforementioned color-magnitude relationships, we simultaneously estimate the distance and coefficient for dust extinction along the line of sight for each star using {\tt GALFASTDUST}, which includes the dust maps of .
 
 \subsection{The Milky Way in 3D and Disk Measurements}
-%Create a 3D map of the galaxy using the aforementioned distances.  Measure the disk scale height as a function of radius, and the disk scale length as a function of height (if possible).
+%Create a 3D map of the Galaxy using the aforementioned distances.  Measure the disk scale height as a function of radius, and the disk scale length as a function of height (if possible).
 With distance information and position on the sky, we construct an X-Y-Z map of the Milky Way (Figure~\ref{fig:xyz_candidates}).
@@
 We also note, as is clear in Figure~\ref{fig:xyz_candidates}, that this is by no means a complete picture of the Milky Way.
-The sample is heavily affected by dust extinction and source confusion within 0.5 kpc of the galactic disk, and entirely beyond the bulge within $|b| < 1^\circ$ of the disk.
-Additionally, because we exclude sources that are beyond the saturation limits of WISE, we lose a large fraction of AGB candidates near the Sun (more toward the galactic anticenter due to a lack of interstellar dust to attenuate the source's intrinsic brightness).
+The sample is heavily affected by dust extinction and source confusion within 0.5 kpc of the Galactic disk, and entirely beyond the bulge within $|b| < 1^\circ$ of the disk.
+Additionally, because we exclude sources that are beyond the saturation limits of WISE, we lose a large fraction of AGB candidates near the Sun (more toward the Galactic anticenter due to a lack of interstellar dust to attenuate the source's intrinsic brightness).
 
@@
 Despite the loss of sources beyond the bulge and near the Sun, we are still able to make measurements of some common disk properties.
-We focus on the +X half of the galaxy due to the higher completeness of sources, and exclude the wedge of the Milky Way containing that solar circle.
+We focus on the +X half of the Galaxy due to the higher completeness of sources, and exclude the wedge of the Milky Way containing that solar circle.
 We then examine the radial and vertical distribution of AGB candidates, and derive measurements of the scale height as a function of radius, as well as the scale length as a function of height (Figure~\ref{fig:horvert_profile}).
@@
 \includegraphics[width=6.5in]{figs/orich_radial_vertical_profile.pdf}
-\caption{Lorem ipsum dolor sit amet, consectetur elit. Fusce, nulla rebds vel pellentesque consequat, ante nulla hendrerit arcu, ac tincidunt mauris lacus sed leo.}
+\caption{Radial (left) and vertical (right) distributions of O-rich AGB candidates in the Milky Way.
+The top plots show the distributions for all recovered stars, and the bottom plots show the distributions for various heights (left) and radii (right).
+The radial distributions are in 1 kpc bins, and vertical in 0.2 kpc bins.
+Uncertainties in each bin are $\sqrt{N}$.}
 \label{fig:horvert_profile}
@@
 \section{Work to be Completed}
-With the initial groundwork completed for the clean selection of the AGB candidate sample, dust extinction along the line of sight, and distance measurements throughout the Milky Way, we can proceed to convolve that data with high quality models of AGB star evolution and the distribution of stars throughout the Milky Way.
+With the initial groundwork completed for the clean selection of the AGB candidate sample, dust extinction along the line of sight, and distance measurements throughout the Milky Way, we can proceed to convolve that data with spectroscopic observations from the SDSS \emph{APO Galactic Evolution Experiment} (APOGEE) as well as high quality models of AGB star evolution and the distribution of stars throughout the Milky Way from TRILEGAL.
+We can also investigate those objects not fitting the pre-defined O-rich and C-rich classification schemes.
+\subsection{Investigate Unclassified AGB Candidates}
+ \cite{2014MNRAS.442.3361N} developed criteria to separate O-rich and C-rich AGB stars based on results from simulations with {\tt DUSTY}, demonstrating that each population should fall along fairly narrow tracks in WISE color-color space.
+As previously mentioned, we use these classification schemes to derive color-magnitude relationships for our candidates and produce our Galactic maps.
+
+ However, as we have shown in Figure~\ref{fig:boundaries}, we can explore wide regions of WISE color-color space and still eliminate the vast majority of contaminants down to the $<$1\% level.
+While this initial effort yields $\sim$215,000 objects qualifying as AGB candidates, the final sample used for the mapping the galaxy numbered only $\sim$69,000 objects.
+This begs the question, what were the other $\sim$150,000 objects, and why did they manage to pass the AGB color-color criteria without explicitly falling in the pre-defined O-rich and C-rich tracks?
+We seek to answer this by exploring other possible contaminants such as R Cor Borealis stars, young stellar objects, and planetary nebulae \citep{2002MNRAS.337..749J} through positional matches to the SIMBAD database, and verification of AGB candidacy with data from SDSS APOGEE (see section~\ref{sec:apogee}).
+
 \subsection{Comparison with {\tt TRILEGAL}}
@@
 {\tt TRILEGAL} \citep{2005A&A...436..895G, 2007ASPC..378...20G} provides us with the possibility of testing several quantities produced from the work done thus far.
-{\tt GALFAST} provided us with an estimation of the dust column along the line of sight to our candidate AGB stars, but with a set galactic model.
+{\tt GALFASTDUST} provided us with an estimation of the dust column along the line of sight to our candidate AGB stars, but with a set Galactic model.
 {\tt TRILEGAL}  allows us to test the validity of that model, alongside several other models of the Milky Way, with respect to our now-derived 3D distribution of AGB stars by providing stellar number densities along lines of sight.
@@
 Finally, {\tt TRILEGAL} also produces estimations of stellar age, metallicity, [C/O], and periodicity.
-Convolved with our existing ALLWISE results and potential GLIMPSE results, we may be able to produce a 3D picture of chemical enrichment and age of the Milky Way, while also being able to see the chemical contribution of AGB stars to their local galactic neighborhoods, giving us an even fuller picture of the evolution of our galaxy.
+Convolved with our existing ALLWISE results and potential GLIMPSE results, we may be able to produce a 3D picture of chemical enrichment and age of the Milky Way, while also being able to see the chemical contribution of AGB stars to their local Galactic neighborhoods, giving us an even fuller picture of the evolution of our Galaxy.
 
-\subsection{Derive Relations Between SDSS Stellar Properties and Position}
+\subsection{ALLWISE AGB Candidates and APOGEE}
+\label{sec:apogee}
 %Match candidate sample to APOGEE red giant survey to get actual stellar parameters as a function of position.
 Use above relationships to estimate circumstellar shell size, age, mass-loss rate, metallicity, anything really.
-Correlate that with X-Y-Z position in the galaxy (or maybe even just R-Z).
-We have matched the ALLWISE AGB candidate sample to the latest results from the SDSS APOGEE survey.
-This provides us with actual measurements of stellar parameters that we can use to characterize our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
+Correlate that with X-Y-Z position in the Galaxy (or maybe even just R-Z).
+The SDSS APOGEE survey observed $10^5$ giant stars down to magnitude $H$ = 12.2 with IR spectroscopy \citep{2012ApJ...755L..25N}.
+We have matched the ALLWISE AGB candidate sample to the latest results from the APOGEE survey, retrieving $\sim$8,000 objects positionally matching our cleanest sample.
+This provides us with precise measurements of stellar parameters that we can use to characterize our sample, provide measurable checks as to the reliability of our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
 
-\subsection{Conclusion About Galactic Evolution w.r.t. SDSS and TRILEGAL}
-%Tie the above to galactic evolution. Try to pick out large structures in the Milky Way such as the Sagittarius over density and the Monoceros Overdensities. Should at least stand out in metallicity.
+\begin{center}\textbf{\textcolor{red}{MAYBE FIGURE SHOWING DISTRIBUTION OF MATCHES?}}\end{center}
+
+With the input from APOGEE, we will correlate various stellar properties (e.g.
+$T_\text{eff}$, $[Fe/H]$) with ALLWISE-2MASS colors for our AGB candidate sample.
+ALLWISE-2MASS IR colors, particularly $W2-W3$ and $W3-W4$, are sensitive to the composition and size of the AGB circumstellar shell, making chemical differences separable in ALLWISE color-color space as demonstrated by \cite{2014MNRAS.442.3361N}.
+We should then be able to use standard machine learning techniques \citep[e.g.
+those found in ][]{2013sdmm.book.....I} to extend estimates of these same stellar parameters to stars that were not initially matched to APOGEE due to lying outside either the photometric or spatial range of APOGEE.
+
+Comparison of the candidate sample to APOGEE will also provide us with a test against the axisymmetric Galactic model used by {\tt GALFASTDUST}.
+Obviously the Milky Way is not a simple axisymmetric system, containing large structures such as spiral arms, as well as the various overdensities observed by \cite{2008ApJ...673..864J} and others that have followed.
+With measurements from APOGEE we will be able to comment on the accuracy of our previous 3D distribution and improve upon observations of substructure in the Milky Way.
+
+\subsection{Conclusions about Galactic Evolution with respect to SDSS and TRILEGAL}
+%Tie the above to Galactic evolution. Try to pick out large structures in the Milky Way such as the Sagittarius over density and the Monoceros Overdensities. Should at least stand out in metallicity.
+With both sets of confirmed and estimated stellar properties, we will be able to create an even more informative 3D map of AGB characteristics (and by proxy, Galactic characteristics).
+
+\begin{center}\textbf{\textcolor{red}{NEED TO ADD STUFF}}\end{center}
 
 \section{Timeline}
-Time it takes for each bit
+\begin{enumerate}
+\item Publish current results of AGB candidate catalog and Galactic distribution discussed in section~\ref{sec:completedWork}, with publicly accessible data -- Q1 2015.
+\item Investigating unclassified AGB candidates -- Q1 2015
+\item Comparison with {\tt TRILEGAL} -- Q2-Q3 2015.
+\item Estimates of stellar parameters from APOGEE -- Q1 2016.
+\item Publish results of comparisons and estimates from {\tt TRILEGAL} and APOGEE -- Q2-Q3 2016
+\item Defense -- Q3-Q4 2016
+\end{enumerate}

### 2014-11-08 thesispropbody_d1.tex
Update 11/7/2014

@@
 Once a candidate sample has been retrieved from the ALLWISE-2MASS database matching our broad color criteria, we narrow them into high-photometric-quality O-rich and C-rich populations using the same criteria from \cite{2014MNRAS.442.3361N}.
-Assuming absolute magnitudes based upon the aforementioned color-magnitude relationships, we simultaneously estimate the distance and coefficient for dust extinction along the line of sight for each star using {\tt GALFASTDUST}, which includes the dust maps of .
+Assuming absolute magnitudes based upon the aforementioned color-magnitude relationships, we simultaneously estimate the distance and coefficient for dust extinction along the line of sight for each star using {\tt GALFASTDUST}.
 
@@
 \begin{enumerate}
-\item Publish current results of AGB candidate catalog and Galactic distribution discussed in section~\ref{sec:completedWork}, with publicly accessible data -- Q1 2015.
-\item Investigating unclassified AGB candidates -- Q1 2015
-\item Comparison with {\tt TRILEGAL} -- Q2-Q3 2015.
-\item Estimates of stellar parameters from APOGEE -- Q1 2016.
-\item Publish results of comparisons and estimates from {\tt TRILEGAL} and APOGEE -- Q2-Q3 2016
-\item Defense -- Q3-Q4 2016
+\item Paper 1: Publication of AGB candidate catalog and preliminary measurements of Milky Way disk and bulge -- \textbf{End of March 2015}
+	\begin{itemize}
+		\item Sample selection completeness and contamination estimates
+		\item O-rich/C-rich classification from color-color criteria
+		\item Identification of unclassified candidates
+		\item Estimates of unconfirmed classifications with Gaussian Mixture Models
+		\item Measurements of disk scale height/length
+		\item Measurements of the sizes of the Galactic bar \& bulge and orientation of the Galactic bar
+		\item Measurement of radius of Galactic disk
+	\end{itemize}
+
+\item Paper 2: Characteristics of the Milky Way galaxy via ALLWISE AGB candidates -- \textbf{End of Sept 2015}
+	\begin{itemize}
+		\item Mass-loss rate, model metallicity, temperature, size, and age of AGB candidates with respect to position
+		\item Interpretations of position-dependent characteristics
+		\item Meaning of these results with respect to improvements in AGB evolutionary tracks
+	\end{itemize}
+
+\item Post doctorate and Job Applications -- \textbf{Oct-Nov 2015}
+\item Paper 3: AGB stellar parameters from APOGEE and extrapolation to entire candidate sample -- \textbf{May 2016}
+	\begin{itemize}
+		\item Matches between ALLWISE candidates and APOGEE targets
+		\item Color-parameter relationships between APOGEE and ALLWISE
+		\item Extrapolation to entire candidate sample through color-parameter relationships
+		\item Comparison to estimations from various machine-learning techniques
+		\item Comparison to estimations from TRILEGAL
+	\end{itemize}
+\item Dissertation preparation -- \textbf{June - August 2016}
+\item Defense -- \textbf{August/September 2016}
 \end{enumerate}

### 2014-11-13 thesispropbody_d1.tex
Submitting for Draft 1

@@
 This was due to the ability to map the Milky Way using the distances and positions of its own stellar residents.
-I propose to continue the mapping of the Milky Way in further detail, using the ubiquitous stars of the Asymptotic Giant Branch (hereon AGB).
+I propose to expand the mapping of the Milky Way, using the ubiquitous stars of the Asymptotic Giant Branch (hereon AGB) as probes to its spatial and chemical structure structure.
 
-\subsection*{\centerline{\ub{First Steps Toward Milky Way Tomography}}}
+\subsection{First Steps Toward Milky Way Tomography}
 Until recently, our knowledge of the basic structural components of the Milky Way was limited to indirect inferences based on stellar population models motivated by other galaxies, and what little information was available from the \emph{High Precision Parallax Collecting Satellite} \cite[HIPPARCOS;][]{1984SSRv...39....1K} catalog and smaller pencil-beam surveys.
@@
 The resulting maps provided a quantitative basis for analysis of the main structural components of the Galaxy, enabled efficient searches for complex substructure, and a robust comparison with various model predictions \citep{2012ApJ...757..166B,2012ARA&A..50..251I}.
-An example of these results is reproduced in Figure 1, showing a panoramic view of the variation in the median $[Fe/H]$ over an unprecedentedly-large volume of the Galaxy.
+Figure 1 shows an example of a panoramic view of the variation in the median $[Fe/H]$ over an unprecedentedly-large volume of the Galaxy.
 %Get this figure and import it here
-
-\begin{center}\textbf{\textcolor{red}{THIS STILL NEEDS WORK AND RESTRUCTURING}}\end{center}
 
@@
 %Fig 1
-\vspace{-0.41in}
+%\vspace{-0.41in}
+\begin{figure}[h]
 \begin{minipage}{0.9\textwidth}
@@
       \begin{minipage}[l]{3in}
-      \hrule \smallskip
-        {\footnotesize\baselineskip0.4cm{{\cap Figure 0:} {
-        The variation 
-        of the median photometric stellar metallicity in cylindrical Galactic coordinates 
+%      \hrule \smallskip
+%        {\footnotesize\baselineskip0.4cm{{\cap Figure 0:} {
+%        The variation 
+%        of the median photometric stellar metallicity in cylindrical Galactic coordinates 
+%        $R$ and $|Z|$  for $\sim$2.5 million stars from SDSS with $14.5 < r <20$,
+%        $0.2 < g-r < 0.4$ (blue turn-off stars), and photometric distance in the 0.8-9 kpc 
+%        range %\citep[details in][]{2008ApJ...684..287I}. 
+%        Each bin 
+%        (50 pc by 50 pc) contained in this map is colored according
+%        to the legend in the top left. The gradient of the median metallicity is essentially 
+%        parallel to the $|Z|$ axis, except in the region of the Monoceros stream, as marked. 
+%        The gray scale background is the best-fit model for the stellar number density distribution 
+%        from \cite{2008ApJ...673..864J}. The inset in the top right illustrates the extent of the data 
+%        volume relative to the rest of Galaxy, and the background image is the Andromeda 
+%        Galaxy. The three squares outline the regions used to measure the 3-dimensional
+%        velocity distribution for halo stars. The arrows illustrate the variation of the velocity
+%        ellipsoid orientation, which always points toward the Galactic center. Adapted from 
+%        \cite{2012ARA&A..50..251I}.
+%\smallskip \hrule}}}
+     \caption{{\footnotesize
+        Variation of median photometric stellar metallicity in cylindrical Galactic coordinates 
         $R$ and $|Z|$  for $\sim$2.5 million stars from SDSS with $14.5 < r <20$,
         $0.2 < g-r < 0.4$ (blue turn-off stars), and photometric distance in the 0.8-9 kpc 
-        range %\citep[details in][]{2008ApJ...684..287I}. 
-        Each bin 
-        (50 pc by 50 pc) contained in this map is colored according
-        to the legend in the top left. The gradient of the median metallicity is essentially 
+        range. %\citep[details in][]{2008ApJ...684..287I}. 
+	The gradient of the median metallicity is essentially 
         parallel to the $|Z|$ axis, except in the region of the Monoceros stream, as marked. 
@@
         \cite{2012ARA&A..50..251I}.
-\smallskip \hrule}}}
-     %
+}}
+
       \end{minipage}\ \hfill \
       \begin{minipage}[r]{3in}
-        \epsfig{file=figs/IBJ_fig5.pdf, height=4.5in}
+	\includegraphics[height=4.5in]{figs/IBJ_fig5.pdf}
+%        \epsfig{file=figs/IBJ_fig5.pdf, height=4.5in}
+
       \end{minipage} \ \hfill \
@@
 \end{minipage}
-\vspace{-28pt}
+\vspace{-10pt}
+\end{figure}
+
 %------------
 
-\section{Work Completed/In Progress}
+\section{Work In Progress/Completed}
 \label{sec:completedWork}
@@
 They're also ubiquitous throughout the Galaxy, being the evolutionary products of low- and intermediate-mass Main Sequence stars \citep{1983ARA&A..21..271I}.
-They have massive stellar winds that enrich their local Galactic neighborhood, and their chemical content can represent the metallicity of the region in which they reside.
+They have massive stellar winds that enrich their local Galactic neighborhood, and their chemical content is representative of their local metallicity.
 As such, our work here is first to obtain a clean, reliable sample of these stars within the Galaxy.
@@
 % Fig 2
+\centering
 \includegraphics[width=2in]{figs/sampleboundaries.jpg}
@@
 \includegraphics[width=2in]{figs/sample_results.jpg}
-\caption{\emph{Left}: SIMBAD, OGLE-III, and MACHO AGB stars (black, red, blue) matched to ALLWISE within 1".
+\caption{{\footnotesize\emph{Left}: SIMBAD, OGLE-III, and MACHO AGB stars (black, red, blue) matched to ALLWISE within 1".
 \emph{Middle}: LRGs, starforming galaxies, AGN, QSO, and Locus Stars (green, red, orange, blue, black).
-\emph{Right}: AGB stars passing ALLWISE color criteria.}
+\emph{Right}: AGB stars passing ALLWISE color criteria.}}
 \label{fig:boundaries}
@@
 % Fig 3
+\centering
 \includegraphics[width=6in]{figs/bulge_lmc_galcoords.pdf}
-\caption{Galactic plot of all AGB stars passing color criteria, highlighting stars in the direction of the bulge (red) and stars in the LMC (blue).}
+\caption{{\footnotesize Galactic plot of all AGB stars passing color criteria, highlighting stars in the direction of the bulge (red) and stars in the LMC (blue).}}
 \label{fig:galplot}
@@
 \includegraphics[width=6in]{figs/bulge_lmc_colormag.pdf}
-\caption{Absolute W1 magnitude vs 2MASS-ALLWISE colors for stars in the Milky Way bulge (top) and LMC (bottom). Stars are separated into O-rich (blue) and C-rich (red).}
+\caption{{\footnotesize Absolute W1 magnitude vs 2MASS-ALLWISE colors for stars in the Milky Way bulge (top) and LMC (bottom). Stars are separated into O-rich (blue) and C-rich (red).}}
 \label{fig:colormag}
@@
 \includegraphics[width=5in]{figs/orich_candidates_xyz.pdf}
-\caption{X-Y-Z distributions of the ALLWISE AGB candidate sample's O-rich stars.
+\caption{{\footnotesize X-Y-Z distributions of the ALLWISE AGB candidate sample's O-rich stars.
 The Sun is set as the yellow dot with (X,Y,Z) = (8.0 kpc, 0 kpc, 0 kpc).
-The gray dashed lines mark (0,0) in each plot, and the contours are spaced by 0.75 logarithmically, starting at 0.}
+The gray dashed lines mark (0,0) in each plot, and the contours are spaced by 0.75 logarithmically, starting at 0.}}
 \label{fig:xyz_candidates}
@@
 \includegraphics[width=6.5in]{figs/orich_radial_vertical_profile.pdf}
-\caption{Radial (left) and vertical (right) distributions of O-rich AGB candidates in the Milky Way.
+\caption{{\footnotesize Radial (left) and vertical (right) distributions of O-rich AGB candidates in the Milky Way.
 The top plots show the distributions for all recovered stars, and the bottom plots show the distributions for various heights (left) and radii (right).
 The radial distributions are in 1 kpc bins, and vertical in 0.2 kpc bins.
-Uncertainties in each bin are $\sqrt{N}$.}
+Uncertainties in each bin are $\sqrt{N}$.}}
 \label{fig:horvert_profile}
@@
 \section{Work to be Completed}
-With the initial groundwork completed for the clean selection of the AGB candidate sample, dust extinction along the line of sight, and distance measurements throughout the Milky Way, we can proceed to convolve that data with spectroscopic observations from the SDSS \emph{APO Galactic Evolution Experiment} (APOGEE) as well as high quality models of AGB star evolution and the distribution of stars throughout the Milky Way from TRILEGAL.
+The initial groundwork is set for the clean selection of the AGB candidate sample, dust extinction along the line of sight, and distance measurements throughout the Milky Way.
+We can proceed to convolve that data with the IR spectroscopic observations from the SDSS \emph{APO Galactic Evolution Experiment} (APOGEE) as well as high quality models of AGB star evolution and Galactic distribution with {\tt TRILEGAL}.
 We can also investigate those objects not fitting the pre-defined O-rich and C-rich classification schemes.
 \subsection{Investigate Unclassified AGB Candidates}
- \cite{2014MNRAS.442.3361N} developed criteria to separate O-rich and C-rich AGB stars based on results from simulations with {\tt DUSTY}, demonstrating that each population should fall along fairly narrow tracks in WISE color-color space.
+ \cite{2014MNRAS.442.3361N} developed criteria to separate O-rich and C-rich AGB stars based on results from simulations with {\tt DUSTY} \citep{1999astro.ph.10475I}, demonstrating that each population should fall along fairly narrow tracks in WISE color-color space.
 As previously mentioned, we use these classification schemes to derive color-magnitude relationships for our candidates and produce our Galactic maps.
 
- However, as we have shown in Figure~\ref{fig:boundaries}, we can explore wide regions of WISE color-color space and still eliminate the vast majority of contaminants down to the $<$1\% level.
+ However, we can explore wide regions of WISE color-color space and still eliminate the selected contaminants down to the $<$1\% level (Figure~\ref{fig:boundaries}).
 While this initial effort yields $\sim$215,000 objects qualifying as AGB candidates, the final sample used for the mapping the galaxy numbered only $\sim$69,000 objects.
-This begs the question, what were the other $\sim$150,000 objects, and why did they manage to pass the AGB color-color criteria without explicitly falling in the pre-defined O-rich and C-rich tracks?
+This begs the question, what were the other $\sim$150,000 objects, and how did they pass the AGB-trained color-color criteria without explicitly falling in the pre-defined O-rich and C-rich tracks?
 We seek to answer this by exploring other possible contaminants such as R Cor Borealis stars, young stellar objects, and planetary nebulae \citep{2002MNRAS.337..749J} through positional matches to the SIMBAD database, and verification of AGB candidacy with data from SDSS APOGEE (see section~\ref{sec:apogee}).
@@
 Correlate that with X-Y-Z position in the Galaxy (or maybe even just R-Z).
-The SDSS APOGEE survey observed $10^5$ giant stars down to magnitude $H$ = 12.2 with IR spectroscopy \citep{2012ApJ...755L..25N}.
-We have matched the ALLWISE AGB candidate sample to the latest results from the APOGEE survey, retrieving $\sim$8,000 objects positionally matching our cleanest sample.
-This provides us with precise measurements of stellar parameters that we can use to characterize our sample, provide measurable checks as to the reliability of our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
+The SDSS APOGEE survey observed $1.6\times10^5$ giant stars down to magnitude $H$ = 12.2 with IR spectroscopy \citep{2012ApJ...755L..25N}.
+We have matched the ALLWISE AGB candidate sample to the latest results from the APOGEE survey, retrieving $\sim$8,500 objects positionally matching our cleanest sample to within 1", with most positions o (Figure~\ref{fig:apogee_allwise}).
+This provides us with precise measurements of stellar parameters unable to be obtained from ALLWISE that we can now use to characterize our sample, provide measurable checks as to the reliability of our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
 
-\begin{center}\textbf{\textcolor{red}{MAYBE FIGURE SHOWING DISTRIBUTION OF MATCHES?}}\end{center}
+\begin{figure}[h]
+% Fig 7
+\centering
+\includegraphics[width=6.5in]{figs/apogee_allwise_overlay.pdf}
+\caption{{\footnotesize APOGEE sources (gray) overlaid with matches within 1" from ALLWISE AGB candidates (red).
+\emph{Inset}: Distribution of matching radii between APOGEE and ALLWISE catalogs out to 1".}}
+\label{fig:apogee_allwise}
+\end{figure}
 
@@
 With measurements from APOGEE we will be able to comment on the accuracy of our previous 3D distribution and improve upon observations of substructure in the Milky Way.
-
-\subsection{Conclusions about Galactic Evolution with respect to SDSS and TRILEGAL}
-%Tie the above to Galactic evolution. Try to pick out large structures in the Milky Way such as the Sagittarius over density and the Monoceros Overdensities. Should at least stand out in metallicity.
-With both sets of confirmed and estimated stellar properties, we will be able to create an even more informative 3D map of AGB characteristics (and by proxy, Galactic characteristics).
-
-\begin{center}\textbf{\textcolor{red}{NEED TO ADD STUFF}}\end{center}
 
@@
 
-\item Post doctorate and Job Applications -- \textbf{Oct-Nov 2015}
+\item Post doctorate job applications -- \textbf{Oct-Nov 2015}
 \item Paper 3: AGB stellar parameters from APOGEE and extrapolation to entire candidate sample -- \textbf{May 2016}
@@
 \end{enumerate}
+
+\section{Potential Future Work}
+This work when completed will be just in time to set the stage for a few ambitious projects, and will be prime for building new and strengthening existing collaborative works.
+\subsection{From Candidates to Confirmations}
+AGB stars are known for their long-period variability of $\Delta V\sim2.5$ mag on timescales of $\sim$1 year \citep{2000PASA...17...18W, 2008PhDT.........6F}.
+With accurate time-series photometry these stars can be confirmed without the need for spectroscopy across tens of thousands of sources.
+A prime candidate for this all-sky time-series data source is the European Space Agency's GAIA space observatory \citep{2001A&A...369..339P,2002MNRAS.331..649B}.
+With unprecedented astrometry and photometric sensitivity in the visual and NIR, as well as between 40 and 250 visits per object over 5 years down to $V$ = 20 mag, it will be trivial for GAIA to identify long period variables in our existing candidate sample \citep{2000ASPC..203...71E}.
+
+\subsection{Investigating the TP-AGB and {\tt COLIBRI}}
+The AGB parameters output by version 2.3 of {\tt TRILEGAL} uses the population synthesis code {\tt COLIBRI}, which includes the thermally-pulsing (TP) phase of the AGB \citep{2013MNRAS.434..488M}.
+This code is calibrated using observations from the Magellanic Clouds, but would benefit greatly from a healthy sample of AGB stars from the Galaxy.
+APOGEE is already providing $\sim$8,300 of our objects with high-quality spectroscopic measurements, however as seen in Figure~\ref{fig:apogee_allwise} it neglects large portions of the Galaxy.
+With a large-scale spectroscopic program aimed at precise measurements of characteristics of the remaining AGB stars from this sample, we'd be able to contribute to this ongoing work with a thorough characteriziation of Galactic AGB stars.
+Specifically, we'd be able to accurately identify populations of TP-AGB stars as well as provide reliable measurements of the C/O ratios with position in the Milky Way alongside our existing IR measurements.

### 2015-02-03 paper1/paper1_abstract.tex
updating for paper 1

\begin{abstract}
We study the structure of the Milky Way disk with candidate asymptotic giant branch (AGB) stars selected from the \emph{Wide-field Infrared Survey Explorer} (\emph{WISE}) catalog. The advantages of our approach, compared to most recent similar works such as those based on SDSS data, are large distance limits due to the high luminosity of AGB stars, small interstellar dust obscuration due to longer wavelengths, and the all-sky coverage of the \emph{WISE} survey. The candidate AGB stars are color-selected with high completeness and low contamination, as quantified using samples of known AGB stars and other objects with known classifications from the SIMBAD and \emph{Sloan Digital Sky Survey} (\emph{SDSS}) databases. Distances to candidate AGB stars are estimated simultaneously with interstellar dust extinction along the line of sight using a 3-dimensional dust distribution model developed to support simulations for the \emph{Large Synoptic Survey Telescope} (\emph{LSST}) and a color -- absolute magnitude relation calibrated using the Large Magellanic Cloud (LMC) and the Galactic bulge. We find that the Galactic disk extends radially out to 15 kpc, with flaring of the disk towards its edge. We present measurements of the vertical scale height and horizontal scale length for double-exponential disk models. We find that the density distribution of AGB candidates within 9 kpc from the Galactic center is consistent with that of a double-exponential profile, while at larger radii the distribution is indistinguishable from a single-exponential profile.
\end{abstract}

### 2015-02-03 paper1/paper1_criteria.tex
updating for paper 1

\section{Object Selection Criteria}
\label{sec:criteria}
We made color-color criteria to create a catalog of AGB candidates. We kept it simple to maximize the number of objects we'd retain while minimizing contamination from non-AGB objects.

### 2015-02-03 paper1/paper1_data.tex
updating for paper 1

\section{Data Sources and Reduction}
\label{sec:data}
\subsection{Data Sources}
In order to analyze the physical distribution of \agb\, stars in the IR, we require reliable identifications of known \agb, stars to generate color-color selection criteria. We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}. \ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II. O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in \cite{2009AcA....59..239S,2011AcA....61..217S}. Data reduction techniques are described in \cite{2008AcA....58...69U}. The resulting samples yielded 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich). From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars). The sample of AGB stars from \simbad\, was obtained by querying for all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars. The total sample of \agb\, stars is 100,393.

Since \sdss\, is an optical survey of unprecedented extent and sensitivity, we use \sdss\, spectroscopic catalogs to select contaminant sources for the NIR-MIR color-color fields of \agb\, stars. Data for active galactic nuclei (AGN; 19,184 objects), quasi-stellar objects (QSOs; 122,550 objects), and star forming/burst galaxies (820,272 objects total) were drawn from \sdss\, DR7, specifically from the NYU Value Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}. Luminous Red Galaxies (LRGs) were selected from the SDSS Luminous Red Galaxy Survey \citep[105,631 objects, ][]{2010ApJ...710.1444K}.  Data for stars in the SDSS stellar locus were drawn from the DR 9 SEGUE Stellar Parameters Pipeline (SSPP) \citep[1,843,190 objects, ][]{2012ApJS..203...21A}.

In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}. The initial \wise All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m. Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$]. The positions of objects in the \wise\, catalog were calibrated to the \twomass\, point source catalog. The 3-band cryogenic data release contains data from $W1$, 2, and 3, and surveyed $30\%$ of the sky between August and October 2010. During the 3-band cryogenic survey, $W1$ and $W2$ operated with nearly the same sensitivity as during the full survey. Warming of the telescope reduced sensitivty in $W3$ and fully saturated $W4$. The NEOWISE post-cryogenic data release contains $W1$ and $W2$ measurements, with sensitivities close to those obtained during the full cryogenic phase. During this phase, \wise\, surveyed $70\%$ of the sky. Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR. The overall total number of sources compiled into \allwise\, totals over 747.6 million.

\subsection{Data Reduction}
We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise. We select only matches within 2" between our main sample of known \agb\, stars and \allwise. We further reduce our sample to unique matches to \allwise\, amongst \ogle, \macho, and \simbad\, by selecting \allwise\, photometry for unique \allwise\, designations. All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}. We require associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band.

We do not enforce the same faint or saturation limits on our contaminant samples. As a show of the strength of our selection criteria, we seek to include matches to each contaminant with reasonable detections in \allwise, and magnitudes reliable enough to perform the necessary color selections. We limit our sample from the \sdss\, DR9 SSPP to objects within 0.5 dex of the $g'-r'$ vs. $g'-i'$ stellar locus \citep{2014MNRAS.440.3430D}, \sdss\, SNR $>$ 5, [$W1/W2/W3/W4$] SNR $>$ 3,  detections in \twomass\, $J$ and $K_s$ bands, and one 2MASS association within 3". Objects from the NYU VAGC were reduced with requirements of SNR $>$ 3 for [$W1/W2/W3/W4$], required detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3". LRGs were required to have SNR $>$ 3 for [$W1/W2/W3$], detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3".  The restriction on $W4$ was dropped here, as it reduces the population of LRGs by an order of magnitude. The population for each sample from initial matching, as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.\\

\vspace{-10pt}
\begin{table}[h]
	\begin{center}
	\caption{AGB and Contaminant Populations}
	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
		\hline
		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
		\hline
		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
		\hline\hline
		Population & MACHO seq1 & seq2 & seq3 & seq4\\
		\hline
		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
		Reduced & 927 & 642 & 263 & 336 \\
		\hline\hline
		Population & OGLE C-rich & O-rich\\
		\hline
		2" match & 11,417 & 38,369 \\
		Reduced & 737 & 2515 \\
		\hline\hline
		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
		\hline
		2" match & 1,508,158 & 18,481 & 102,178 & 113,392 & 799,761 \\
		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
		\hline

		\label{tab:pop}
	\end{tabular}}
	\end{center}
\end{table}

### 2015-02-03 paper1/paper1_distribution.tex
updating for paper 1

\section{\agb\, Candidate Distribution}
\label{sec:distribution}
Put words here

### 2015-02-03 paper1/paper1_intro.tex
updating for paper 1

\section{Introduction}
The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars. This standard view was exemplified by the \cite{1980ApJS...44...73B} and \cite{1989ARA&A..27..555G} models, described in detail by, e.g., \cite{1993ARA&A..31..575M}. It was further motivated by observations of other galaxies, as well as what little information was available from \emph{High Precision Parallax Collecting Satellite} \citep[\emph{HIPPARCOS}, ][]{1984SSRv...39....1K} and smaller pencil-beam surveys. In these, the \mw\, is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.

The advent of the \emph{Sloan Digital Sky Survey} \citep[\sdss, ][]{2000AJ....120.1579Y} alleviated these limitations, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known. This new influx of data enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances. In turn, this led to the large scale ``tomography" of the \mw\; via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}. The resulting maps revealed rich, complex substructure in the distribution of the \mw's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V}, deeply shaking the existing view of the Galaxy.

In order to move forward from where \sdss\, tomography left off, we require observations that span an area larger than that of \sdss\, with Galactic objects that can be seen through interstellar dust out to large distances. Stars from the Asymptotic Giant Branch (\agb) are perfect candidates as probes to the \mw. \agb\, stars are represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H}, so they are bound to reside throughout the galaxy wherever other stars are present. During this phase, they produce substantial stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4} M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in SiO and amorphous carbon as they progress through being oxygen-rich to being carbon-rich. These winds collect into circumstellar shells that, when warmed by the stellar photosphere, shine brightly in the near- and mid-infrared (NIR \& MIR respectively).

The \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C} is a space-based observatory that has imaged the entire sky in the MIR (3.4, 4.6, 12, and 22$\mu$m). Additionally, \wise\, has been positionally matched to the Two-Micron All-Sky Survey (\twomass), a four-year mission characterizing the full sky in the NIR. Thus, the \wise\, catalog presents with hundreds of millions of sources with photometry of unprecedented sensitivity in the NIR and MIR--ideal for capturing \agb\, stars at Galactic distances. In Section~\ref{sec:data}, we describe in detail the data that we use for our study.

In Section~\ref{sec:criteria}, we describe the color-color criteria used to isolate \agb\, stars in the \wise\, dataset, and the color-magnitude relationships that were derived for these stars from the Large Magellanic Cloud and the \mw\, bulge.
In Section~\ref{sec:distribution}, we describe the spatial density distribution of \agb\, candidates in the Milky Way.
Our conclusions can be found in section~\ref{sec:conclusions}.

### 2015-02-03 thesispropbody_d2.tex
updating for paper 1

--- 2014-11-13 thesispropbody_d1.tex
@@
 This was due to the ability to map the Milky Way using the distances and positions of its own stellar residents.
-I propose to expand the mapping of the Milky Way, using the ubiquitous stars of the Asymptotic Giant Branch (hereon AGB) as probes to its spatial and chemical structure structure.
+I propose to expand the mapping of the Milky Way, using the ubiquitous stars of the Asymptotic Giant Branch (hereon AGB) as probes to its spatial and chemical structure.
 
@@
 %Fig 1
-%\vspace{-0.41in}
+%\vspace{-10pt}
 \begin{figure}[h]
+\vspace{-40pt}
 \begin{minipage}{0.9\textwidth}
@@
       \begin{minipage}[l]{3in}
-%      \hrule \smallskip
-%        {\footnotesize\baselineskip0.4cm{{\cap Figure 0:} {
-%        The variation 
-%        of the median photometric stellar metallicity in cylindrical Galactic coordinates 
-%        $R$ and $|Z|$  for $\sim$2.5 million stars from SDSS with $14.5 < r <20$,
-%        $0.2 < g-r < 0.4$ (blue turn-off stars), and photometric distance in the 0.8-9 kpc 
-%        range %\citep[details in][]{2008ApJ...684..287I}. 
-%        Each bin 
-%        (50 pc by 50 pc) contained in this map is colored according
-%        to the legend in the top left. The gradient of the median metallicity is essentially 
-%        parallel to the $|Z|$ axis, except in the region of the Monoceros stream, as marked. 
-%        The gray scale background is the best-fit model for the stellar number density distribution 
-%        from \cite{2008ApJ...673..864J}. The inset in the top right illustrates the extent of the data 
-%        volume relative to the rest of Galaxy, and the background image is the Andromeda 
-%        Galaxy. The three squares outline the regions used to measure the 3-dimensional
-%        velocity distribution for halo stars. The arrows illustrate the variation of the velocity
-%        ellipsoid orientation, which always points toward the Galactic center. Adapted from 
-%        \cite{2012ARA&A..50..251I}.
-%\smallskip \hrule}}}
+
      \caption{{\footnotesize
-        Variation of median photometric stellar metallicity in cylindrical Galactic coordinates 
+        Median photometric stellar metallicity in cylindrical Galactic coordinates 
         $R$ and $|Z|$  for $\sim$2.5 million stars from SDSS with $14.5 < r <20$,
-        $0.2 < g-r < 0.4$ (blue turn-off stars), and photometric distance in the 0.8-9 kpc 
+        $0.2 < g-r < 0.4$, and photometric distance in the 0.8-9 kpc 
         range. %\citep[details in][]{2008ApJ...684..287I}. 
 	The gradient of the median metallicity is essentially 
-        parallel to the $|Z|$ axis, except in the region of the Monoceros stream, as marked. 
+        parallel to the $|Z|$ axis, except in the Monoceros stream. The three squares outline 
+        the regions used to measure the 3-dimensional
+        velocity distribution for halo stars. The arrows illustrate the variation of the velocity
+        ellipsoid orientation, which always points toward the Galactic center.
         The gray scale background is the best-fit model for the stellar number density distribution 
-        from \cite{2008ApJ...673..864J}. The inset in the top right illustrates the extent of the data 
-        volume relative to the rest of Galaxy, and the background image is the Andromeda 
-        Galaxy. The three squares outline the regions used to measure the 3-dimensional
-        velocity distribution for halo stars. The arrows illustrate the variation of the velocity
-        ellipsoid orientation, which always points toward the Galactic center. Adapted from 
+         \citep{2008ApJ...673..864J}. \emph{Top right}: the extent of the data 
+        volume relative to the rest of Galaxy, superimposed on the Andromeda 
+        Galaxy.  Adapted from 
         \cite{2012ARA&A..50..251I}.
@@
 \end{minipage}
-\vspace{-10pt}
+\vspace{-40pt}
 \end{figure}
@@
 \label{sec:completedWork}
-The coverage of SDSS in the plane of the Galaxy, and even in the Galactic halo, is thus far limited to the regions targeted by the \emph{Sloan Extension for Galactic Understanding and Exploration} \citep[SEGUE, ][]{2012ApJ...757..166B}.
-More detailed studies of Galactic structure need to rely on the extensive observational capabilities of surveys of the entire sky, preferably with instruments that either display weak sensitivity to dust extinction along the line of sight or are immune.
-The all-sky survey conducted by the \emph{Wide-field Infrared Survey Explorer} \citep[WISE,][]{2012wise.rept....1C} gives us the advantage of being able to observe the whole sky in the infrared (IR) at 3.4, 4.6, 12, and 22$\mu$m with unprecedented depth and resolution.
-While the near- and mid-infrared (hereon NIR and MIR) are not immune to dust extinction along the line of sight, the influence of dust on observations is minimal as compared to optical wavelengths \citep{2012ApJ...757..166B}.
+The coverage of SDSS in the plane of the Galaxy, and even in the Galactic halo, is thus far limited to the regions targeted by the \emph{Sloan Extension for Galactic Understanding and Exploration} \citep[SEGUE;][]{2012ApJ...757..166B}.
+More detailed studies of Galactic structure need to rely on the extensive observational capabilities of surveys of the entire sky, using instruments with weak sensitivity to dust extinction.
+The all-sky survey conducted by the \emph{Wide-field Infrared Survey Explorer} \citep[WISE;][]{2010AJ....140.1868W, 2012wise.rept....1C} gives us the advantage of being able to observe the whole sky in the infrared (IR) at 3.4, 4.6, 12, and 22$\mu$m with unprecedented depth and resolution.
+The near- and mid-IR (hereon NIR and MIR) are not immune to dust extinction along the line of sight, but the influence of dust on observations is minimal as compared to optical wavelengths \citep{2012ApJ...757..166B, 2014MNRAS.440.3430D}.
 
-While observations in the NIR and MIR allow us to minimize the influence of dust extinction, and using WISE (particularly the latest data products released as ALLWISE) provides access to the entire sky, a proper survey of the Galaxy requires reliable objects that can be observed throughout the entirety of the Milky Way.
+While observations in the NIR and MIR allow us to minimize the influence of dust extinction, and  WISE (particularly the latest data products released as ALLWISE) provides access to the entire sky, a proper survey of the Galaxy requires reliable objects that can be observed throughout the entirety of the Milky Way.
 AGB stars are the perfect objects to serve as those probes of the Galaxy.
-AGB stars with dusty circumstellar shells shine brightly in the IR, particularly with the SiO present in the shells of O-rich stars and the AmC in shells of C-rich stars \citep{2011A&A...534A..79I}.
-They are bright enough to be detected by ALLWISE not only at the Galactic center (even with dust extinction), but beyond out to the edge of the Galactic disk \citep{2013RAA....13..323T}.
-They're also ubiquitous throughout the Galaxy, being the evolutionary products of low- and intermediate-mass Main Sequence stars \citep{1983ARA&A..21..271I}.
-They have massive stellar winds that enrich their local Galactic neighborhood, and their chemical content is representative of their local metallicity.
+AGB stars with dusty circumstellar shells shine brightly in the MIR, particularly with the SiO present in the shells of O-rich stars and the amorphous carbon in shells of C-rich stars \citep{2011A&A...534A..79I}.
+They are bright enough to be detected by ALLWISE not only near the Galactic center (even with dust extinction), but beyond out to the far edge of the Galactic disk \citep{2013RAA....13..323T}.
+They're also ubiquitous throughout the Galaxy, being the evolutionary products of 0.8-8 $M_\odot$ Main Sequence stars \citep{1983ARA&A..21..271I}.
 As such, our work here is first to obtain a clean, reliable sample of these stars within the Galaxy.
-Once obtained, we can then use the sample of ALLWISE AGB candidates in conjunction with high-confidence models of the Galaxy and of the consequences of AGB star evolution to describe the Milky Way in detail, parameterized by the physical characteristics of these AGB stars.
+Once obtained, we can then use the sample of ALLWISE AGB candidates in conjunction with high-confidence models of the Galaxy and of the consequences of AGB star evolution to describe the Milky Way in detail, parameterized by the physical characteristics of AGB stars.
 
 \subsection{Development of Initial Color Criteria for Selecting AGB Stars}
-We begin by selecting a known sample of AGB stars and matching to the ALLWISE-2MASS data release.
+We begin with a known sample of AGB stars, matching them to the ALLWISE-2MASS data release.
 The MACHO project \citep{2008AJ....136.1242F} and the Optical Gravitational Lens Experiment (OGLE) III Online Catalog of Variable Stars\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}} both contain catalogs of AGB stars, confirmed from their variability.
@@
 Thus we need a reliable sample of contaminants that we can exclude in color-color space to increase the likelihood that our candidates represent the objects we wish to study.
-We gather AGN, QSOs, and star forming galaxies from the Sloan Digital Sky Survey \citep{2000AJ....120.1579Y} data release (DR) 7, specifically the NYU Value-Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}, and the LRGs from the SDSS Luminous Red Galaxy Survey \citep{2010ApJ...710.1444K}.
+We gather spectroscopically classified AGN, QSOs, and star forming galaxies from the Sloan Digital Sky Survey \citep{2000AJ....120.1579Y} DR 7, specifically the NYU Value-Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}, and LRGs from the SDSS Luminous Red Galaxy Survey \citep{2010ApJ...710.1444K}.
 Data for stars in the SDSS stellar locus \citep{2014MNRAS.440.3430D} were drawn from the DR 9 SEGUE Stellar Parameters Pipeline (SSPP).
@@
 
-The color-color criteria set are fairly simple (Figure~\ref{fig:boundaries}): $J-K_s > 1.1$, $W2-W3 > 0.3$, and $W3-W4 < -0.83(W2-W3) + 3.37$.
-While we wish to maximize the completeness of our AGB star sample, we emphasize reliability of the sample and so minimize contamination below the 1\% level for each contaminant population and in total.
+The color-color criteria are fairly simple (Figure~\ref{fig:boundaries}): $J-K_s > 1.1$, $W2-W3 > 0.3$, and $W3-W4 < -0.83(W2-W3) + 3.37$.
+These criteria are set to minimize contamination to  $<1\%$ while retaining as many AGB stars as possible and requiring few cuts in color-color space.
 We note that in our selection criteria we exclude a substantial population of AGB stars with small or non-existent circumstellar shells.
@@
 
-As a first step toward creating a 3D map of our Galaxy using candidate AGB stars, we set out to find reliable color-magnitude relationships for our stars.
+As a first step toward creating a 3D map of our Galaxy, we derive reliable color-magnitude relationships for our AGB candidates.
 From our known sample, we select objects that fit within the narrow color-color criteria for O-rich and C-rich AGB stars set by \cite{2014MNRAS.442.3361N}.
-We then choose stars in the direction of regions of a) known distance and b) high star concentration in order to derive our color-magnitude relationships for these stars: the Milky Way bulge and the Large Magellanic Cloud (Figure~\ref{fig:galplot}).
+We then choose stars in the direction of regions of known distance and high stellar number density: the Milky Way bulge and the Large Magellanic Cloud (red \& blue respectively in Figure~\ref{fig:galplot}).
 For the LMC, we minimize foreground contamination by selecting stars with $b < -20^\circ$ and also meeting the following criteria:
@@
 \noindent With this selection of stars, we demonstrate that AGB stars possess an intrinsically-narrow magnitude distribution, as seen in \cite{1985A&A...152L...1H} and \cite{2002MNRAS.337..749J}. 
-We however do not use this sample further to define the color-magnitude relationships for AGB stars in the IR.
+%We however do not use this sample further to define the color-magnitude relationships for AGB stars in the IR.
 The LSST dust extinction code that we use to estimate the dust extinction along the line of sight (hereon {\tt GALFASTDUST}) doesn't include dust in or the shape of the LMC.
-We neglect to use the Small Magellanic Cloud entirely due to the low count of recovered objects from our initial color-color criteria.
+We neglect to use the Small Magellanic Cloud entirely due to the low object count.
 We instead solely use the sample in the direction of the Milky Way bulge for the necessary color-magnitude calibration (Figure~\ref{fig:colormag}).
@@
 
-Color-magnitude relations are then found for each IR color with respect to $M_{w1}$ as derived from the Milky Way bulge using linear regression.
-With $W1$ directly obtained from ALLWISE and $M_\text{w1}$ calculated with respect to the Milky Way bulge, we use {\tt GALFASTDUST} to simultaneously estimate distance moduli ($DM$) and extinction coefficients ($A_\lambda$) for each object.
+Color-magnitude relations are found for each IR color with respect to $M_{w1}$ as derived from the Milky Way bulge using linear regression.
+With $M_\text{w1}$ calculated with respect to the Milky Way bulge, we use {\tt GALFASTDUST} to simultaneously estimate distance moduli ($DM$) and extinction coefficients ($A_\lambda$) for each object.
 {\tt GALFASTDUST} returns the color excess $E(B-V)$ for a given ($l$, $b$) and distance $D$ in kpc by placing the object within the axisymmetric galaxy model used by \cite{2005AJ....130..659A}, normalizing $E(B-V)$ to the dust maps of \cite{1998ApJ...500..525S}, and interpolating the dust column along the line of sight.
 We convert $E(B-V)$ to $A_\lambda$ using the relations in the appendix of \cite{2014MNRAS.440.3430D}.
-These estimated $DM$ lie within a small degree of uncertainty ($\sim$0.5 mag), due to scatter about the initial color-magnitude relationship seen in Figure~\ref{fig:colormag}.
+These estimated $DM$ lie within a relatively small degree of uncertainty ($\sim$1 mag), due to scatter about the initial color-magnitude relationship seen in Figure~\ref{fig:colormag}.
 
@@
 %Create a 3D map of the Galaxy using the aforementioned distances.  Measure the disk scale height as a function of radius, and the disk scale length as a function of height (if possible).
-With distance information and position on the sky, we construct an X-Y-Z map of the Milky Way (Figure~\ref{fig:xyz_candidates}).
-We note that the overwhelming majority of AGB candidates in the Milky Way are O-rich, with the population C-rich AGB candidates of high photometric quality being essentially nonexistent, so we move forward solely with an analysis of candidates classified as such.
+With distance information and position on the sky, we construct an ($X,Y,Z $) map of the Milky Way (Figure~\ref{fig:xyz_candidates}).
+We note that the overwhelming majority of AGB candidates in the Milky Way are O-rich, with the Galactic population C-rich AGB candidates with high photometric quality being effectively nonexistent, so we move forward solely with an analysis O-rich candidates.
 We also note, as is clear in Figure~\ref{fig:xyz_candidates}, that this is by no means a complete picture of the Milky Way.
-The sample is heavily affected by dust extinction and source confusion within 0.5 kpc of the Galactic disk, and entirely beyond the bulge within $|b| < 1^\circ$ of the disk.
-Additionally, because we exclude sources that are beyond the saturation limits of WISE, we lose a large fraction of AGB candidates near the Sun (more toward the Galactic anticenter due to a lack of interstellar dust to attenuate the source's intrinsic brightness).
+The sample is still affected by dust extinction and source confusion, particularly within 0.5 kpc of the Galactic disk, and entirely beyond the bulge within $|b| < 1^\circ$ of the disk.
+Additionally, because we exclude sources that are beyond the saturation limits of ALLWISE, we lose a large fraction of AGB candidates near the Sun (more toward the Galactic anticenter due to a lack of interstellar dust).
 
@@
 Despite the loss of sources beyond the bulge and near the Sun, we are still able to make measurements of some common disk properties.
-We focus on the +X half of the Galaxy due to the higher completeness of sources, and exclude the wedge of the Milky Way containing that solar circle.
-We then examine the radial and vertical distribution of AGB candidates, and derive measurements of the scale height as a function of radius, as well as the scale length as a function of height (Figure~\ref{fig:horvert_profile}).
+We focus on the +X half of the Galaxy due to the higher completeness of sources, and exclude the wedge of the Milky Way containing saturation region.
+We then examine the radial and vertical distribution of AGB candidates, and derive measurements of the scale height as a function of $R$, as well as the scale length as a function of $\lvert Z\rvert$ (Figure~\ref{fig:horvert_profile}).
 
@@
 We can also investigate those objects not fitting the pre-defined O-rich and C-rich classification schemes.
+
 \subsection{Investigate Unclassified AGB Candidates}
@@
  However, we can explore wide regions of WISE color-color space and still eliminate the selected contaminants down to the $<$1\% level (Figure~\ref{fig:boundaries}).
-While this initial effort yields $\sim$215,000 objects qualifying as AGB candidates, the final sample used for the mapping the galaxy numbered only $\sim$69,000 objects.
-This begs the question, what were the other $\sim$150,000 objects, and how did they pass the AGB-trained color-color criteria without explicitly falling in the pre-defined O-rich and C-rich tracks?
+While this initial effort yields $\sim$215,000 objects qualifying as AGB candidates, the final sample used for mapping the galaxy contained only $\sim$69,000 objects.
+While many objects may have been lost due to saturation/reliability limits, this still begs the question: what were the other $\sim$150,000 objects, and how did they pass the AGB-trained color-color criteria without explicitly falling in the pre-defined O-rich and C-rich tracks?
 We seek to answer this by exploring other possible contaminants such as R Cor Borealis stars, young stellar objects, and planetary nebulae \citep{2002MNRAS.337..749J} through positional matches to the SIMBAD database, and verification of AGB candidacy with data from SDSS APOGEE (see section~\ref{sec:apogee}).
 
-\subsection{Comparison with {\tt TRILEGAL}}
+\subsection{Comparisons with {\tt TRILEGAL}}
 %Same as above, but use TRILEGAL and compare
@@
 {\tt GALFASTDUST} provided us with an estimation of the dust column along the line of sight to our candidate AGB stars, but with a set Galactic model.
-{\tt TRILEGAL}  allows us to test the validity of that model, alongside several other models of the Milky Way, with respect to our now-derived 3D distribution of AGB stars by providing stellar number densities along lines of sight.
+{\tt TRILEGAL}  allows us to test the validity of that model, alongside several other models of the Milky Way, with respect to our now-derived 3D distribution of AGB stars by providing stellar number densities along given lines of sight.
 
-With the simulated photometry output by {\tt TRILEGAL}, along with various choices of dust chemistry used as input for these simulations, we can use our color-color information as proxies for measurements of photospheric temperature and mass-loss rate, measurements unobtainable from the photometry of ALLWISE.
+With {\tt TRILEGAL}'s simulated photometry, along with various choices of dust chemistry used as input for these simulations, we can use our color-color information as proxies for measurements of photospheric temperature and mass-loss rate, measurements unobtainable from the photometry of ALLWISE.
 {\tt TRILEGAL} also has the added benefit of producing photometry for any given photometric set.
@@
 Correlate that with X-Y-Z position in the Galaxy (or maybe even just R-Z).
-The SDSS APOGEE survey observed $1.6\times10^5$ giant stars down to magnitude $H$ = 12.2 with IR spectroscopy \citep{2012ApJ...755L..25N}.
-We have matched the ALLWISE AGB candidate sample to the latest results from the APOGEE survey, retrieving $\sim$8,500 objects positionally matching our cleanest sample to within 1", with most positions o (Figure~\ref{fig:apogee_allwise}).
-This provides us with precise measurements of stellar parameters unable to be obtained from ALLWISE that we can now use to characterize our sample, provide measurable checks as to the reliability of our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
+The SDSS APOGEE survey observed $1.6\times10^5$ giant stars to $H < 12.2$ with NIR spectroscopy \citep{2012ApJ...755L..25N}.
+We have matched the ALLWISE AGB candidate sample to the latest results from the APOGEE survey, retrieving $\sim$8,500 objects positionally matching our cleanest sample to within 1", with most positions $0.4"$ apart (Figure~\ref{fig:apogee_allwise}).
+This provides us with precise measurements of stellar properties unable to be obtained from ALLWISE that we can now use to characterize our sample, provide measurable checks as to the reliability of our sample, as well as improve upon the existing capabilities and outputs of {\tt TRILEGAL}.
 
@@
 This work when completed will be just in time to set the stage for a few ambitious projects, and will be prime for building new and strengthening existing collaborative works.
+
 \subsection{From Candidates to Confirmations}
@@
 With accurate time-series photometry these stars can be confirmed without the need for spectroscopy across tens of thousands of sources.
-A prime candidate for this all-sky time-series data source is the European Space Agency's GAIA space observatory \citep{2001A&A...369..339P,2002MNRAS.331..649B}.
-With unprecedented astrometry and photometric sensitivity in the visual and NIR, as well as between 40 and 250 visits per object over 5 years down to $V$ = 20 mag, it will be trivial for GAIA to identify long period variables in our existing candidate sample \citep{2000ASPC..203...71E}.
+A prime candidate for such all-sky time-series data is the European Space Agency's Gaia space observatory \citep{2001A&A...369..339P,2002MNRAS.331..649B}.
+With unprecedented astrometry and photometric sensitivity in the optical and NIR, as well as between 40 and 250 visits per object over 5 years down to $V$ = 20 mag, it will be trivial for Gaia to identify long period variables in our existing candidate sample \citep{2000ASPC..203...71E}.
 

### 2015-03-31 paper1/paper1_criteria.tex
Updating paper

@@
 \label{sec:criteria}
-We made color-color criteria to create a catalog of AGB candidates. We kept it simple to maximize the number of objects we'd retain while minimizing contamination from non-AGB objects.
+We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
+
+The color-color criteria for AGB selection are as follows:
+\begin{align} 
+(J-K_s) > 1.1\label{eq:criteria1}\\
+(W2-W3) > 0.3\label{eq:criteria2}\\
+(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
+\end{align}
+The criteria in (1) and (2) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies).
+Criteria in (3) also reject stars from the stellar locus, but primarily function to reject IR-bright extragalactic sources.
+
+\noindent Sample completeness $\eta$ is defined as
+\begin{eqnarray*}
+\eta &=& \frac{N - n_\text{missed}}{N}
+\end{eqnarray*}
+where $N$ is the total number of objects in that sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries.
+Figure~\ref{fig:completeness} shows the distribution of completeness amongst AGB sources after the application of the above criteria.
+Sample contamination is typically defined as
+\begin{eqnarray*}
+\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
+\end{eqnarray*}
+where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected}$ is the total number of selected sources, including AGB objects and spurious objects \citep{2013sdmm.book.....I}.
+The contamination map is shown in Figure~\ref{fig:contamination}.
+
+\begin{figure}[h]
+\includegraphics[width=6.7in]{figs/completeness_map.pdf}
+\caption{Words \label{fig:completeness}.}
+\end{figure}
+\begin{figure}[h]
+\includegraphics[width=6.7in]{figs/contamination_map.pdf}
+\caption{Words \label{fig:contamination}.}
+\end{figure}
+
+A fairly large issue for creating the boundaries are the O-rich stars.
+Either the objects from the OGLE O-rich AGB star catalog were mostly mis-matched to WISE, taking on colors of the stellar locus, or the issue is more physical.
+It could be that the warm O-rich AGB photospheres are still visible and not heavily enshrouded in dust, thus appearing similar to Main Sequence stars in the NIR.
+Previous work on warm O-rich AGB stars has shown that their circumstellar shells are not prominent, and their NIR photometry reflects the Rayleigh Jeans tail of a cool 2000-4000K blackbody \citep{1974ApJ...189...89D}.
+
+The results of the applied critiera are summarized in Table~\ref{tab:completeness}. Of the objects remaining in the sample, 52.94\% of the AGBs have $\lvert b\rvert \le 10^\circ$.
+
+\begin{table}[h]
+    \begin{center}
+        \caption{Sample Completeness and Contamination}
+        \scalebox{0.85}{
+            \begin{tabular}{l c c c c c c}
+		\hline
+		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
+		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
+		\hline
+		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
+		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
+		\hline
+		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
+		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
+		\hline
+		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
+		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
+		\hline
+
+                \label{tab:completeness}
+            \end{tabular}}    
+    \end{center}
+\end{table}

### 2015-03-31 paper1/paper1_data.tex
Updating paper

@@
 We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise.
-We select only matches within 2" between our main sample of known \agb\, stars and \allwise.
-We further reduce our sample to unique matches to \allwise\, amongst \ogle, \macho, and \simbad\, by selecting \allwise\, photometry for unique \allwise\, designations.
-All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}.
-We require associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band.
-
-We do not enforce the same faint or saturation limits on our contaminant samples.
+We select only matches within 3" between our main sample of known \agb\, stars and \allwise.
+All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}, with no confusion or contamination flags in [$W1/W2/W3/W4$].
+We require only single associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band.
+%We do not enforce the same faint or saturation limits on our contaminant samples.
 As a show of the strength of our selection criteria, we seek to include matches to each contaminant with reasonable detections in \allwise, and magnitudes reliable enough to perform the necessary color selections.
@@
 The restriction on $W4$ was dropped here, as it reduces the population of LRGs by an order of magnitude.
-The population for each sample from initial matching, as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.\\
+The population for each sample from initial matching as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.
+The WISE color-color distributions for the AGB and contaminant samples are shown in Figure~\ref{fig:distros}.
+\\
 
 \vspace{-10pt}
+%\begin{table}[h]
+%	\begin{center}
+%	\caption{AGB and Contaminant Populations}
+%	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
+%		\hline
+%		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
+%		\hline
+%		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
+%		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
+%		\hline\hline
+%		Population & MACHO seq1 & seq2 & seq3 & seq4\\
+%		\hline
+%		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
+%		Reduced & 927 & 642 & 263 & 336 \\
+%		\hline\hline
+%		Population & OGLE C-rich & O-rich\\
+%		\hline
+%		2" match & 11,417 & 38,369 \\
+%		Reduced & 737 & 2515 \\
+%		\hline\hline
+%		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
+%		\hline
+%		2" match & 1,508,158 & 18,481 & 102,178 &  & 799,761 \\
+%		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
+%		\hline
+%
+%		\label{tab:pop}
+%	\end{tabular}}
+%	\end{center}
+%\end{table}
+
 \begin{table}[h]
@@
 	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
+		\hline\hline
+		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
 		\hline
-		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
+		3" match & 1,689 & 14,209 & 9,027 & 406 & 1,081 \\
+		Reduced & 684 & 1,782 & 3,241 & 43 & 511 \\ 
 		\hline
-		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
-		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
+		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
+		\hline
+		3" match & 5,279 & 3,519 & 2,619 & 3,070 \\
+		Reduced & 277 & 185 & 73 & 61 \\ 
+		\hline
+		Population & OGLE-III C-rich & O-rich \\ 
+		\hline
+		3" match & 11,542 & 38,848 \\
+		Reduced & 249 & 730 \\ 
+		\hline
+		Population & DR12 SSPP & DR7 LRG & QSO & AGN & Galaxies \\ 
+		\hline
+		3" match & 1,578,329 & 104,345 & 103,590 & 18,528 & 841,712 \\ 
+		Reduced & 67,508 & 84 & 3,977 & 1,069 & 44,314 \\ 
 		\hline\hline
-		Population & MACHO seq1 & seq2 & seq3 & seq4\\
-		\hline
-		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
-		Reduced & 927 & 642 & 263 & 336 \\
-		\hline\hline
-		Population & OGLE C-rich & O-rich\\
-		\hline
-		2" match & 11,417 & 38,369 \\
-		Reduced & 737 & 2515 \\
-		\hline\hline
-		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
-		\hline
-		2" match & 1,508,158 & 18,481 & 102,178 & 113,392 & 799,761 \\
-		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
-		\hline
 
@@
 \end{table}
+
+\begin{figure}
+\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
+\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
+\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
+\caption{\emph{Left:} Known AGB stars after data reduction. \emph{Right:} All contaminants after data reduction.\label{fig:distros}}
+\end{figure}

### 2015-04-01 paper1/paper1_criteria.tex
updating the paper

--- 2015-02-03 paper1/paper1_criteria.tex
@@
 \label{sec:criteria}
-We made color-color criteria to create a catalog of AGB candidates. We kept it simple to maximize the number of objects we'd retain while minimizing contamination from non-AGB objects.
+We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
+
+The color-color criteria for AGB selection are as follows:
+\begin{align} 
+(J-K_s) > 1.1\label{eq:criteria1}\\
+(W2-W3) > 0.3\label{eq:criteria2}\\
+(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
+\end{align}
+The criteria in (\ref{eq:criteria1}) and (\ref{eq:criteria2}) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies).
+(\ref{eq:criteria3}) also rejects stars from the stellar locus, but primarily functions to remove IR-bright extragalactic sources.
+
+\noindent Sample completeness $\eta$ is defined as
+\begin{eqnarray*}
+\eta &=& \frac{N - n_\text{missed}}{N}
+\end{eqnarray*}
+where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries.
+Figure~\ref{fig:completeness} shows the distribution of sample completeness amongst AGB sources after the application of the above criteria.
+The vast majority (79.07\%) of Galactic AGB stars is recovered after criteria (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}) are applied.
+The largest losses occur at the edges of the Galactic disk ($\lvert b\rvert\approx10^\circ$) and in regions of high stellar number density both in the Galaxy and the Magellanic Clouds.
+In the color-color diagrams, selection completeness degrades near the borders of the selction area, as objects that straddle these boundaries may exhibit enough emission in other color-color spaces to be removed by our criteria.
+Of the remaining sample of 5,709 objects, 52.94\% lie within the disk ($\lvert b\rvert<10^\circ$).
+
+\begin{figure}[h]
+\includegraphics[width=6.7in]{figs/completeness_map.pdf}
+\caption{Sample selection completeness maps in ($l,b$) space (\emph{top}) and the color-color space of our selection criteria (\emph{bottom}).
+Color scale shows the completeness fraction per bin, with 4-deg$^2$ bins on the Galactic map and 0.1 dex bins on each axis of the color-color diagrams.
+Selection criteria are shown as dashed lines.\label{fig:completeness}}
+\end{figure}
+
+\cite{2013sdmm.book.....I} defines sample contamination as
+\begin{eqnarray*}
+\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
+\end{eqnarray*}
+where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
+The contamination map is shown in Figure~\ref{fig:contamination}.
+Most bins in Figure~\ref{fig:contamination} exhibit 0\% contamination, with the overall contamination level at 0.38\%.
+What contaminants do remain exist primarily at the very fringes of the AGB star distribution in ($J-K_S$) vs ($W2 - W3$) space and near the redder boundary in ($W3-W4$) vs ($W2 - W3$), where bluer extragalactic sources creep into the selection region.
+The results of the applied critiera on both the collective AGB and contaminant samples are summarized in Table~\ref{tab:completeness}.
+
+\begin{figure}[h]
+\includegraphics[width=6.7in]{figs/contamination_map.pdf}
+\caption{Sample contamination in Galactic ($l,b$) space (\emph{top}) and color-color space (\emph{bottom}). Boundaries and binning the same as Fig~\ref{fig:completeness}.\label{fig:contamination}}
+\end{figure}
+
+%One large issue with the boundaries are the O-rich AGB stars.
+Either the objects from the OGLE O-rich AGB star catalog were mostly mis-matched to WISE, taking on colors of the stellar locus, or the issue is more physical.
+It could be that the warm O-rich AGB photospheres are still visible and not heavily enshrouded in dust, thus appearing similar to Main Sequence stars in the NIR.
+Previous work on warm O-rich AGB stars has shown that their circumstellar shells are not prominent, and their NIR photometry reflects the Rayleigh Jeans tail of a cool 2000-4000K blackbody \citep{1974ApJ...189...89D}.
+
+\begin{table}[h]
+    \begin{center}
+        \caption{Sample Selection Completeness and Contamination}
+        \scalebox{0.85}{
+            \begin{tabular}{l c c c c c c}
+		\hline
+		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
+		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
+		\hline
+		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
+		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
+		\hline
+		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
+		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
+		\hline
+		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
+		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
+		\hline
+
+                \label{tab:completeness}
+            \end{tabular}}    
+    \end{center}
+\end{table}

### 2015-04-01 paper1/paper1_data.tex
updating the paper

@@
 We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}.
+
 \ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
@@
 The resulting samples yielded 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich).
+
 From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars).
-The sample of AGB stars from \simbad\, was obtained by querying for all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
+Following \cite{2008AJ....136.1242F}, the objects are divided into sequences (seq) 1-4.
+Sequence 1 primarily contains Miras pulsating in their fundamental modes, whereas Sequences 2-4 contain semi-regular variables in various pulsation modes.
+
+The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
 The total sample of \agb\, stars is 100,393.
+Objects are classified spectroscopically, though by a variety of methods owing to the heterogeneous data housed within \simbad.
 
@@
 In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
-The initial \wise All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
+The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
 Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$].
@@
 \includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
-\caption{\emph{Left:} Known AGB stars after data reduction. \emph{Right:} All contaminants after data reduction.\label{fig:distros}}
+\caption{Logarithmic number densities for objects in \wise\, and \twomass\, color-color space, binned in 0.1 dex on each axis.
+\emph{Left:} The combined AGB sample matched to ALLWISE.
+\emph{Right:} The combined contaminant sample.\label{fig:distros}}
 \end{figure}

### 2015-05-05 paper1/paper1_criteria.tex
Update to paper 1

@@
 \label{sec:criteria}
-We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
+In creating color-color criteria to generate a catalog of AGB candidates, we seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
 
@@
 \end{align}
-The criteria in (\ref{eq:criteria1}) and (\ref{eq:criteria2}) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies).
+The criteria in (\ref{eq:criteria1}) and (\ref{eq:criteria2}) are primarily concerned with rejecting objects from the stellar locus, and other objects whose NIR spectra are dominated by the Rayleigh-Jeans tail.
 (\ref{eq:criteria3}) also rejects stars from the stellar locus, but primarily functions to remove IR-bright extragalactic sources.
+
+We also experiment with other criteria to focus more closely on the high-reliability AGB population, instead of trying to capture the most AGB-like objects. These criteria are as follows:
+\begin{align} 
+	(W1-W2) < 1\label{eq:criteria4}\\
+	(W2-W3) < 1\label{eq:criteria5}
+\end{align}
+Criteria (\ref{eq:critera4}) restricts NIR excesses. Criteria (\ref{eq:criteria5}) primarily removes extragalactic sources, while only slightly cutting into the distribution of known AGB stars.
 
@@
 The largest losses occur at the edges of the Galactic disk ($\lvert b\rvert\approx10^\circ$) and in regions of high stellar number density both in the Galaxy and the Magellanic Clouds.
-In the color-color diagrams, selection completeness degrades near the borders of the selction area, as objects that straddle these boundaries may exhibit enough emission in other color-color spaces to be removed by our criteria.
+In the color-color diagrams, selection completeness degrades near the borders of the selection area, as objects that straddle these boundaries may exhibit enough emission in other color-color spaces to be removed by our criteria.
 Of the remaining sample of 5,709 objects, 52.94\% lie within the disk ($\lvert b\rvert<10^\circ$).
+
+\begin{table}[h]
+    \begin{center}
+        \caption{Samples Recovered wih Application of Criteria (\%)}
+        \scalebox{0.85}{
+            \begin{tabular}{l r r r r r r r r r r r r}
+\hline
+Object Type & Reduced & (1) & (2) & (3) & (1,2) & (1,2,3) & | & \textbf{(4)} & \textbf{(5)} & \textbf{(1,2,4)} & \textbf{(1,2,5)} & \textbf{(1,2,4,5)}\\
+\hline
+All AGBs & 7220 & 95.15 & 84.89 & 95.21 & 82.62 & 79.07 & | & \textbf{94.22} & \textbf{88.82} & \textbf{77.05} & \textbf{72.71} & \textbf{71.14}\\ 
+O-rich AGB & 3147 & 98.98 & 100.00 & 100.00 & 98.98 & 98.98 & | & \textbf{98.44} & \textbf{97.33} & \textbf{97.43} & \textbf{96.31} & \textbf{94.88}\\ 
+C-rich AGB & 540 & 99.26 & 99.81 & 99.26 & 99.07 & 98.33 & | & \textbf{57.59} & \textbf{63.33} & \textbf{56.85} & \textbf{62.59} & \textbf{54.44}\\ 
+Unclassed AGB & 3533 & 91.11 & 69.15 & 90.32 & 65.53 & 58.39 & | & \textbf{96.07} & \textbf{85.14} & \textbf{61.99} & \textbf{53.24} & \textbf{52.53}\\ 
+DR12 SSPP & 67508 & 76.16 & 99.16 & 0.91 & 76.07 & 0.05 & | & \textbf{86.05} & \textbf{0.89} & \textbf{65.67} & \textbf{0.03} & \textbf{0.03}\\ 
+DR7 LRG & 84 & 96.43 & 100.00 & 0.00 & 96.43 & 0.00 & | & \textbf{89.29} & \textbf{0.00} & \textbf{85.71} & \textbf{0.00} & \textbf{0.00}\\ 
+QSO & 3783 & 73.54 & 99.97 & 0.13 & 73.54 & 0.11 & | & \textbf{34.50} & \textbf{0.11} & \textbf{27.54} & \textbf{0.08} & \textbf{0.08}\\ 
+Galaxy & 42066 & 78.55 & 100.00 & 0.01 & 78.54 & 0.00 & | & \textbf{96.39} & \textbf{0.01} & \textbf{75.31} & \textbf{0.00} & \textbf{0.00}\\ 
+AGN & 1011 & 79.33 & 100.00 & 0.00 & 79.33 & 0.00 & | & \textbf{96.34} & \textbf{0.00} & \textbf{75.87} & \textbf{0.00} & \textbf{0.00}\\
+
+\hline
+                \label{tab:criteria}
+            \end{tabular}}    
+    \end{center}
+\end{table}
 
@@
 What contaminants do remain exist primarily at the very fringes of the AGB star distribution in ($J-K_S$) vs ($W2 - W3$) space and near the redder boundary in ($W3-W4$) vs ($W2 - W3$), where bluer extragalactic sources creep into the selection region.
-The results of the applied critiera on both the collective AGB and contaminant samples are summarized in Table~\ref{tab:completeness}.
+The results of the applied criteria on both the collective AGB and contaminant samples are summarized in Table~\ref{tab:completeness}.
+
+We note that while we do achieve a low rate of contamination, most of our contaminant sources are out of the plane of the Galaxy, while many of our \agb\, sources are within the Galactic disk.
+Considering how our extragalactic contaminant sources are from various iterations of \sdss, we know that their spatial distributions should be effectively uniform across the sky {\color{red} [cite me]}.
+For contaminants unaffected by interstellar reddening, they would occupy the same regions of color-color space already marked by our existing contaminant sample, and would similarly be removed by (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}).
+This would produce the same zero-level contamination for all extragalactic sources aside from QSOs.
+We can adopt a QSO spatial density by dividing the total number of spectroscopically identified QSOs by the covered area on the sky.
+Using this, we recover a QSO spatial density of {\color{red} some number per square degree}.
+Using that density, and assuming ubiquity of sources on the sky, complete detection of QSOs {\color{red} down to some magnitude limit}, and a lack of interstellar reddening, we can estimate a maximum QSO contamination fraction of {\color{red} some number}.
+We use this number to report our QSO contamination fraction in Table~\ref{tab:completeness}, as QSOs significantly affected by interstellar reddening would exist further outside of the limits of (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}).
+A more rigorous study of QSO spatial density is outside the scope of this paper.
+
+The remaining contaminant sample is from the \sdss\, DR12 SSPP, with the number of remaining species found within shown in Figure~\ref{fig:sspp_histo}.
+{\color{red} the figures don't match up with each other.
+The number of red stars on the right side of the fig are significantly less than those in the histogram.
+Reconcile this.}
 
@@
 \caption{Sample contamination in Galactic ($l,b$) space (\emph{top}) and color-color space (\emph{bottom}). Boundaries and binning the same as Fig~\ref{fig:completeness}.\label{fig:contamination}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.7in]{figs/contaminating_sspp_objects.pdf}
+\includegraphics[width=6.7in]{figs/contaminating_sspp_objects_classes.pdf}
+\caption{Some histogram\label{fig:sspp_histo}}
 \end{figure}

### 2015-05-05 paper1/paper1_intro.tex
Update to paper 1

@@
 \section{Introduction}
-The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.
+%The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.
 This standard view was exemplified by the \cite{1980ApJS...44...73B} and \cite{1989ARA&A..27..555G} models, described in detail by, e.g., \cite{1993ARA&A..31..575M}.
 It was further motivated by observations of other galaxies, as well as what little information was available from \emph{High Precision Parallax Collecting Satellite} \citep[\emph{HIPPARCOS}, ][]{1984SSRv...39....1K} and smaller pencil-beam surveys.
-In these, the \mw\, is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
-
-The advent of the \emph{Sloan Digital Sky Survey} \citep[\sdss, ][]{2000AJ....120.1579Y} alleviated these limitations, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
-This new influx of data enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
-In turn, this led to the large scale ``tomography" of the \mw\; via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
+In these,
+The structure of the \mw\, has for many years been uncertain, with historical models assuming three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo \citep{1980ApJS...44...73B, 1989ARA&A..27..555G,1993ARA&A..31..575M}.
+The advent of the \emph{Sloan Digital Sky Survey} \citep[\sdss, ][]{2000AJ....120.1579Y} has since provided more detail, using accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
+%This new influx of data enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
+These new data led to the large scale ``tomography" of the \mw\; via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
 The resulting maps revealed rich, complex substructure in the distribution of the \mw's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V}, deeply shaking the existing view of the Galaxy.
 
-In order to move forward from where \sdss\, tomography left off, we require observations that span an area larger than that of \sdss\, with Galactic objects that can be seen through interstellar dust out to large distances.
-Stars from the Asymptotic Giant Branch (\agb) are perfect candidates as probes to the \mw.
-\agb\, stars are represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H}, so they are bound to reside throughout the galaxy wherever other stars are present.
-During this phase, they produce substantial stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4} M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in SiO and amorphous carbon as they progress through being oxygen-rich to being carbon-rich.
-These winds collect into circumstellar shells that, when warmed by the stellar photosphere, shine brightly in the near- and mid-infrared (NIR \& MIR respectively).
+In order to move forward from where \sdss\, tomography left off, we require observations that span an area larger than that of \sdss\,, with Galactic objects that can be seen through interstellar dust out to large distances.
+Stars from the Asymptotic Giant Branch (\agb) are perfect candidates to serve as these probes to the \mw.
+\agb\, stars represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H} -- the mass-range with the highest number of stars as inferred from the  \cite{2001MNRAS.322..231K} initial mass function that can also reach the final stages of stellar evolution within a galactic timescale.
+Because of this, they are bound to reside throughout the galaxy wherever other stars are present.
+This phase of evolution is marked by two distinct periods: the early \agb\, (E-\agb)  and the thermally-pulsing \agb\, phases.
+During the latter of the two, \agb\, stars produce substantial dust-driven stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4}$ $M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in oxides (SiO, Al$_2$O$_3$, etc.) and carbon-rich molecules (SiC, AmC, etc.), with the chemical dominance being highly dependent upon the metallicity of the host galaxy \citep{2005A&A...434..691M}.
+Galaxies such as the Milky Way are expected to have a substantial population of oxygen-rich AGB stars \citep{1985A&A...152L...1H}, whereas low-metallicity galaxies such as the Magellanic clouds have been shown to possess AGB populations dominated by carbon-rich stars \citep{2011AJ....142..103B}.
+In both cases, the other species of AGB star is rarely seen, as richness in one chemical type (e.g.
+oxides)  necessitates the almost complete capture of the other chemical type (e.g.
+carbon) in CO \citep{1983ARA&A..21..271I}.
+Over time, these chemically-rich winds create vast circumstellar shells that, when warmed by the stellar photosphere, emit in the near- and mid-infrared (NIR \& MIR respectively).
+Although the molecular species present in oxygen- and carbon-rich winds are vastly different, emission from both types produce strong IR emission near 10$\mu$m, visible out to the Magellanic clouds \citep{2011A&A...534A..79I} and beyond.
+Thus, if these stars can be pinpointed by a survey with a wide area of observation and high positional accuracy, they can be used as markers for a map of the Milky Way \citep{2013RAA....13..323T}.
 
-The \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C} is a space-based observatory that has imaged the entire sky in the MIR (3.4, 4.6, 12, and 22$\mu$m).
+Such a survey can be found in the \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C}, a space-based observatory that has imaged the entire sky in the MIR (3.4, 4.6, 12, and 22$\mu$m).
 Additionally, \wise\, has been positionally matched to the Two-Micron All-Sky Survey (\twomass), a four-year mission characterizing the full sky in the NIR.
-Thus, the \wise\, catalog presents with hundreds of millions of sources with photometry of unprecedented sensitivity in the NIR and MIR--ideal for capturing \agb\, stars at Galactic distances.
+Thus, the \wise\, catalog presents with hundreds of millions of sources with photometry of unprecedented sensitivity and positional accuracy in the NIR and MIR--ideal for capturing \agb\, stars at Galactic distances.
+In this paper, we use samples of known Galactic and Magellanic \agb\, stars to formulate color-color criteria with \wise\, and \twomass\, photometry that can produce a reliable catalog of IR \agb\, candidates.
+We then use these candidates in conjunction with estimates of Galactic dust extinction along the line of sight to produce a Galaxy-wide number density distribution of \agb\, stars.
+
 In Section~\ref{sec:data}, we describe in detail the data that we use for our study.
-
 In Section~\ref{sec:criteria}, we describe the color-color criteria used to isolate \agb\, stars in the \wise\, dataset, and the color-magnitude relationships that were derived for these stars from the Large Magellanic Cloud and the \mw\, bulge.

### 2015-05-05 paper1/paper1_intro.tex
ZI edits

--- 2015-02-03 paper1/paper1_intro.tex
@@
 \section{Introduction}
+
+- brief intro to Gal structure
+
+- SDSS and other recent results do not go into the plane
+
+- mention a few studies that do (GLIMPSE etc) and ongoing data taking with DECam
+
+- AGB stars: luminous in IR, relatively easy to find; mention MACHO and OGLE studies, 
+   refer to Jackson et al. IRAS-based study, lead to WISE
+
+- brief description of WISE
+
+- paper outline
+
 The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.
@@
 In these, the \mw\, is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
+
+Review is \citet{IBJ2012} \citep{IBJ2012}
 

### 2015-05-05 paper1/paper1_data.tex

@@
 \label{sec:data}
+
 \subsection{Data Sources}
-In order to analyze the physical distribution of \agb\, stars in the IR, we require reliable identifications of known \agb, stars to generate color-color selection criteria.
-We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}.
-
-\ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
-O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in \cite{2009AcA....59..239S,2011AcA....61..217S}.
-Data reduction techniques are described in \cite{2008AcA....58...69U}.
-The resulting samples yielded 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich).
-
-From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars).
-Following \cite{2008AJ....136.1242F}, the objects are divided into sequences (seq) 1-4.
-Sequence 1 primarily contains Miras pulsating in their fundamental modes, whereas Sequences 2-4 contain semi-regular variables in various pulsation modes.
-
-The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
-The total sample of \agb\, stars is 100,393.
-Objects are classified spectroscopically, though by a variety of methods owing to the heterogeneous data housed within \simbad.
-
-Since \sdss\, is an optical survey of unprecedented extent and sensitivity, we use \sdss\, spectroscopic catalogs to select contaminant sources for the NIR-MIR color-color fields of \agb\, stars.
-Data for active galactic nuclei (AGN; 19,184 objects), quasi-stellar objects (QSOs; 122,550 objects), and star forming/burst galaxies (820,272 objects total) were drawn from \sdss\, DR7, specifically from the NYU Value Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}.
-Luminous Red Galaxies (LRGs) were selected from the SDSS Luminous Red Galaxy Survey \citep[105,631 objects, ][]{2010ApJ...710.1444K}.
-Data for stars in the SDSS stellar locus were drawn from the DR 9 SEGUE Stellar Parameters Pipeline (SSPP) \citep[1,843,190 objects, ][]{2012ApJS..203...21A}.
-
+\label{sec:sources}
 In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
@@
 During the 3-band cryogenic survey, $W1$ and $W2$ operated with nearly the same sensitivity as during the full survey.
-Warming of the telescope reduced sensitivty in $W3$ and fully saturated $W4$.
+Warming of the telescope reduced sensitivity in $W3$ and fully saturated $W4$.
 The NEOWISE post-cryogenic data release contains $W1$ and $W2$ measurements, with sensitivities close to those obtained during the full cryogenic phase.
@@
 Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR.
-The overall total number of sources compiled into \allwise\, totals over 747.6 million.
+The overall number of sources compiled into \allwise\, totals over 747.6 million.
+
+In order to generate a reliable, high-confidence catalog of Galactic candidate \agb\, stars, we must first define color-color criteria from known \agb\, star samples.
+We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}.
+
+\ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
+O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in detail in \cite{2009AcA....59..239S,2011AcA....61..217S}.
+{\color{red}[Describe the selection bit in a little more detail, along with their sample completeness, selection biases, and contamination fractions]} Data reduction techniques are described in \cite{2008AcA....58...69U}.
+The resulting samples yield 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich).
+
+From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars).
+{\color{red}Why were these objects selected?
+Howe were they selected?
+What is their completeness, selection bias, and contamination fraction?} Following \cite{2008AJ....136.1242F}, the objects are divided into sequences (seq) 1-4.
+Sequence 1 primarily contains Mira variables pulsating in their fundamental modes, whereas Sequences 2-4 contain semi-regular variables in various pulsation modes.
+
+The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
+Objects are classified spectroscopically, though by a variety of methods owing to the heterogeneous data housed within \simbad.
+Together with \macho\, and \ogle, the total sample of \agb\, stars is 100,393.
+Because there is a high likelihood that samples between \ogle, \macho, and \simbad overlap, we retain only unique objects after the initial data reduction in section~\ref{sec:reduction}.
+
+We use \sdss\, spectroscopic catalogs to find and quantify regions in NIR-MIR color-color space populated by plausible contaminant sources.
+These include any Galactic stellar objects and planetary nebulae, as well as a host of extragalactic sources.
+Data for active galactic nuclei (AGN; 19,184 objects), quasi-stellar objects (QSOs; 122,550 objects), and star forming/burst galaxies (820,272 objects total) were drawn from \sdss\, DR7, specifically from the NYU Value Added Galaxy Catalog\footnote{\url{http://sdss.physics.nyu.edu/vagc/}} \citep[VAGC]{2005AJ....129.2562B}.
+Luminous Red Galaxies (LRGs) were selected from the SDSS Luminous Red Galaxy Survey \citep[105,631 objects, ][]{2010ApJ...710.1444K}.
+Data for stars in the SDSS stellar locus were drawn from the DR 9 SEGUE Stellar Parameters Pipeline (SSPP) \citep[1,843,190 objects, ][]{2012ApJS..203...21A}.
+{\color{red} Include bit about YSOs and PNe from SIMBAD}
 
 \subsection{Data Reduction}
+\label{sec:reduction}
 We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise.
-We select only matches within 3" between our main sample of known \agb\, stars and \allwise.
-All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}, with no confusion or contamination flags in [$W1/W2/W3/W4$].
-We require only single associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band.
-%We do not enforce the same faint or saturation limits on our contaminant samples.
-As a show of the strength of our selection criteria, we seek to include matches to each contaminant with reasonable detections in \allwise, and magnitudes reliable enough to perform the necessary color selections.
-We limit our sample from the \sdss\, DR9 SSPP to objects within 0.5 dex of the $g'-r'$ vs.
-$g'-i'$ stellar locus \citep{2014MNRAS.440.3430D}, \sdss\, SNR $>$ 5, [$W1/W2/W3/W4$] SNR $>$ 3,  detections in \twomass\, $J$ and $K_s$ bands, and one 2MASS association within 3".
-Objects from the NYU VAGC were reduced with requirements of SNR $>$ 3 for [$W1/W2/W3/W4$], required detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3".
-LRGs were required to have SNR $>$ 3 for [$W1/W2/W3$], detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3".
-The restriction on $W4$ was dropped here, as it reduces the population of LRGs by an order of magnitude.
+We select only matches within 3" between  each sample and \allwise.
+All samples of \agb\, were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}, with no flags for confusion or contamination as a spurious source in any band.
+We also require only single associations with \twomass\, objects within 3", detections in [$J/K_S/W1/W2/W3/W4$], and SNR $>$ 3 in each \allwise\, band.
+
 The population for each sample from initial matching as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.
@@
 	\begin{center}
-	\caption{AGB and Contaminant Populations}
+
 	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
@@
 
-		\label{tab:pop}
 	\end{tabular}}    
+	\caption{\agb\, and contaminant populations matched to WISE before and after sample reduction in section~\ref{sec:reduction}.
+MACHO sequences (seq1-seq4) are from \cite{2008AJ....136.1242F} and described briefly in section~\ref{sec:sources}.\label{tab:pop}}
 	\end{center}
+
 \end{table}

### 2015-05-05 paper1/paper1_intro.tex

--- 2015-05-05 paper1/paper1_intro.tex
@@
 \section{Introduction}
+
+- brief intro to Gal structure
+
+- SDSS and other recent results do not go into the plane
+
+- mention a few studies that do (GLIMPSE etc) and ongoing data taking with DECam
+
+- AGB stars: luminous in IR, relatively easy to find; mention MACHO and OGLE studies, 
+   refer to Jackson et al. IRAS-based study, lead to WISE
+
+- brief description of WISE
+
+- paper outline
+
+The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.
+This standard view was exemplified by the \cite{1980ApJS...44...73B} and \cite{1989ARA&A..27..555G} models, described in detail by, e.g., \cite{1993ARA&A..31..575M}.
+It was further motivated by observations of other galaxies, as well as what little information was available from \emph{High Precision Parallax Collecting Satellite} \citep[\emph{HIPPARCOS}, ][]{1984SSRv...39....1K} and smaller pencil-beam surveys.
+In these, the \mw\, is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
+
+Review is \citet{IBJ2012} \citep{IBJ2012}
+
+The advent of the \emph{Sloan Digital Sky Survey} \citep[\sdss, ][]{2000AJ....120.1579Y} alleviated these limitations, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
+This new influx of data enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
+In turn, this led to the large scale ``tomography" of the \mw\; via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
+The resulting maps revealed rich, complex substructure in the distribution of the \mw's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V}, deeply shaking the existing view of the Galaxy.
+
+In order to move forward from where \sdss\, tomography left off, we require observations that span an area larger than that of \sdss\, with Galactic objects that can be seen through interstellar dust out to large distances.
+Stars from the Asymptotic Giant Branch (\agb) are perfect candidates as probes to the \mw.
+\agb\, stars are represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H}, so they are bound to reside throughout the galaxy wherever other stars are present.
+During this phase, they produce substantial stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4} M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in SiO and amorphous carbon as they progress through being oxygen-rich to being carbon-rich.
+These winds collect into circumstellar shells that, when warmed by the stellar photosphere, shine brightly in the near- and mid-infrared (NIR \& MIR respectively).
+=======
 %The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.

### 2015-05-05 paper1/paper1_data.tex
Something

@@
 \ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
-O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in detail in \cite{2009AcA....59..239S,2011AcA....61..217S}.
+LPVs were classified into 3 categories: \ogle\, Small Amplitude Red Giants (OSARGs), Semi-Regular Variables (SRVs), and Miras.
+All AGB stars  O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in detail in \cite{2009AcA....59..239S,2011AcA....61..217S}.
 {\color{red}[Describe the selection bit in a little more detail, along with their sample completeness, selection biases, and contamination fractions]} Data reduction techniques are described in \cite{2008AcA....58...69U}.

### 2015-05-05 paper1/paper1_data.tex
ZI edits May 5

@@
-\section{Data Sources and Reduction}
+\section{Input Catalogs and Data Preparation}
 \label{sec:data}
 
-\subsection{Data Sources}
+Our principal data source is the merged WISE-2MASS catalog. In order to derive
+WISE-2MASS color-based selection criteria, we use several training samples
+described below: a heterogeneous sample of bright \agb\ stars listed in 
+SIMBAD database and more homogneous fainter samples selected using 
+optical time-domain surveys OGLE-III and MACHO. The latter samples 
+include \agb\ stars from the LMC and SMC and are also used to 
+calibrate color-absolute magnitude relations that underlie our distance
+estimates. In order to assess what infrared populations could contribute to
+contamination of selected \agb\ candidates, we utilize a number of SDSS
+catalogs, also described below. We conclude this section by describing
+how these catalogs auxiliary catalogs were  positionally merged with the 
+WISE-2MASS catalog.
+
+Now rearrange the old text below in these subsections:
+
+\subsection{WISE-2MASS catalog}
+
+\subsection{AGB stars from SIMBAD database}
+
+\subsection{AGB stars from optical time-domain surveys}
+A few sentences of introduction...
+\subsubsection{MACHO}
+\subsubsection{OGLE-III}
+
+\subsection{Imaging and spectroscopic catalogs from SDSS}
+A few sentences of introduction...
+\subsubsection{Galactic sources} 
+\subsubsection{Extragalactic populations}
+
+\subsection{Merged sample}
+
+How was it merged, which further cuts and why...
+
+\subsection{Completeness and contamination of the initial \agb\ sample}
+
+In the next section you are using the \agb\ sample from this section to 
+estimate completeness and contamination of color-selected samples. Here
+you need to summarize what do you believe completeness and contamination
+are for your initial sample constructed using SIMBAD and optical surveys.
+In particular, it is important to describe expected distance limits.
+
+\subsection{XXX old stuff from here to the end of this section: Data Sources}
 \label{sec:sources}

### 2015-05-05 paper1/paper1_intro.tex
ZI edits May 5

--- 2015-05-05 paper1/paper1_intro.tex
@@
 
-- brief intro to Gal structure
+The structure of the \mw\ holds clues to the processes of formation and evolution of galaxies.
+Historical models typically assumed three discrete components described by simple analytic 
+expressions: the thin disk, thick disk, and halo \citep{1980ApJS...44...73B, 1989ARA&A..27..555G,1993ARA&A..31..575M}. 
+Recent surveys, such as the Sloan Digital Sky Survey \citep[\sdss, ][]{2000AJ....120.1579Y} 
+and the Two Micron All Sky Survey \citep[\twomass, ][]{2006AJ....131.1163S}, have provided 
+much more detail about these components. For example, SDSS data have constrained stellar distributions 
+in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
+The resulting maps revealed rich, complex substructure in the distribution of the \mw's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V}, deeply shaking the older view of a smooth Galaxy.
 
-- SDSS and other recent results do not go into the plane
+The large distance limit of about 100 kpc for some halo populations (e.g. RR Lyrae and BHB stars 
+with SDSS, XXX Sesar 2010 ref and red giants with 2MASS, XXX Majewski ref), as well as about 10 kpc 
+for very numerous main sequence stars, is not reached when studying disk component at low galactic
+latitudes. As shown, for example, by Berry et al., the extinction due to interstellar dust limits 
+the SDSS sample of main sequence stars to heliocentric distances of only a few kpc. In order
+to avoid detrimental effect of dust extinction, stellar samples need to be probed at longer infrared
+wavelengths. Several infrared surveys covering the Galactic plane have recently become available
+(gives refs for WISE and GLIMPSE, VVV: arXiv:1111.5511).
 
-- mention a few studies that do (GLIMPSE etc) and ongoing data taking with DECam
+Among populations suitable for studying Galactic structure using infrared surveys, Asymptotic Giant 
+Branch (\agb) stars stand out.
+\agb\ stars represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H}.
+Stars from this mass range can reach the final stages of stellar evolution within the Galactic timescale ($\sim$10 Gyr, \citealt{1983ARA&A..21..271I}) and thus are bound to reside throughout the Galaxy wherever other stars are present.
+This phase of stellar evolution is marked by two distinct
+episodes with different observational characteristics: the early \agb\, phase (E-\agb)  and the thermally-pulsing \agb\, phase (TP).
+During the thermally-pulsing phase, \agb\, stars produce substantial dust-driven stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4}$ $M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in oxides (SiO, Al$_2$O$_3$, etc.) or carbon-rich molecules (SiC, AmC, etc.) The dominant chemical composition is highly dependent upon the metallicity of the host galaxy \citep{2005A&A...434..691M}.
+High-metallicity galaxies like  the Milky Way have a substantial population of oxygen-rich AGB stars \citep{1985A&A...152L...1H}, whereas low-metallicity galaxies such as the Magellanic clouds are dominated by carbon-rich \agb\ stars \citep{2011AJ....142..103B}.
+In both cases, the other species of AGB star is rarely seen, as richness in one chemical type (e.g.
+oxides)  necessitates the almost complete capture of the other chemical type (e.g.
+carbon) in CO \citep{1983ARA&A..21..271I}.
 
-- AGB stars: luminous in IR, relatively easy to find; mention MACHO and OGLE studies, 
-   refer to Jackson et al. IRAS-based study, lead to WISE
+The dust-rich winds in TP phase  create vast circumstellar shells that are warmed by the stellar photosphere and emit
+copiously in the near- and mid-infrared (NIR \& MIR, respectively). Together with high bolometric luminosity
+(10$^3$--10$^4$ \Mo), this redistribution of the output radiation to the infrared wavelength makes \agb\ stars
+excellent disk probe when infrared survey data are available. Indeed, they were detected all the way to the 
+Galactic center even with the IRAS survey \citep{2002MNRAS.337..749J}. Such disk studies can now be significantly
+improved thanks to the much more sensitive WISE survey.
 
-- brief description of WISE
+The \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C}, is a space observatory that has imaged essentially the entire sky in the MIR (3.4, 4.6, 12, and 22 $\mu$m).
+\wise\, catalog has been positionally matched to the \twomass\ catalog, with the matched catalog listing NIR and MIR 7-band photometry for
+hundreds of millions of sources. Given the depths of the two surveys, this catalog should contain AGB stars to many
+kpc beyond the Galactic center. Here we develop selection methods for \agb\ stars using WISE and 2MASS data,
+and analyse the resulting samples.
 
-- paper outline
-
-The formation of galaxies like the \mw\, was long thought to be a steady process that created a smooth distribution of stars.
-This standard view was exemplified by the \cite{1980ApJS...44...73B} and \cite{1989ARA&A..27..555G} models, described in detail by, e.g., \cite{1993ARA&A..31..575M}.
-It was further motivated by observations of other galaxies, as well as what little information was available from \emph{High Precision Parallax Collecting Satellite} \citep[\emph{HIPPARCOS}, ][]{1984SSRv...39....1K} and smaller pencil-beam surveys.
-In these, the \mw\, is modeled by three discrete components described by simple analytic expressions: the thin disk, thick disk, and halo.
-
-Review is \citet{IBJ2012} \citep{IBJ2012}
-
-The advent of the \emph{Sloan Digital Sky Survey} \citep[\sdss, ][]{2000AJ....120.1579Y} alleviated these limitations, providing accurate digital multi-band optical photometry across a quarter of the sky, as well as the largest optical spectroscopic catalog thus far known.
-This new influx of data enabled the development and application of photometric parallax methods, using color-magnitude relations to estimate stellar distances.
-In turn, this led to the large scale ``tomography" of the \mw\; via stellar distributions in the 7-dimensional space spanned by spatial coordinates \citep{2008ApJ...673..864J}, velocity components \citep{2010ApJ...716....1B}, and metallicity \citep{2008ApJ...684..287I}.
-The resulting maps revealed rich, complex substructure in the distribution of the \mw's stars \citep[e.g.][]{2000AJ....120..963I,2000ApJ...540..825Y,2001ApJ...554L..33V,2002ApJ...569..245N,2003ApJ...599.1082M,2006ApJ...642L.137B,2006ApJ...651L..29G,2006AJ....132..714V}, deeply shaking the existing view of the Galaxy.
-
-In order to move forward from where \sdss\, tomography left off, we require observations that span an area larger than that of \sdss\, with Galactic objects that can be seen through interstellar dust out to large distances.
-Stars from the Asymptotic Giant Branch (\agb) are perfect candidates as probes to the \mw.
-\agb\, stars are represent the last stage of evolution for stars between 0.8 and 8 $M_\odot$ \citep{1983ARA&A..21..271I, 2005ARA&A..43..435H}, so they are bound to reside throughout the galaxy wherever other stars are present.
-During this phase, they produce substantial stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4} M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in SiO and amorphous carbon as they progress through being oxygen-rich to being carbon-rich.
-These winds collect into circumstellar shells that, when warmed by the stellar photosphere, shine brightly in the near- and mid-infrared (NIR \& MIR respectively).
-
-The \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C} is a space-based observatory that has imaged the entire sky in the MIR (3.4, 4.6, 12, and 22$\mu$m).
-Additionally, \wise\, has been positionally matched to the Two-Micron All-Sky Survey (\twomass), a four-year mission characterizing the full sky in the NIR.
-Thus, the \wise\, catalog presents with hundreds of millions of sources with photometry of unprecedented sensitivity in the NIR and MIR--ideal for capturing \agb\, stars at Galactic distances.
-In Section~\ref{sec:data}, we describe in detail the data that we use for our study.
-
-In Section~\ref{sec:criteria}, we describe the color-color criteria used to isolate \agb\, stars in the \wise\, dataset, and the color-magnitude relationships that were derived for these stars from the Large Magellanic Cloud and the \mw\, bulge.
-In Section~\ref{sec:distribution}, we describe the spatial density distribution of \agb\, candidates in the Milky Way.
-Our conclusions can be found in section~\ref{sec:conclusions}.
+In Section~\ref{sec:data}, we describe in detail the WISE, 2MASS and other auxiliary data used in our study. 
+In Section~\ref{sec:criteria}, we use samples of known Galactic and Magellanic \agb\, stars to derive 
+WISE-2MASS color-based selection criteria, and calibrate color-absolute magnitude relations. 
+In Section~\ref{sec:distribution}, we describe the spatial density distribution of selected \agb\, candidates
+from the Milky Way. Our conclusions are summarized in Section~\ref{sec:conclusions}.

### 2015-06-22 paper1/paper1_data.tex
Edited sections 2.2, 2.3, and 2.5

@@
 
-Now rearrange the old text below in these subsections:
-
 \subsection{WISE-2MASS catalog}
+Write a few sentences about the WISE survey...
 
 \subsection{AGB stars from SIMBAD database}
+The SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} is comprised of basic data, classifications, and measurements for all types of astronomical objects from a variety of surveys.
+Classifications are hierarchical, and are classified with increasing specificity based on the available data, though each star can have multiple classifications.
+AGB stars have subcategories of C-stars ([C/O] $>$ 1) and S-stars ([C/O] $\sim$ 1).
+Where subclassification is unclear, the main classification remains as AGB star.
+OH/IR stars are subsets of peculiar stars that specifically show strong OH maser emission and strong IR emission, identifying them clearly as AGB stars.
+Mira-type variables are a subset of variable stars, and are classified based upon their periodicity.
+Semi-regular variables are not included as they contain an unknown mixture of AGB stars and red supergiants.
+The sample of AGB stars from SIMBAD was obtained by querying all objects with types classified as C-stars (23,628), S-stars (1,514), AGB stars (4,459), OH/IR stars (1,265), and Mira-type variables (10,828).
+Without duplicate objects between the five types, there is a total of 39,348 AGB stars from SIMBAD.
 
 \subsection{AGB stars from optical time-domain surveys}
-A few sentences of introduction...
+Many of the identifications in the SIMBAD catalog are from spectroscopic studies, however many more AGB stars have been found through their characteristic optical periodicity.
+The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo} (MACHO) survey have are two time-domain surveys that have captured and catalogued such variability.
+
 \subsubsection{MACHO}
+
 \subsubsection{OGLE-III}
@@
 \subsection{Merged sample}
+Each AGB sample as well as the SDSS sources were matched separately to the ALLWISE data products via the {\tt GATOR} tool at the NASA/IPAC Infrared Science Archive\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}}.
+The accepted matching radius was 3".
+The matched sample was further reduced by enforcing the WISE 5$\sigma$ faint limits of [$W1/W2/W3/W4$] $<$ [$16.83/15.6/11.32/8.0$], saturation limits of [$W1/W2/W3/W4$] $>$ [2.0/1.5/-3.0/-4.0], signal-to-noise ratios greater than 3 in each WISE band.
+We constrain the possibility for a mismatched extragalactic IR source by asserting that the extended source flag $\le$ 1, corresponding either to a point source (flag = 0) or a source whose profile-fit photometry is goodness-of-fit is $>$ 3 in any WISE band but is not coincident with any extended sources from the 2MASS survey.
+We also require only one match to a 2MASS source within 3".
 
-How was it merged, which further cuts and why...
+{\color{red}{How was it merged, which further cuts and why...}}
 

### 2015-07-06 paper1/paper1_data.tex
Updating paper with new figures

@@
 how these catalogs auxiliary catalogs were  positionally merged with the 
-WISE-2MASS catalog.
+WISE-2MASS catalog and summarize the object counts before and after quality control in Table 1 {\color{red} [reference for table 1 when made]}.
 
@@
 \subsection{AGB stars from SIMBAD database}
-The SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} is comprised of basic data, classifications, and measurements for all types of astronomical objects from a variety of surveys.
+The SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} is comprised of astrometry, classifications, and photometry for all types of astronomical objects from a variety of surveys.
 Classifications are hierarchical, and are classified with increasing specificity based on the available data, though each star can have multiple classifications.
@@
 Many of the identifications in the SIMBAD catalog are from spectroscopic studies, however many more AGB stars have been found through their characteristic optical periodicity.
-The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo} (MACHO) survey have are two time-domain surveys that have captured and catalogued such variability.
+The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo Objects} (MACHO) survey are two time-domain microlensing surveys that have captured and catalogued such variability.
+In Sections~\ref{sec:macho} and \ref{sec:ogle} we disc.uss the origins of the data from each source and how they were then used for this study
 
-\subsubsection{MACHO}
+\subsubsection{MACHO}\label{sec:macho}
+The \emph{MACHO} project {\color{red} [cite Alcock et al.
+1992 and 1997]} was a two-color optical microlensing survey searching for massive compact halo objects in the Milky Way bulge and the Magellanic Clouds.
+As a consequence of the time-domain survey, the \emph{MACHO} catalog contains 8 years of observations for several million stars, many of which were observed to be variable.
+%The {\tt VIZIER}\footnote{\url{http://vizier.u-strasbg.fr/viz-bin/VizieR?-source=II\%2F247}} service contains variable stars in the Magellanic Clouds, of which many are expected to be AGB stars. 
+Using light curves of these variable stars, {\color{red}[cite Fraser 2008]} identified a population of {\color{red}[some number]} LPVs in the Magellanic Clouds and Miky Way bulge, with observations in Kron-Cousins $V$ and $R$ with matches to objects from 2MASS.
+{\color{red}[some small tidbit about the reduction procedure]}.
 
-\subsubsection{OGLE-III}
+\subsubsection{OGLE-III}\label{sec:ogle}
+The \emph{OGLE-III} Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall \emph{OGLE-III} experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars in  $V$- and $I$-magnitudes.
+These observations saturated at $I = 12.5$ mag, and are limited at the faint end for $I = 20.5$ mag {\color{red}[cite Zebrun, Soszynski, and Wozniak 2001]}.
+The \emph{OGLE-III} CVS contains a database of LPVs classified on object type (\emph{OGLE} Small Amplitude Red Giant [OSARG], Mira, Semi-regular Variable star [SRV]), evolutionary status (AGB, RGB), and spectral type (O-rich, C-rich), with observations focused on the Magellanic Clouds.
+There are also obserations of the Milky Way bulge, however those objects are not yet accessible through the \emph{OGLE-III} CVS site\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}.
+Object type is based on variability, with OSARGs having $10<P<100$ days and $I$-band amplitudes $0.005 < A < 0.13$ mag {\color{red} [cite Soszynski et al 2004a]}.
+Miras and SRVs were separated using $I$-band amplitude and position in period-NIR magnitude space {\color{red} [cite Soszynski et al 2005]}.
+The evolutionary status was determined via color-magnitude diagrams, where AGBs are separated by being above the TRGB in color-NIR magnitude space {\color{red}[cite Kiss \& Bedding 2003]}.
+The spectral type is based on divisions in color-color space and period-luminosity space, where O-rich and C-rich stars show clear separation (Figure~\ref{fig:oglepmag}).
+From OGLE we have 52,976 objects.
+
+\begin{figure}[h]
+\includegraphics[width=3in]{figs/ogle_2mass_pmag.png}
+\includegraphics[width=3in]{figs/ogle_2mass_colcol.png}
+\caption{Words \label{fig:oglepmag}}
+\end{figure}
 

### 2015-07-16 paper1/paper1_data.tex
Updated OGLE section 2.3.2

--- 2015-06-22 paper1/paper1_data.tex
@@
 Many of the identifications in the SIMBAD catalog are from spectroscopic studies, however many more AGB stars have been found through their characteristic optical periodicity.
-The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo} (MACHO) survey have are two time-domain surveys that have captured and catalogued such variability.
+The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo Objects} (MACHO) survey are two time-domain microlensing surveys that have captured and catalogued such variability.
+In Sections~\ref{sec:macho} and \ref{sec:ogle} we disc.uss the origins of the data from each source and how they were then used for this study
 
-\subsubsection{MACHO}
+\subsubsection{MACHO}\label{sec:macho}
+The \emph{MACHO} project \citep{1992ASPC...34..193A,1997ApJ...482...89A} was a two-color optical microlensing survey searching for massive compact halo objects in the Milky Way bulge and the Magellanic Clouds.
+As a consequence of the time-domain survey, the \emph{MACHO} catalog contains 8 years of observations for several million stars, many of which were observed to be variable.
+Using Fourier decomposition, \cite{2008PhDT.........6F} identified and classified a population of 46,832 LPVs in the LMC, 3,943 in the SMC {\color{red}[check this]} , and {\color{red}[some other number]} in the Milky Way Bulge with matches to 2MASS.
+We note that due to high scatter and uncertainty of Galactic objects in period-magnitude space, we only take sequences 1 and 2 (Mira-type variables) for the Milky Way Bulge sample.
+Each object carries observations in Kron-Cousins $V$ \& $R$, and 2MASS photometry.
+AGB stars are separated from RGBs by identifying the tip of the RGB ($K_s < 12.3$ in the LMC and $K_s < 12.8$ in the SMC).
+{\color{red}[some small tidbit about the reduction procedure]}.
 
-\subsubsection{OGLE-III}
+\subsubsection{OGLE-III}\label{sec:ogle}
+The \emph{OGLE-III} Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall \emph{OGLE-III} experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars.
+With high-precision photometry \citep[$~0.001$ mag][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
+
+The \emph{OGLE-III} CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with LPVs classified by object type, evolutionary status, and spectral type.
+Object type is based chiefly on variability and luminosity.
+\emph{OGLE} Small Amplitude Red Giants (hereon OSARGs) are weakly-variable ($0.005 < A_I < 0.13$ mag), with relatively short periods \citep[$10 < P < 100$ days;][]{2004AcA....54..129S}.
+Additionally, because these objects are multiperiodic, OSARGs are selected using their period ratios and their distances from established period-luminosity (log $_{10} P-L$) sequences from \cite{1999IAUS..191..151W}.
+For more detail in their selection, see \cite{2007AcA....57..201S}.
+Mira variables and Semi-Regular Variables (hereon SRVs) are identified using $I$-band amplitudes and $P-W_I$, where $W_I$ is the reddening-free Wesenheit index \citep{2005AcA....55..331S}:
+\begin{eqnarray}
+W_I &=& I - 1.55(V-I)
+\end{eqnarray}
+
+As red giant branch stars (hereon RGBs) can contaminate the same log$_{10} P-L$ space as AGB stars, evolutionary type is distinguished in two ways.
+Stars above the tip of the RGBÑ[$[I/K_s] = [14.56/12.05]$ mag in LMC \citep{2004AcA....54..129S, 2007AcA....57..201S}Ñare AGBs by virtue of luminosity.
+Below that threshold, a narrow sequence of stars appears to share the same slope and intercept with AGB stars above the TRGB.
+Additionally they share similar primary-secondary period ratios as known AGB stars \citep{2005AcA....55..331S}.
+However, as there is still some significant contamination by RGBs, we neglect all OGLE objects below the TRGB moving forward.
+
+For individual log$_{10} P-L$ sequences, spectral type (O-rich or C-rich) is easily seen as separations in log$_{10} P-L$ space as well as visible-NIR color-color space.
+The initial classifications are based on spectroscopically-observed stars and were extended more generally to log$_{10} P-L$ and visible-NIR color-color criteria \citep{2005AcA....55..331S, 2007AcA....57..201S}.
+This clear separation can be seen in Figure~\ref{fig:oglepmag}.
+
+From the initial \emph{OGLE-III} sample of 52,976 LPVs, we retain 49,547 after matching to ALLWISE within 3" and ensuring only 1 match to \emph{2MASS}.
+The $I$-band median of \emph{OGLE-III} LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag.
+Out of the entire sample only 13 objects (Miras) fall beyond the 20.5 mag faint limit, and 41 beyond the saturation limit.
+As such, we expect this to effectively represent the complete sample of oscillating AGB stars in the LMC.
+
+\begin{figure}[h]
+\includegraphics[width=3in]{figs/ogle_2mass_period_mag_threeplot.png}
+\includegraphics[width=3in]{figs/ogle_2mass_color_color_threeplot.png}
+\caption{O-rich (\emph{blue}) and C-rich (\emph{red}) OGLE-III LPVs matched to 2MASS. \emph{Left}: $W_I$ vs log period in days. \emph{Right:} OGLE-III $(V-I)$ vs 2MASS $(J-K_s)$. \label{fig:oglepmag}}
+\end{figure}
 

### 2015-07-28 paper1/paper1_data.tex
Changes made up to and including section 2.4

@@
 Our principal data source is the merged WISE-2MASS catalog. In order to derive
-WISE-2MASS color-based selection criteria, we use several training samples
-described below: a heterogeneous sample of bright \agb\ stars listed in 
-SIMBAD database and more homogneous fainter samples selected using 
-optical time-domain surveys OGLE-III and MACHO. The latter samples 
-include \agb\ stars from the LMC and SMC and are also used to 
-calibrate color-absolute magnitude relations that underlie our distance
-estimates. In order to assess what infrared populations could contribute to
-contamination of selected \agb\ candidates, we utilize a number of SDSS
-catalogs, also described below. We conclude this section by describing
-how these catalogs auxiliary catalogs were  positionally merged with the 
-WISE-2MASS catalog and summarize the object counts before and after quality control in Table 1 {\color{red} [reference for table 1 when made]}.
+WISE-2MASS color-based selection criteria, we use %several training samples
+%described below: a heterogeneous sample of bright \agb\ stars listed in 
+%SIMBAD database and more homogneous fainter samples selected using 
+%optical time-domain surveys OGLE-III and MACHO. The latter samples 
+%include \agb\ stars from the LMC and SMC and are also used to 
+%calibrate color-absolute magnitude relations that underlie our distance
+%estimates. In order to assess what infrared populations could contribute to
+%contamination of selected \agb\ candidates, we utilize a number of SDSS
+%catalogs, also described below. We conclude this section by describing
+%how these catalogs auxiliary catalogs were  positionally merged with the 
+%WISE-2MASS catalog and summarize the object counts before and after quality control in Table 1 
+a homogenous sample of AGB stars from the OGLE-III Catalog of Variable Stars supplying objects in the Magellanic Clouds. This sample serves to calibrate color-color and color-absolute magnitude
+relations that underlie our distance estimates. In order to assess what infrared populations
+could contribute to contamination of selected \agb\ candidates, we utilize
+extragalactic sources from SDSS data release 7 pulled from the NYU Value-added Galaxy Catalog,
+the SDSS unofficial Luminous Red Galaxy sample,
+and young stellar objects obtained identified in \wise\ data, also described below. 
+We conclude this section by describing how these auxiliary catalogs were positionally merged
+with the WISE-2MASS catalog, and summarize the object counts before and after quality control in 
+Table 1 {\color{red} [reference for table 1 when made]}.
 
 \subsection{WISE-2MASS catalog}
-Write a few sentences about the WISE survey...
+In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
+The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
+Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$].
+The positions of objects in the \wise\, catalog were calibrated to the \twomass\, point source catalog.
 
-\subsection{AGB stars from SIMBAD database}
-The SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} is comprised of astrometry, classifications, and photometry for all types of astronomical objects from a variety of surveys.
+The 3-band cryogenic data release contains data from $W1$, 2, and 3, and surveyed $30\%$ of the sky between August and October 2010.
+During the 3-band cryogenic survey, $W1$ and $W2$ operated with nearly the same sensitivity as during the full survey.
+Warming of the telescope reduced sensitivity in $W3$ and fully saturated $W4$.
+The NEOWISE post-cryogenic data release contains $W1$ and $W2$ measurements, with sensitivities close to those obtained during the full cryogenic phase.
+During this phase, \wise\, surveyed $70\%$ of the sky.
+Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR.
+The overall number of sources compiled into \allwise\, totals over 747.6 million.
+
+%\subsection{AGB stars from SIMBAD database}
+%The SIMBAD database\footnote{\url{http://simbad.u-strasbg.fr/simbad/}} is comprised of astrometry, classifications, and photometry for all types of astronomical objects from a variety of surveys.
 Classifications are hierarchical, and are classified with increasing specificity based on the available data, though each star can have multiple classifications.
@@
 
-\subsection{AGB stars from optical time-domain surveys}
-Many of the identifications in the SIMBAD catalog are from spectroscopic studies, however many more AGB stars have been found through their characteristic optical periodicity.
+%\subsection{AGB stars from optical time-domain surveys}
+%Many of the identifications in the SIMBAD catalog are from spectroscopic studies, however many more AGB stars have been found through their characteristic optical periodicity.
 The \emph{Optical Gravitational Lens Experiment} (OGLE) and the \emph{MAssive Compacy Halo Objects} (MACHO) survey are two time-domain microlensing surveys that have captured and catalogued such variability.
@@
 
-\subsubsection{MACHO}\label{sec:macho}
-The \emph{MACHO} project \citep{1992ASPC...34..193A,1997ApJ...482...89A} was a two-color optical microlensing survey searching for massive compact halo objects in the Milky Way bulge and the Magellanic Clouds.
+%\subsubsection{MACHO}\label{sec:macho}
+%The \emph{MACHO} project \citep{1992ASPC...34..193A,1997ApJ...482...89A} was a two-color optical microlensing survey searching for massive compact halo objects in the Milky Way bulge and the Magellanic Clouds.
 As a consequence of the time-domain survey, the \emph{MACHO} catalog contains 8 years of observations for several million stars, many of which were observed to be variable.
-Using Fourier decomposition, \cite{2008PhDT.........6F} identified and classified a population of 46,832 LPVs in the LMC, 3,943 in the SMC {\color{red}[check this]} , and {\color{red}[some other number]} in the Milky Way Bulge with matches to 2MASS.
+%Using Fourier decomposition, \cite{2008PhDT.........6F} identified and classified a population of 46,832 LPVs in the LMC, 3,943 in the SMC {\color{red}[check this]} , and {\color{red}[some other number]} in the Milky Way Bulge with matches to 2MASS.
 We note that due to high scatter and uncertainty of Galactic objects in period-magnitude space, we only take sequences 1 and 2 (Mira-type variables) for the Milky Way Bulge sample.
@@
 
-\subsubsection{OGLE-III}\label{sec:ogle}
+\subsection{AGB stars from the OGLE-III Catalog of Variable Stars}\label{sec:ogle}
 The \emph{OGLE-III} Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall \emph{OGLE-III} experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars.
@@
 
-\subsection{Imaging and spectroscopic catalogs from SDSS}
-A few sentences of introduction...
-\subsubsection{Galactic sources} 
-\subsubsection{Extragalactic populations}
+\subsection{Extragalactic catalogs from SDSS}
+SDSS is a multi-year survey of roughly 25\% of the sky collecting both photometric (\emph{u, g, r, i, z}-bands for 700+ million objects) and spectroscopic measurements (1.6+ million) of stars and extragalactic sources \citep{2000AJ....120.1579Y}.
+Though its ground-based nature and sensitivity to interstellar extinction prevent SDSS from being an all-sky survey, it represents the largest spectrophotometric catalog to date.
+For this study we collect extragalactic objects from the NYU Value-added Galaxy Catalog, containing objects matched between SDSS and several auxiliary surveys spanning multiple wavelength regimes, and the unofficial SDSS Luminous Red Galaxy sample, isolating high-confidence luminous red galaxies from the SDSS database.
+
+\subsubsection{The NYU Value-added Galaxy Catalog}
+In order to cull the full set of SDSS spectrophotometry and produce an easily-referenced extragalactic catalog for investigating galaxy formation and evolution, the NYU Value-added Galaxy Catalog \citep[NYU-VAGC;][]{2005AJ....129.2562B} was created as of SDSS data-release 7 (DR7).
+
+The NYU-VAGC contains matches between SDSS spectrophotometry and the following catalogs: the FIRST radio survey, the 2MASS Point Source Catalog, the 2MASS Extended Source Catalog, the Two-degree Field Galaxy Redshift Survey, the IRAS Point Source Catalog Redshift Survey, and Reference Catalog 3 (RC3.9b).
+As a consequence, the NYU-VAGC retains SDSS optical photometry in addition to spectral classifications (QSO, galaxy, star) and subclassifications (AGN, starforming galaxy, M3 star, etc.) for 2,506,741 objects across 10,417 square degrees of the SDSS footprint.
+Main and secondary classifications are made by comparing measured optical spectra to the SDSS spectral library.
+See \cite{2005AJ....129.2562B} for the full set of reduction and inclusion criteria.
+This catalog is available via an online data repository\footnote{\url{http://sdss.physics.nyu.edu/vagc/}}, along with a full description of the data and other subsamples not used in this study.
+
+\subsubsection{The Unofficial Luminous Red Galaxy Sample}
+Subselected for the study of Baryon Acoustic Oscillations by \cite{2010ApJ...710.1444K}, the Luminous Red Galaxy sample \citep[LRGs;][]{2001AJ....122.2267E} is also sourced from SDSS DR7.
+LRGs contaminate the color-color space of AGB stars due to the severe redshift of their intrinsically-bright optical fluxes.
+We obtain the unofficial LRG sample from the data repository\footnote{\url{http://cosmo.nyu.edu/~eak306/SDSS-LRG.html}} of \cite{2010ApJ...710.1444K}.
+The initial sample is volume-limited, containing 105,631 objects spanning a redshift range of $0.16 < z < 0.47$.
+
+\subsection{WISE+2MASS Young Stellar Objects}
+Young Stellar Objects (YSOs) represent a unique contaminant in our search for dust-enshrouded AGB stars.
+Similar to AGBs, YSOs are luminous sources surrounded by warm, dusty environments.
+That warm dust can then glow brightly in the MIR, easily contaminating AGB color-color space.
+To characterize and later eliminate this potential contaminant we select as our base sample 290 YSOs from \cite{2011ApJS..196....4R}, a survey searching for YSOs in the Taurus Molecular Cloud using data from WISE and ancillary data from SDSS and 2MASS.
+As YSO tend to be embedded within high-extinction clouds of this sort, we expect that the dust-reddened NIR/MIR emission from these objects should match to similar objects in the LMC when we later calibrate our color-color criteria.
 

### 2015-07-28 paper1/paper1_intro.tex
Edited sections 2.2, 2.3, and 2.5
Changes made up to and including section 2.4

@@
 The large distance limit of about 100 kpc for some halo populations (e.g. RR Lyrae and BHB stars 
-with SDSS, XXX Sesar 2010 ref and red giants with 2MASS, XXX Majewski ref), as well as about 10 kpc 
+with SDSS, \citealt{2010ApJ...708..717S} and red giants with 2MASS, \citealt{2003ApJ...599.1082M}), as well as about 10 kpc 
 for very numerous main sequence stars, is not reached when studying disk component at low galactic
@@
 to avoid detrimental effect of dust extinction, stellar samples need to be probed at longer infrared
-wavelengths. Several infrared surveys covering the Galactic plane have recently become available
-(gives refs for WISE and GLIMPSE, VVV: arXiv:1111.5511).
+wavelengths.
+Several infrared surveys covering the Galactic plane have recently become available (WISE, \citealt{2010AJ....140.1868W, 2012wise.rept....1C}, GLIMPSE \citealt{2009PASP..121..213C, 2003PASP..115..953B}, VVV \citealt{2012A&A...537A.107S}).
 
@@
 
-The \emph{Wide-field Infrared Survey Explorer} \citep[\wise, ][]{2010AJ....140.1868W, 2012wise.rept....1C}, is a space observatory that has imaged essentially the entire sky in the MIR (3.4, 4.6, 12, and 22 $\mu$m).
-\wise\, catalog has been positionally matched to the \twomass\ catalog, with the matched catalog listing NIR and MIR 7-band photometry for
+The \emph{Wide-field Infrared Survey Explorer}, \wise, is a space observatory that has imaged essentially the entire sky in the MIR (3.4, 4.6, 12, and 22 $\mu$m).
+The \wise\, catalog has been positionally matched to the \twomass\ catalog, with the matched catalog listing NIR and MIR 7-band photometry for
 hundreds of millions of sources. Given the depths of the two surveys, this catalog should contain AGB stars to many
@@
 
-In Section~\ref{sec:data}, we describe in detail the WISE, 2MASS and other auxiliary data used in our study. 
+In Section~\ref{sec:data}, we describe in detail the WISE, 2MASS and other auxiliary data used in our study and the data reduction process. 
 In Section~\ref{sec:criteria}, we use samples of known Galactic and Magellanic \agb\, stars to derive 

### 2015-07-30 paper1/paper1_data.tex
Updates!

@@
 As YSO tend to be embedded within high-extinction clouds of this sort, we expect that the dust-reddened NIR/MIR emission from these objects should match to similar objects in the LMC when we later calibrate our color-color criteria.
+
+\subsection{WISE+2MASS Stellar Locus}
+We begin to extract the stellar locus from ALLWISE by centering on the LMC ($276.5^\circ < l < 284^\circ$, $-38.2^\circ < b < -28^\circ$) by necessitating that every object have 1 2MASS association, [$W1/W2/W3$] signal-to-noise $>$ 1, detections in 2MASS J, H, and K,  and confusion \& contamination flags set to 0 for [$W1/W2/W3$].
+We then adapt the color-color criteria from \cite{2014MNRAS.440.3430D}.
+Their locus focuses on stars with effective temperatures $3540 < T_\text{eff} < 7200 K$.
+We fit their criteria for $J-K_s$ vs $W1 - W2$ within 3$\sigma$ of the locus with degree-3 polynomials.
+The resulting fit for the 3$\sigma$ stellar locus follows:
+
+\begin{eqnarray}
+(J - K_s) & < & 61.67(W1-W2)^3 - 17.88(W1-W2)^2 + 1.68(W1-W2) + 0.99\\
+(J - K_s) & > & 40.19(W1-W2)^3 - 9.78(W1-W2)^2 + 0.54(W1-W2) + 0.64
+\end{eqnarray}
+
+The resulting locus sample, as well as the original \cite{2014MNRAS.440.3430D} bounds and locus are showin in Fig~\ref{fig:locus}. Note that our sample goes beyond the limits of the stated locus.
+
+\begin{figure}[h]
+\includegraphics[width=3in]{figs/lmc_locus_3sig.png}
+\caption{The WISE-2MASS stellar locus. \label{fig:locus}}
+\end{figure}
 

### 2015-08-06 paper1/paper1_data.tex
Updating paper with section 2.6

@@
 with the WISE-2MASS catalog, and summarize the object counts before and after quality control in 
-Table 1 {\color{red} [reference for table 1 when made]}.
+Tables~\ref{tab:reductions} \& 2.
 
@@
 \subsection{AGB stars from the OGLE-III Catalog of Variable Stars}\label{sec:ogle}
-The \emph{OGLE-III} Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall \emph{OGLE-III} experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars.
+The OGLE-III Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall OGLE-III experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars.
 With high-precision photometry \citep[$~0.001$ mag][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
 
-The \emph{OGLE-III} CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with LPVs classified by object type, evolutionary status, and spectral type.
+The OGLE-III CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with LPVs classified by object type, evolutionary status, and spectral type.
 Object type is based chiefly on variability and luminosity.
-\emph{OGLE} Small Amplitude Red Giants (hereon OSARGs) are weakly-variable ($0.005 < A_I < 0.13$ mag), with relatively short periods \citep[$10 < P < 100$ days;][]{2004AcA....54..129S}.
+OGLE Small Amplitude Red Giants (hereon OSARGs) are weakly-variable ($0.005 < A_I < 0.13$ mag), with relatively short periods \citep[$10 < P < 100$ days;][]{2004AcA....54..129S}.
 Additionally, because these objects are multiperiodic, OSARGs are selected using their period ratios and their distances from established period-luminosity (log $_{10} P-L$) sequences from \cite{1999IAUS..191..151W}.
@@
 
-From the initial \emph{OGLE-III} sample of 52,976 LPVs, we retain 49,547 after matching to ALLWISE within 3" and ensuring only 1 match to \emph{2MASS}.
-The $I$-band median of \emph{OGLE-III} LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag.
+From the initial OGLE-III sample of 52,976 LPVs, we retain 43,209 after matching to ALLWISE within 3", and ensuring only 1 match to 2MASS as well as all objects needing to be in the LMC.
+The $I$-band median of OGLE-III LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag.
 Out of the entire sample only 13 objects (Miras) fall beyond the 20.5 mag faint limit, and 41 beyond the saturation limit.
@@
 The NYU-VAGC contains matches between SDSS spectrophotometry and the following catalogs: the FIRST radio survey, the 2MASS Point Source Catalog, the 2MASS Extended Source Catalog, the Two-degree Field Galaxy Redshift Survey, the IRAS Point Source Catalog Redshift Survey, and Reference Catalog 3 (RC3.9b).
-As a consequence, the NYU-VAGC retains SDSS optical photometry in addition to spectral classifications (QSO, galaxy, star) and subclassifications (AGN, starforming galaxy, M3 star, etc.) for 2,506,741 objects across 10,417 square degrees of the SDSS footprint.
+As a consequence, the NYU-VAGC retains SDSS optical photometry in addition to spectral classifications (QSO, galaxy, star) and subclassifications (AGN, starforming galaxy, starburst galaxy, etc.) for 441,707 objects across 10,417 square degrees of the SDSS footprint.
 Main and secondary classifications are made by comparing measured optical spectra to the SDSS spectral library.
@@
 This catalog is available via an online data repository\footnote{\url{http://sdss.physics.nyu.edu/vagc/}}, along with a full description of the data and other subsamples not used in this study.
+The populations for each species (QSO, AGN, starforming galaxy (SF), starburst galaxy (SB)) are found in Table~\ref{tab:reductions}.
 
@@
 That warm dust can then glow brightly in the MIR, easily contaminating AGB color-color space.
-To characterize and later eliminate this potential contaminant we select as our base sample 290 YSOs from \cite{2011ApJS..196....4R}, a survey searching for YSOs in the Taurus Molecular Cloud using data from WISE and ancillary data from SDSS and 2MASS.
+To characterize and later eliminate this potential contaminant we select as our base sample 290 YSOs from \cite{2011ApJS..196....4R}, a survey searching for YSOs in the Taurus Molecular Cloud using data from WISE, ancillary data from SDSS and 2MASS, and covering {\color{red} some number} deg$^2$.
 As YSO tend to be embedded within high-extinction clouds of this sort, we expect that the dust-reddened NIR/MIR emission from these objects should match to similar objects in the LMC when we later calibrate our color-color criteria.
@@
 \includegraphics[width=3in]{figs/lmc_locus_3sig.png}
-\caption{The WISE-2MASS stellar locus. \label{fig:locus}}
+\caption{The WISE-2MASS stellar locus.
+(\emph{black dots}) objects from the stellar locus in the direction of the LMC.
+(\emph{blue line}) WISE-2MASS color-color stellar locus derived from \cite{2014MNRAS.440.3430D}.
+(\emph{red line}) 3$\sigma$ boundaries for WISE-2MASS color-color locus.
+\label{fig:locus}}
 \end{figure}
 
-\subsection{Merged sample}
-Each AGB sample as well as the SDSS sources were matched separately to the ALLWISE data products via the {\tt GATOR} tool at the NASA/IPAC Infrared Science Archive\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}}.
-The accepted matching radius was 3".
-The matched sample was further reduced by enforcing the WISE 5$\sigma$ faint limits of [$W1/W2/W3/W4$] $<$ [$16.83/15.6/11.32/8.0$], saturation limits of [$W1/W2/W3/W4$] $>$ [2.0/1.5/-3.0/-4.0], signal-to-noise ratios greater than 3 in each WISE band.
-We constrain the possibility for a mismatched extragalactic IR source by asserting that the extended source flag $\le$ 1, corresponding either to a point source (flag = 0) or a source whose profile-fit photometry is goodness-of-fit is $>$ 3 in any WISE band but is not coincident with any extended sources from the 2MASS survey.
-We also require only one match to a 2MASS source within 3".
+\subsection{Merged Samples}
+The OGLE-III AGB sample, as well as the SDSS and YSO samples, were matched to the ALLWISE data products via the {\tt GATOR} tool at the NASA/IPAC Infrared Science Archive\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}}.
+The accepted matching radius was 3" execept in the case of YSOs, which were limited to 1".
+YSO objects were obtained from the VIZIER service already possessing WISE observations, so we were able to accept a smaller matching radius as we only sought the extra 2MASS information.
 
-{\color{red}{How was it merged, which further cuts and why...}}
+The final OGLE-2MASS-WISE AGB sample is produced from the obejct sample in section~\ref{sec:ogle}.
+We apply the point source saturation limits of [$W1/W2/W3/K_S$] $<$ [$2.0/1.5/-3.0/8.5$].
+We also enforce, in order, the 5$\sigma$ faint limits for [$W1/W2/W3/K_S$]  $<$ [$16.83/15.6/11.32/15.5$].
+The WISE point source saturation and 5$\sigma$ faint limits can be found in \cite{2012wise.rept....1C}.
+The 2MASS faint and saturation limits are found in \cite{2006AJ....131.1163S}.
+We also enforce [$W1/W2/W3$] SNR $> 3$, and contamination \& confusion flags ({\tt cc\_flag}) set to 0 in each of those bands.
+The VAGC, LRG, and YSO samples are filtered through the same criteria.
+Note that we deliberately neglect $W4$ because, after applying the 5$\sigma$ faint limits as well as {\tt cc\_flag}=0, our sample is too heavily reduced.
+The resulting populations from each reduction step can be found in Table~\ref{tab:reductions}.
 
-\subsection{Completeness and contamination of the initial \agb\ sample}
+\begin{table}[h]
+	\begin{center}
 
-In the next section you are using the \agb\ sample from this section to 
+	\scalebox{0.85}{
+%		\begin{tabular}{l r r r r r r r r r r}
+%		\hline\hline
+%		Population & Original & WISE match &  [$W1-3/K_S$] sat & [$W1-2$] fnt & [$W1-3$] fnt & [$W-3/K_S$] fnt & [$W-3$] SNR $> 3$ \\ 
+%		\hline
+%		OGLE-2MASS & 46,385 & 43,209& 43,201 & 43,200 & 19,358 & 19,358 & 18,995 \\
+%		VAGC-QSO & 122,550 & 6,902 & 6,900 & 6,891& 6,728 & 5,283 & 5,279\\
+%		VAGC-AGN & 19,184 & 3,098 & 3,098 & 3,098 & 3,005 & 2,878 & 2,876\\
+%		VAGC-SF & 232,845 & 36,539 & 36,538 & 36,537 & 35,678 & 34,299 & 34,285\\
+%		VAGC-SB & 67,128 & 10,359 & 10,358 & 10,358 & 10,105 & 9,676 & 9,673\\
+%		DR7-LRGs & 105,631& 75,543 & 75,527 & 75,527 & 1,184 & 1,167 & 1,079\\
+%		YSOs & 290 & 274 & 214 & 214 & 214 & 213 & 213\\
+%		WISE-2MASS Locus & - & 25,254 & 25,156 & 24,382 & 3,154 & 3,070 & 3,056\\
+		\begin{tabular}{l r r r r r r r r r r}
+		\hline\hline
+		Population & OGLE & QSO & AGN & SF & SB & LRGs & YSOs & Locus & \textbf{Total} \\ 
+		\hline
+		Original & 46,385 & 122,550 & 19,184 & 232,845 & 67,128 & 105,631 & 290 & 25,254 & $\mathbf{619,267}$ \\
+		WISE match & 43,209 & 6,902 & 3,098 & 36,539 & 10,359 & 75,543 & 274 & 25,254 & $\mathbf{201,178}$ \\
+		$W1$-3, $K_S$ saturation limit& 43,201 & 6,900 & 3,098 & 36,538 & 10,358 & 75,527 & 214 & 25,156 & $\mathbf{200,992}$ \\
+		 $W1$-2 faint limit & 43,200 & 6,891 & 3,098 & 36,537 & 10,358 & 75,527 & 214 & 24,382 & $\mathbf{200,207}$  \\
+		 $W1$-3 faint limit & 19,358 & 6,728 & 3,005 & 35,678 & 10,105 & 1,184 & 214 & 3,154 & $\mathbf{79,426}$  \\
+		 $W1$-3, $K_S$ faint limit & 19,358 & 5,283 & 2,878 & 34,299 & 9,676 & 1,167 & 213 & 3,070 & $\mathbf{75,944}$ \\
+		 $W1$-3 SNR $> 3$ & 18,995 & 5,279 & 2,876 & 34,285 & 9,673 & 1,079 & 213 & 3,056 & $\mathbf{75,456}$\\
+		\hline\hline
+		\end{tabular}}    
+	\caption{Sample populations with respect to each reduction step. \label{tab:reductions}}
+	\end{center}
+
+\end{table}
+
+\subsection{Completeness and Contamination of the Initial \agb\ Sample}
+Sample completeness $\eta$ is defined as
+\begin{eqnarray*}
+\eta &=& \frac{N - n_\text{missed}}{N}
+\end{eqnarray*}
+where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries. \cite{2013sdmm.book.....I} defines sample contamination as
+\begin{eqnarray*}
+\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
+\end{eqnarray*}
+where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
+Our goal is to get the lowest AGB sample (~1\% contamination) while still retaining a large number of AGB stars.
+
+The difficulty with determining contamination is that none of the contaminant sample, save for the locus stars,  falls within the region of the LMC.
+However, knowing that there is essentially a uniform distribution of extragalactic sources on the sky, we take the number density of objects in the SDSS footprint and scale it up by the area centered on the LMC.
+This should provide a reasonable estimate for the number of sources that we would find within the region of the LMC, and by consequence estimate our final contamination levels.
+
+{\color{red} In the next section you are using the \agb\ sample from this section to 
 estimate completeness and contamination of color-selected samples. Here
@@
 are for your initial sample constructed using SIMBAD and optical surveys.
-In particular, it is important to describe expected distance limits.
+In particular, it is important to describe expected distance limits. }
 

### 2015-08-06 paper1/paper1_intro.tex
Updating paper with section 2.6

@@
 for very numerous main sequence stars, is not reached when studying disk component at low galactic
-latitudes. As shown, for example, by Berry et al., the extinction due to interstellar dust limits 
+latitudes. As shown, for example, by \cite{2012ApJ...757..166B}, the extinction due to interstellar dust limits 
 the SDSS sample of main sequence stars to heliocentric distances of only a few kpc. In order

### 2015-08-10 paper1/paper1_criteria.tex
Finished section 2

--- 2015-03-31 paper1/paper1_criteria.tex
@@
 \label{sec:criteria}
-We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
+\subsection{Overview}
+{\color{red}Some words that introduce the section.
+What do we do here?
+Calculate completeness and contamination for various color cuts.
+Re-emphasize that the extragalactic sources are not in the region of the LMC, so we estimate their numbers (and thus the contamination fraction) by scaling their number density per square degree up by the space occupied by the LMC.
+Then we add some more color cuts to reject YSOs.
+We should also do something with O-rich and C-rich criteria from Nikutta et al 2014, however it does use W4 which we neglect.}
 
-The color-color criteria for AGB selection are as follows:
-\begin{align} 
-(J-K_s) > 1.1\label{eq:criteria1}\\
-(W2-W3) > 0.3\label{eq:criteria2}\\
-(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
-\end{align}
-The criteria in (1) and (2) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies).
-Criteria in (3) also reject stars from the stellar locus, but primarily function to reject IR-bright extragalactic sources.
-
-\noindent Sample completeness $\eta$ is defined as
+\subsection{Completeness and Contamination of Initial Samples}
+Sample completeness $\eta$ is defined as
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $N$ is the total number of objects in that sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries.
-Figure~\ref{fig:completeness} shows the distribution of completeness amongst AGB sources after the application of the above criteria.
-Sample contamination is typically defined as
+where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries. \cite{2013sdmm.book.....I} defines sample contamination as
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected}$ is the total number of selected sources, including AGB objects and spurious objects \citep{2013sdmm.book.....I}.
-The contamination map is shown in Figure~\ref{fig:contamination}.
+where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
+Our goal is to get the lowest AGB sample ($\le$1\% contamination) while still retaining a large number of AGB stars.
+
+The difficulty with determining contamination is that none of the contaminant sample, save for the locus stars,  falls within the region occupied by our AGB sample: the LMC.
+However, knowing that there is essentially a uniform distribution of extragalactic sources on the sky {\color{red}[a citation would be great]}, we solve for the number density of objects in the 10,417 sq.
+deg.
+DR7 footprint.
+We then scale these number densities up by the $\sim76.5$ deg$^2$ area occupied by the LMC.
+This should provide a reasonable estimate for the number of sources that we would find within the region of the LMC, and by consequence estimate our final levels of extragalactic contamination in WISE-2MASS color-color space.
+
+We employ the following color criteria to isolate AGB stars from their potential contaminants:
+\begin{eqnarray}
+(J - K_s) &>& 1.1\\
+(W2 - W3) &<&  2.2
+\end{eqnarray}
+\noindent Figure~\ref{fig:colorcuts} shows completeness distributions for the OGLE-2MASS-WISE AGB sample, and corresponding contamination distributions for each contaminant sample.
+Tables~\ref{tab:criteria_completeness} \&~\ref{tab:criteria_contamination} show the completeness and contamination fractions for each color criterion.
 
 \begin{figure}[h]
-\includegraphics[width=6.7in]{figs/completeness_map.pdf}
-\caption{Words \label{fig:completeness}.}
-\end{figure}
-\begin{figure}[h]
-\includegraphics[width=6.7in]{figs/contamination_map.pdf}
-\caption{Words \label{fig:contamination}.}
+\includegraphics[width=3.3in]{figs/completeness_contamination_jkcut.pdf}\includegraphics[width=3.3in]{figs/completeness_contamination_w23cut.pdf}
+\caption{\emph{Left:} Cumulative completeness and contamination distributions before $(J-K_s)$ color cut.
+\emph{Right:} Cumulative completeness and contamination after $(J-K_s)$ cut and before $(W2-W3)$ cuts.
+Colors are the same as Figure~\ref{fig:wisecontaminants}.
+Thick black line in top panels is OGLE-2MASS-WISE AGB completeness as a function of color.
+Bin width for both panels is 0.1 dex.
+\label{fig:colorcuts}}
 \end{figure}
 
-A fairly large issue for creating the boundaries are the O-rich stars.
+\begin{table}[h]
+	\begin{center}
+	\scalebox{0.85}{
+		\begin{tabular}{l r r r}
+		\hline\hline
+		Criteria & Completness \\
+		\hline
+		$(J-K_s) > 1.1$ & 89.62\%\\
+		$(W2-W3) < 2.2$ & 82.40\%\\
+		\hline\hline
+		\end{tabular}}    
+	\caption{Sample completeness with respect to each successive cut on WISE-2MASS color.\label{tab:criteria_completeness}}
+	\end{center}
+
+\end{table}
+\begin{table}[h]
+	\begin{center}
+	\scalebox{0.85}{		
+		\begin{tabular}{l r r r r r r r r}
+		\hline\hline
+		Critiera & QSO & AGN & LRG & SF & SB & Locus \\
+		\hline
+		$(J-K_s) > 1.1$ & 0.18\% & 0.09 \%& 0.04\% & 0.18\% & 0.31\% & 2.26\%\\
+		$(W2-W3) < 2.2$ & 0\% & 0\%& 0.01\% & 0\% & 0\% & 0.83\%\\
+		\hline\hline
+		\end{tabular}
+		}    
+	\caption{Sample contamination with respect to each successive cut on WISE-2MASS color.\label{tab:criteria_contamination}}
+	\end{center}
+\end{table}
+
+Figure~\ref{fig:lmc_completeness} shows the distribution of recovered AGB stars on the sky as well as in color-color space.
+We reserve an accounting of YSOs for the next section, as there is significant overlap between YSOs and AGBs, and they are not evenly distributed across the sky.
+
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/ogle_completeness_map.png}
+\caption{Map of completeness fractions for AGB stars in the LMC \label{fig:lmc_completeness}}
+\end{figure}
+
+\subsection{Accounting for YSOs}
+{\color{red}Talk about how you reject YSOs}
+
+%In creating color-color criteria to generate a catalog of AGB candidates, we seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.
+%
+%The color-color criteria for AGB selection are as follows:
+%\begin{align} 
+%	(J-K_s) > 1.1\label{eq:criteria1}\\
+%	(W2-W3) > 0.3\label{eq:criteria2}\\
+%	(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
+%\end{align}
+%The criteria in (\ref{eq:criteria1}) and (\ref{eq:criteria2}) are primarily concerned with rejecting objects from the stellar locus, and other objects whose NIR spectra are dominated by the Rayleigh-Jeans tail.
+(\ref{eq:criteria3}) also rejects stars from the stellar locus, but primarily functions to remove IR-bright extragalactic sources.
+%
+%We also experiment with other criteria to focus more closely on the high-reliability AGB population, instead of trying to capture the most AGB-like objects. These criteria are as follows:
+%\begin{align} 
+%	(W1-W2) < 1\label{eq:criteria4}\\
+%	(W2-W3) < 1\label{eq:criteria5}
+%\end{align}
+%Criteria (\ref{eq:critera4}) restricts NIR excesses. Criteria (\ref{eq:criteria5}) primarily removes extragalactic sources, while only slightly cutting into the distribution of known AGB stars.
+%
+%\noindent Sample completeness $\eta$ is defined as
+%\begin{eqnarray*}
+%\eta &=& \frac{N - n_\text{missed}}{N}
+%\end{eqnarray*}
+%where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries.
+Figure~\ref{fig:completeness} shows the distribution of sample completeness amongst AGB sources after the application of the above criteria.
+The vast majority (79.07\%) of Galactic AGB stars is recovered after criteria (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}) are applied.
+The largest losses occur at the edges of the Galactic disk ($\lvert b\rvert\approx10^\circ$) and in regions of high stellar number density both in the Galaxy and the Magellanic Clouds.
+In the color-color diagrams, selection completeness degrades near the borders of the selection area, as objects that straddle these boundaries may exhibit enough emission in other color-color spaces to be removed by our criteria.
+Of the remaining sample of 5,709 objects, 52.94\% lie within the disk ($\lvert b\rvert<10^\circ$).
+%
+%\begin{table}[h]
+%    \begin{center}
+%        \caption{Samples Recovered wih Application of Criteria (\%)}
+%        \scalebox{0.85}{
+%            \begin{tabular}{l r r r r r r r r r r r r}
+%\hline
+%Object Type & Reduced & (1) & (2) & (3) & (1,2) & (1,2,3) & | & \textbf{(4)} & \textbf{(5)} & \textbf{(1,2,4)} & \textbf{(1,2,5)} & \textbf{(1,2,4,5)}\\
+%\hline
+%All AGBs & 7220 & 95.15 & 84.89 & 95.21 & 82.62 & 79.07 & | & \textbf{94.22} & \textbf{88.82} & \textbf{77.05} & \textbf{72.71} & \textbf{71.14}\\ 
+%O-rich AGB & 3147 & 98.98 & 100.00 & 100.00 & 98.98 & 98.98 & | & \textbf{98.44} & \textbf{97.33} & \textbf{97.43} & \textbf{96.31} & \textbf{94.88}\\ 
+%C-rich AGB & 540 & 99.26 & 99.81 & 99.26 & 99.07 & 98.33 & | & \textbf{57.59} & \textbf{63.33} & \textbf{56.85} & \textbf{62.59} & \textbf{54.44}\\ 
+%Unclassed AGB & 3533 & 91.11 & 69.15 & 90.32 & 65.53 & 58.39 & | & \textbf{96.07} & \textbf{85.14} & \textbf{61.99} & \textbf{53.24} & \textbf{52.53}\\ 
+%DR12 SSPP & 67508 & 76.16 & 99.16 & 0.91 & 76.07 & 0.05 & | & \textbf{86.05} & \textbf{0.89} & \textbf{65.67} & \textbf{0.03} & \textbf{0.03}\\ 
+%DR7 LRG & 84 & 96.43 & 100.00 & 0.00 & 96.43 & 0.00 & | & \textbf{89.29} & \textbf{0.00} & \textbf{85.71} & \textbf{0.00} & \textbf{0.00}\\ 
+%QSO & 3783 & 73.54 & 99.97 & 0.13 & 73.54 & 0.11 & | & \textbf{34.50} & \textbf{0.11} & \textbf{27.54} & \textbf{0.08} & \textbf{0.08}\\ 
+%Galaxy & 42066 & 78.55 & 100.00 & 0.01 & 78.54 & 0.00 & | & \textbf{96.39} & \textbf{0.01} & \textbf{75.31} & \textbf{0.00} & \textbf{0.00}\\ 
+%AGN & 1011 & 79.33 & 100.00 & 0.00 & 79.33 & 0.00 & | & \textbf{96.34} & \textbf{0.00} & \textbf{75.87} & \textbf{0.00} & \textbf{0.00}\\ 
+%
+%\hline
+%                \label{tab:criteria}
+%            \end{tabular}}    
+%    \end{center}
+%\end{table}
+%
+%
+%\begin{figure}[h]
+%\includegraphics[width=6.7in]{figs/completeness_map.pdf}
+%\caption{Sample selection completeness maps in ($l,b$) space (\emph{top}) and the color-color space of our selection criteria (\emph{bottom}).
+Color scale shows the completeness fraction per bin, with 4-deg$^2$ bins on the Galactic map and 0.1 dex bins on each axis of the color-color diagrams.
+Selection criteria are shown as dashed lines.\label{fig:completeness}}
+%\end{figure}
+%
+%\cite{2013sdmm.book.....I} defines sample contamination as
+%\begin{eqnarray*}
+%\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
+%\end{eqnarray*}
+%where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
+The contamination map is shown in Figure~\ref{fig:contamination}.
+Most bins in Figure~\ref{fig:contamination} exhibit 0\% contamination, with the overall contamination level at 0.38\%.
+What contaminants do remain exist primarily at the very fringes of the AGB star distribution in ($J-K_S$) vs ($W2 - W3$) space and near the redder boundary in ($W3-W4$) vs ($W2 - W3$), where bluer extragalactic sources creep into the selection region.
+The results of the applied criteria on both the collective AGB and contaminant samples are summarized in Table~\ref{tab:completeness}.
+%
+%We note that while we do achieve a low rate of contamination, most of our contaminant sources are out of the plane of the Galaxy, while many of our \agb\, sources are within the Galactic disk.
+Considering how our extragalactic contaminant sources are from various iterations of \sdss, we know that their spatial distributions should be effectively uniform across the sky {\color{red} [cite me]}.
+For contaminants unaffected by interstellar reddening, they would occupy the same regions of color-color space already marked by our existing contaminant sample, and would similarly be removed by (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}).
+This would produce the same zero-level contamination for all extragalactic sources aside from QSOs.
+We can adopt a QSO spatial density by dividing the total number of spectroscopically identified QSOs by the covered area on the sky.
+Using this, we recover a QSO spatial density of {\color{red} some number per square degree}.
+Using that density, and assuming ubiquity of sources on the sky, complete detection of QSOs {\color{red} down to some magnitude limit}, and a lack of interstellar reddening, we can estimate a maximum QSO contamination fraction of {\color{red} some number}.
+We use this number to report our QSO contamination fraction in Table~\ref{tab:completeness}, as QSOs significantly affected by interstellar reddening would exist further outside of the limits of (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}).
+A more rigorous study of QSO spatial density is outside the scope of this paper.
+%
+%The remaining contaminant sample is from the \sdss\, DR12 SSPP, with the number of remaining species found within shown in Figure~\ref{fig:sspp_histo}.
+{\color{red} the figures don't match up with each other.
+The number of red stars on the right side of the fig are significantly less than those in the histogram.
+Reconcile this.}
+%
+%\begin{figure}[h]
+%\includegraphics[width=6.7in]{figs/contamination_map.pdf}
+%\caption{Sample contamination in Galactic ($l,b$) space (\emph{top}) and color-color space (\emph{bottom}). Boundaries and binning the same as Fig~\ref{fig:completeness}.\label{fig:contamination}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%\includegraphics[width=6.7in]{figs/contaminating_sspp_objects.pdf}
+%\includegraphics[width=6.7in]{figs/contaminating_sspp_objects_classes.pdf}
+%\caption{Some histogram\label{fig:sspp_histo}}
+%\end{figure}
+%
+%%One large issue with the boundaries are the O-rich AGB stars.
 Either the objects from the OGLE O-rich AGB star catalog were mostly mis-matched to WISE, taking on colors of the stellar locus, or the issue is more physical.
@@
 Previous work on warm O-rich AGB stars has shown that their circumstellar shells are not prominent, and their NIR photometry reflects the Rayleigh Jeans tail of a cool 2000-4000K blackbody \citep{1974ApJ...189...89D}.
-
-The results of the applied critiera are summarized in Table~\ref{tab:completeness}. Of the objects remaining in the sample, 52.94\% of the AGBs have $\lvert b\rvert \le 10^\circ$.
-
-\begin{table}[h]
-    \begin{center}
-        \caption{Sample Completeness and Contamination}
-        \scalebox{0.85}{
-            \begin{tabular}{l c c c c c c}
-		\hline
-		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
-		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
-		\hline
-		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
-		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
-		\hline
-		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
-		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
-		\hline
-		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
-		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
-		\hline
-
-                \label{tab:completeness}
-            \end{tabular}}    
-    \end{center}
-\end{table}
+%
+%\begin{table}[h]
+%    \begin{center}
+%        \caption{Sample Selection Completeness and Contamination}
+%        \scalebox{0.85}{
+%            \begin{tabular}{l c c c c c c}
+%		\hline
+%		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
+%		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
+%		\hline
+%		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
+%		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
+%		\hline
+%		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
+%		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
+%		\hline
+%		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
+%		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
+%		\hline
+%
+%                \label{tab:completeness}
+%            \end{tabular}}    
+%    \end{center}
+%\end{table}
+%

### 2015-08-10 paper1/paper1_data.tex
Finished section 2

@@
 the SDSS unofficial Luminous Red Galaxy sample,
-and young stellar objects obtained identified in \wise\ data, also described below. 
+and young stellar objects identified in \wise\ data, also described below. 
 We conclude this section by describing how these auxiliary catalogs were positionally merged
-with the WISE-2MASS catalog, and summarize the object counts before and after quality control in 
-Tables~\ref{tab:reductions} \& 2.
+with the WISE-2MASS catalog, summarize the object counts before and after quality control (Table~\ref{tab:reductions}), and summarize the effects of WISE-2MASS color-color cuts (Tables~\ref{tab:criteria_completeness} \& ~\ref{tab:criteria_contamination}) on the AGB and contaminant samples.
 
@@
 In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
-The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
+The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times on average with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
 Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$].
-The positions of objects in the \wise\, catalog were calibrated to the \twomass\, point source catalog.
+The positions of objects in the \wise\, catalog were calibrated using the \twomass\, point source catalog.
 
@@
 During this phase, \wise\, surveyed $70\%$ of the sky.
-Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR.
+Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in higher signal-to-noise for most sources.
 The overall number of sources compiled into \allwise\, totals over 747.6 million.
@@
 \subsection{AGB stars from the OGLE-III Catalog of Variable Stars}\label{sec:ogle}
-The OGLE-III Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall OGLE-III experiment, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars.
-With high-precision photometry \citep[$~0.001$ mag][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
+The OGLE-III Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall OGLE-III dataset, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars, with typically 700-900 data points per star in the $I$-band and about 50 data points per star in $V$-band.
+With high-precision photometry \citep[$~0.01$ mag][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
+The variability of these stars covers a wide range of periodicity ($4 < P < 2000$ days) and amplitudes ($0.005 < A_I < 5.7$ mag).
 
-The OGLE-III CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with LPVs classified by object type, evolutionary status, and spectral type.
+The OGLE-III CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with long-period variables (LPVs) classified by object type, evolutionary status, and spectral type.
 Object type is based chiefly on variability and luminosity.
 OGLE Small Amplitude Red Giants (hereon OSARGs) are weakly-variable ($0.005 < A_I < 0.13$ mag), with relatively short periods \citep[$10 < P < 100$ days;][]{2004AcA....54..129S}.
-Additionally, because these objects are multiperiodic, OSARGs are selected using their period ratios and their distances from established period-luminosity (log $_{10} P-L$) sequences from \cite{1999IAUS..191..151W}.
-For more detail in their selection, see \cite{2007AcA....57..201S}.
+Additionally, because these objects are multiperiodic, OSARGs are selected using their period ratios and their distances from established period-luminosity (log$P-L$) sequences from \cite{1999IAUS..191..151W}.
+For a more detailed description of their selection, see \cite{2007AcA....57..201S}.
 Mira variables and Semi-Regular Variables (hereon SRVs) are identified using $I$-band amplitudes and $P-W_I$, where $W_I$ is the reddening-free Wesenheit index \citep{2005AcA....55..331S}:
@@
 
-As red giant branch stars (hereon RGBs) can contaminate the same log$_{10} P-L$ space as AGB stars, evolutionary type is distinguished in two ways.
-Stars above the tip of the RGBÑ[$[I/K_s] = [14.56/12.05]$ mag in LMC \citep{2004AcA....54..129S, 2007AcA....57..201S}Ñare AGBs by virtue of luminosity.
+As red giant branch stars (RGBs) can contaminate the same log($P)-L$ space as AGB stars, evolutionary type is distinguished in two ways.
+Stars above the tip of the RGB ($K_s = 12.05$ mag in LMC; \citealt{2004AcA....54..129S, 2007AcA....57..201S}) are classified as AGBs.
 Below that threshold, a narrow sequence of stars appears to share the same slope and intercept with AGB stars above the TRGB.
 Additionally they share similar primary-secondary period ratios as known AGB stars \citep{2005AcA....55..331S}.
-However, as there is still some significant contamination by RGBs, we neglect all OGLE objects below the TRGB moving forward.
+However, as there is still some contamination by RGBs using period ratios, we reject all OGLE objects below the TRGB.
 
-For individual log$_{10} P-L$ sequences, spectral type (O-rich or C-rich) is easily seen as separations in log$_{10} P-L$ space as well as visible-NIR color-color space.
-The initial classifications are based on spectroscopically-observed stars and were extended more generally to log$_{10} P-L$ and visible-NIR color-color criteria \citep{2005AcA....55..331S, 2007AcA....57..201S}.
+For individual log$(P)-L$ sequences, spectral type (O-rich or C-rich) is easily seen as separations in log$(P)-L$ space as well as visible-NIR color-color space.
+The initial classifications are based on spectroscopically-observed stars and were extended more generally to log$(P)-L$ and visible-NIR color-color criteria \citep{2005AcA....55..331S, 2007AcA....57..201S}.
 This clear separation can be seen in Figure~\ref{fig:oglepmag}.
 
-From the initial OGLE-III sample of 52,976 LPVs, we retain 43,209 after matching to ALLWISE within 3", and ensuring only 1 match to 2MASS as well as all objects needing to be in the LMC.
+From the initial OGLE-III sample of 52,976 LPVs, we retain 43,262 after matching to ALLWISE within 3", and ensuring only 1 match to 2MASS as well as all objects needing to be in the LMC.
 The $I$-band median of OGLE-III LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag.
-Out of the entire sample only 13 objects (Miras) fall beyond the 20.5 mag faint limit, and 41 beyond the saturation limit.
+Out of the entire sample only 13 objects (Miras) fall beyond the 20.5 mag 5$\sigma$ faint limit, and 41 beyond the saturation limit.
 As such, we expect this to effectively represent the complete sample of oscillating AGB stars in the LMC.
+
+Out of the 43,262 objects matched to ALLWISE within 3", 1,633 are Mira variables, 30,768 are OSARGs, and 10,861 are SRVs.
+The right side of Figure~\ref{fig:oglepmag} shows color-color plots for OGLE-2MASS colors, with a line at $(J-K_s) = 1.4$ drawn to separate O-rich and C-rich Miras.
+Out of the total of 457 O-rich Miras, 36 (7.9\%) have $(J-K_s) > 1.4$.
+Conversely, out of the 1,176 C-rich total Miras, 1,138 (96.8\%) have $(J-K_s) > 1.4$.
+The distribution of these objects on the sky is summarized in Figure~\ref{fig:oglegalplot}.
 
@@
 \includegraphics[width=3in]{figs/ogle_2mass_color_color_threeplot.png}
-\caption{O-rich (\emph{blue}) and C-rich (\emph{red}) OGLE-III LPVs matched to 2MASS. \emph{Left}: $W_I$ vs log period in days. \emph{Right:} OGLE-III $(V-I)$ vs 2MASS $(J-K_s)$. \label{fig:oglepmag}}
+\caption{O-rich (\emph{blue}) and C-rich (\emph{red}) OGLE-III LPVs matched to 2MASS.
+\emph{Left}: $W_I$ vs log period in days.
+\emph{Right:} OGLE-III $(V-I)$ vs 2MASS $(J-K_s)$.
+The dashed line represents $(J-K_s) = 1.4$.
+\label{fig:oglepmag}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=3in]{figs/ogle_2mass_wise_miras.pdf}
+\caption{O-rich (\emph{blue}) and C-rich (\emph{red}) OGLE-III Miras with ($J-K_s$) $>$ 1.4. \label{fig:oglegalplot}}
 \end{figure}
@@
 \subsection{Extragalactic catalogs from SDSS}
-SDSS is a multi-year survey of roughly 25\% of the sky collecting both photometric (\emph{u, g, r, i, z}-bands for 700+ million objects) and spectroscopic measurements (1.6+ million) of stars and extragalactic sources \citep{2000AJ....120.1579Y}.
+SDSS is a multi-year survey of roughly 25\% of the sky that collected both photometric (\emph{u, g, r, i, z}-bands for 700+ million objects) and spectroscopic measurements (1.6+ million) of stars and extragalactic sources \citep{2000AJ....120.1579Y}.
 Though its ground-based nature and sensitivity to interstellar extinction prevent SDSS from being an all-sky survey, it represents the largest spectrophotometric catalog to date.
-For this study we collect extragalactic objects from the NYU Value-added Galaxy Catalog, containing objects matched between SDSS and several auxiliary surveys spanning multiple wavelength regimes, and the unofficial SDSS Luminous Red Galaxy sample, isolating high-confidence luminous red galaxies from the SDSS database.
+For this study we collect extragalactic objects from the NYU Value-added Galaxy Catalog, containing objects matched between SDSS and several auxiliary surveys spanning multiple wavelength regimes, and the SDSS Luminous Red Galaxy sample, isolating high-confidence luminous red galaxies from the SDSS database.
+The color-color distributions of these sample sets are shown in Figure~\ref{fig:wisecontaminants}.
 
@@
 
-\subsubsection{The Unofficial Luminous Red Galaxy Sample}
+\subsubsection{The Luminous Red Galaxy Sample}
 Subselected for the study of Baryon Acoustic Oscillations by \cite{2010ApJ...710.1444K}, the Luminous Red Galaxy sample \citep[LRGs;][]{2001AJ....122.2267E} is also sourced from SDSS DR7.
-LRGs contaminate the color-color space of AGB stars due to the severe redshift of their intrinsically-bright optical fluxes.
-We obtain the unofficial LRG sample from the data repository\footnote{\url{http://cosmo.nyu.edu/~eak306/SDSS-LRG.html}} of \cite{2010ApJ...710.1444K}.
+LRGs contaminate the IR color-color space of AGB stars due to their large redshifts ($z\sim0.3$).
+We obtain the  LRG sample from the data repository\footnote{\url{http://cosmo.nyu.edu/~eak306/SDSS-LRG.html}} of \cite{2010ApJ...710.1444K}.
 The initial sample is volume-limited, containing 105,631 objects spanning a redshift range of $0.16 < z < 0.47$.
+
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/contaminant_color_color_distr.pdf}
+\caption{Extragalactic sources from SDSS DR7's VAGC and LRG samples.
+Objects are (\emph{black}) starforming galaxies, (\emph{blue}) starburst galaxies, (\emph{red}) QSOs, (\emph{green}) AGN, and (\emph{purple}) LRGs.
+\label{fig:wisecontaminants}}
+\end{figure}
 
@@
 That warm dust can then glow brightly in the MIR, easily contaminating AGB color-color space.
-To characterize and later eliminate this potential contaminant we select as our base sample 290 YSOs from \cite{2011ApJS..196....4R}, a survey searching for YSOs in the Taurus Molecular Cloud using data from WISE, ancillary data from SDSS and 2MASS, and covering {\color{red} some number} deg$^2$.
-As YSO tend to be embedded within high-extinction clouds of this sort, we expect that the dust-reddened NIR/MIR emission from these objects should match to similar objects in the LMC when we later calibrate our color-color criteria.
+To characterize and later eliminate this potential contaminant, we select as our base sample 290 YSOs from \cite{2011ApJS..196....4R}, a survey searching for YSOs in the Taurus Molecular Cloud using data from WISE, ancillary data from SDSS and 2MASS, and covering 260 deg$^2$.
+As YSO tend to be embedded within high-extinction clouds, we expect that the dust-reddened NIR/MIR emission from these objects should match to similar objects in the LMC when we later calibrate our color-color criteria.
 
 \subsection{WISE+2MASS Stellar Locus}
-We begin to extract the stellar locus from ALLWISE by centering on the LMC ($276.5^\circ < l < 284^\circ$, $-38.2^\circ < b < -28^\circ$) by necessitating that every object have 1 2MASS association, [$W1/W2/W3$] signal-to-noise $>$ 1, detections in 2MASS J, H, and K,  and confusion \& contamination flags set to 0 for [$W1/W2/W3$].
+The color-color space occupied by AGB stars showing significant photospheric emission is heavily populated by naked stars from the main sequence.
+By sheer number, these stars would drown out sources that could be AGB stars with thin circumstellar envelopes.
+To address this, we extract the stellar locus from ALLWISE in the direction of the LMC ($276.5^\circ < l < 284^\circ$, $-38.2^\circ < b < -28^\circ$).
+We require that every object have 1 2MASS association, [$W1/W2/W3$] signal-to-noise $>$ 1, detections in 2MASS J, H, and K,  and confusion \& contamination flags set to 0 for [$W1/W2/W3$].
 We then adapt the color-color criteria from \cite{2014MNRAS.440.3430D}.
@@
 
-\subsection{Merged Samples}
+\subsection{Merged Samples\label{sec:merged}}
 The OGLE-III AGB sample, as well as the SDSS and YSO samples, were matched to the ALLWISE data products via the {\tt GATOR} tool at the NASA/IPAC Infrared Science Archive\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}}.
@@
 The final OGLE-2MASS-WISE AGB sample is produced from the obejct sample in section~\ref{sec:ogle}.
-We apply the point source saturation limits of [$W1/W2/W3/K_S$] $<$ [$2.0/1.5/-3.0/8.5$].
-We also enforce, in order, the 5$\sigma$ faint limits for [$W1/W2/W3/K_S$]  $<$ [$16.83/15.6/11.32/15.5$].
+We apply the point source saturation limits of [$W1/W2/W3/K_s$] $<$ [$2.0/1.5/-3.0/8.5$].
+We also enforce, in order, the 5$\sigma$ faint limits for [$W1/W2/W3/K_s$]  $<$ [$16.83/15.6/11.32/15.5$].
 The WISE point source saturation and 5$\sigma$ faint limits can be found in \cite{2012wise.rept....1C}.
 The 2MASS faint and saturation limits are found in \cite{2006AJ....131.1163S}.
-We also enforce [$W1/W2/W3$] SNR $> 3$, and contamination \& confusion flags ({\tt cc\_flag}) set to 0 in each of those bands.
+To ensure quality measurements in the desired bands, we also enforce [$W1/W2/W3$] SNR $> 3$, and contamination \& confusion flags ({\tt cc\_flag}) set to 0 in each of those bands.
 The VAGC, LRG, and YSO samples are filtered through the same criteria.
@@
 	\scalebox{0.85}{
-%		\begin{tabular}{l r r r r r r r r r r}
-%		\hline\hline
-%		Population & Original & WISE match &  [$W1-3/K_S$] sat & [$W1-2$] fnt & [$W1-3$] fnt & [$W-3/K_S$] fnt & [$W-3$] SNR $> 3$ \\ 
-%		\hline
-%		OGLE-2MASS & 46,385 & 43,209& 43,201 & 43,200 & 19,358 & 19,358 & 18,995 \\
-%		VAGC-QSO & 122,550 & 6,902 & 6,900 & 6,891& 6,728 & 5,283 & 5,279\\
-%		VAGC-AGN & 19,184 & 3,098 & 3,098 & 3,098 & 3,005 & 2,878 & 2,876\\
-%		VAGC-SF & 232,845 & 36,539 & 36,538 & 36,537 & 35,678 & 34,299 & 34,285\\
-%		VAGC-SB & 67,128 & 10,359 & 10,358 & 10,358 & 10,105 & 9,676 & 9,673\\
-%		DR7-LRGs & 105,631& 75,543 & 75,527 & 75,527 & 1,184 & 1,167 & 1,079\\
-%		YSOs & 290 & 274 & 214 & 214 & 214 & 213 & 213\\
-%		WISE-2MASS Locus & - & 25,254 & 25,156 & 24,382 & 3,154 & 3,070 & 3,056\\
+
 		\begin{tabular}{l r r r r r r r r r r}
@@
 		WISE match & 43,209 & 6,902 & 3,098 & 36,539 & 10,359 & 75,543 & 274 & 25,254 & $\mathbf{201,178}$ \\
-		$W1$-3, $K_S$ saturation limit& 43,201 & 6,900 & 3,098 & 36,538 & 10,358 & 75,527 & 214 & 25,156 & $\mathbf{200,992}$ \\
+		$W1$-3, $K_s$ saturation limit& 43,201 & 6,900 & 3,098 & 36,538 & 10,358 & 75,527 & 214 & 25,156 & $\mathbf{200,992}$ \\
 		 $W1$-2 faint limit & 43,200 & 6,891 & 3,098 & 36,537 & 10,358 & 75,527 & 214 & 24,382 & $\mathbf{200,207}$  \\
 		 $W1$-3 faint limit & 19,358 & 6,728 & 3,005 & 35,678 & 10,105 & 1,184 & 214 & 3,154 & $\mathbf{79,426}$  \\
-		 $W1$-3, $K_S$ faint limit & 19,358 & 5,283 & 2,878 & 34,299 & 9,676 & 1,167 & 213 & 3,070 & $\mathbf{75,944}$ \\
+		 $W1$-3, $K_s$ faint limit & 19,358 & 5,283 & 2,878 & 34,299 & 9,676 & 1,167 & 213 & 3,070 & $\mathbf{75,944}$ \\
 		 $W1$-3 SNR $> 3$ & 18,995 & 5,279 & 2,876 & 34,285 & 9,673 & 1,079 & 213 & 3,056 & $\mathbf{75,456}$\\
@@
 		\end{tabular}}    
-	\caption{Sample populations with respect to each reduction step. \label{tab:reductions}}
+	\caption{Sample populations with respect to each reduction step.
+Please see section~\ref{sec:merged} for the appropriate saturation limits, faint limits, and WISE matching radii.\label{tab:reductions}}
 	\end{center}
@@
 
-\subsection{Completeness and Contamination of the Initial \agb\ Sample}
-Sample completeness $\eta$ is defined as
-\begin{eqnarray*}
-\eta &=& \frac{N - n_\text{missed}}{N}
-\end{eqnarray*}
-where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries. \cite{2013sdmm.book.....I} defines sample contamination as
-\begin{eqnarray*}
-\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
-\end{eqnarray*}
-where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
-Our goal is to get the lowest AGB sample (~1\% contamination) while still retaining a large number of AGB stars.
-
-The difficulty with determining contamination is that none of the contaminant sample, save for the locus stars,  falls within the region of the LMC.
-However, knowing that there is essentially a uniform distribution of extragalactic sources on the sky, we take the number density of objects in the SDSS footprint and scale it up by the area centered on the LMC.
-This should provide a reasonable estimate for the number of sources that we would find within the region of the LMC, and by consequence estimate our final contamination levels.
-
-{\color{red} In the next section you are using the \agb\ sample from this section to 
-estimate completeness and contamination of color-selected samples. Here
-you need to summarize what do you believe completeness and contamination
-are for your initial sample constructed using SIMBAD and optical surveys.
-In particular, it is important to describe expected distance limits. }
-
-\subsection{XXX old stuff from here to the end of this section: Data Sources}
-\label{sec:sources}
-In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
+%
+%\subsection{XXX old stuff from here to the end of this section: Data Sources}
+%\label{sec:sources}
+%In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}.
 The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m.
@@
 The overall number of sources compiled into \allwise\, totals over 747.6 million.
-
-In order to generate a reliable, high-confidence catalog of Galactic candidate \agb\, stars, we must first define color-color criteria from known \agb\, star samples.
+%
+%In order to generate a reliable, high-confidence catalog of Galactic candidate \agb\, stars, we must first define color-color criteria from known \agb\, star samples.
 We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}.
-
-\ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
+%
+%\ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II.
 LPVs were classified into 3 categories: \ogle\, Small Amplitude Red Giants (OSARGs), Semi-Regular Variables (SRVs), and Miras.
@@
 The resulting samples yield 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich).
-
-From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars).
+%
+%From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars).
 {\color{red}Why were these objects selected?
@@
 Sequence 1 primarily contains Mira variables pulsating in their fundamental modes, whereas Sequences 2-4 contain semi-regular variables in various pulsation modes.
-
-The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
+%
+%The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars.
 Objects are classified spectroscopically, though by a variety of methods owing to the heterogeneous data housed within \simbad.
@@
 Because there is a high likelihood that samples between \ogle, \macho, and \simbad overlap, we retain only unique objects after the initial data reduction in section~\ref{sec:reduction}.
-
-We use \sdss\, spectroscopic catalogs to find and quantify regions in NIR-MIR color-color space populated by plausible contaminant sources.
+%
+%We use \sdss\, spectroscopic catalogs to find and quantify regions in NIR-MIR color-color space populated by plausible contaminant sources.
 These include any Galactic stellar objects and planetary nebulae, as well as a host of extragalactic sources.
@@
 {\color{red} Include bit about YSOs and PNe from SIMBAD}
-
-\subsection{Data Reduction}
-\label{sec:reduction}
-We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise.
+%
+%\subsection{Data Reduction}
+%\label{sec:reduction}
+%We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise.
 We select only matches within 3" between  each sample and \allwise.
 All samples of \agb\, were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}, with no flags for confusion or contamination as a spurious source in any band.
-We also require only single associations with \twomass\, objects within 3", detections in [$J/K_S/W1/W2/W3/W4$], and SNR $>$ 3 in each \allwise\, band.
-
-The population for each sample from initial matching as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.
+We also require only single associations with \twomass\, objects within 3", detections in [$J/K_s/W1/W2/W3/W4$], and SNR $>$ 3 in each \allwise\, band.
+%
+%The population for each sample from initial matching as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.
 The WISE color-color distributions for the AGB and contaminant samples are shown in Figure~\ref{fig:distros}.
 \\
-
-\vspace{-10pt}
+%
+%\vspace{-10pt}
+%%\begin{table}[h]
+%%	\begin{center}
+%%	\caption{AGB and Contaminant Populations}
+%%	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
+%%		\hline
+%%		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
+%%		\hline
+%%		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
+%%		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
+%%		\hline\hline
+%%		Population & MACHO seq1 & seq2 & seq3 & seq4\\
+%%		\hline
+%%		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
+%%		Reduced & 927 & 642 & 263 & 336 \\
+%%		\hline\hline
+%%		Population & OGLE C-rich & O-rich\\
+%%		\hline
+%%		2" match & 11,417 & 38,369 \\
+%%		Reduced & 737 & 2515 \\
+%%		\hline\hline
+%%		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
+%%		\hline
+%%		2" match & 1,508,158 & 18,481 & 102,178 &  & 799,761 \\
+%%		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
+%%		\hline
+%%
+%%		\label{tab:pop}
+%%	\end{tabular}}
+%%	\end{center}
+%%\end{table}
+%
 %\begin{table}[h]
 %	\begin{center}
-%	\caption{AGB and Contaminant Populations}
+%	
 %	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
+%		\hline\hline
+%		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
 %		\hline
-%		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
+%		3" match & 1,689 & 14,209 & 9,027 & 406 & 1,081 \\
+%		Reduced & 684 & 1,782 & 3,241 & 43 & 511 \\ 
 %		\hline
-%		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
-%		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
+%		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
+%		\hline
+%		3" match & 5,279 & 3,519 & 2,619 & 3,070 \\
+%		Reduced & 277 & 185 & 73 & 61 \\ 
+%		\hline
+%		Population & OGLE-III C-rich & O-rich \\ 
+%		\hline
+%		3" match & 11,542 & 38,848 \\
+%		Reduced & 249 & 730 \\ 
+%		\hline
+%		Population & DR12 SSPP & DR7 LRG & QSO & AGN & Galaxies \\ 
+%		\hline
+%		3" match & 1,578,329 & 104,345 & 103,590 & 18,528 & 841,712 \\ 
+%		Reduced & 67,508 & 84 & 3,977 & 1,069 & 44,314 \\ 
 %		\hline\hline
-%		Population & MACHO seq1 & seq2 & seq3 & seq4\\
-%		\hline
-%		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
-%		Reduced & 927 & 642 & 263 & 336 \\
-%		\hline\hline
-%		Population & OGLE C-rich & O-rich\\
-%		\hline
-%		2" match & 11,417 & 38,369 \\
-%		Reduced & 737 & 2515 \\
-%		\hline\hline
-%		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
-%		\hline
-%		2" match & 1,508,158 & 18,481 & 102,178 &  & 799,761 \\
-%		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
-%		\hline
 %
-%		\label{tab:pop}
+%		
 %	\end{tabular}}    
+%	\caption{\agb\, and contaminant populations matched to WISE before and after sample reduction in section~\ref{sec:reduction}.
+MACHO sequences (seq1-seq4) are from \cite{2008AJ....136.1242F} and described briefly in section~\ref{sec:sources}.\label{tab:pop}}
 %	\end{center}
+%	
 %\end{table}
-
-\begin{table}[h]
-	\begin{center}
-
-	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
-		\hline\hline
-		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
-		\hline
-		3" match & 1,689 & 14,209 & 9,027 & 406 & 1,081 \\
-		Reduced & 684 & 1,782 & 3,241 & 43 & 511 \\ 
-		\hline
-		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
-		\hline
-		3" match & 5,279 & 3,519 & 2,619 & 3,070 \\
-		Reduced & 277 & 185 & 73 & 61 \\ 
-		\hline
-		Population & OGLE-III C-rich & O-rich \\ 
-		\hline
-		3" match & 11,542 & 38,848 \\
-		Reduced & 249 & 730 \\ 
-		\hline
-		Population & DR12 SSPP & DR7 LRG & QSO & AGN & Galaxies \\ 
-		\hline
-		3" match & 1,578,329 & 104,345 & 103,590 & 18,528 & 841,712 \\ 
-		Reduced & 67,508 & 84 & 3,977 & 1,069 & 44,314 \\ 
-		\hline\hline
-
-	\end{tabular}}    
-	\caption{\agb\, and contaminant populations matched to WISE before and after sample reduction in section~\ref{sec:reduction}.
-MACHO sequences (seq1-seq4) are from \cite{2008AJ....136.1242F} and described briefly in section~\ref{sec:sources}.\label{tab:pop}}
-	\end{center}
-
-\end{table}
-
-\begin{figure}
-\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
-\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
-\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
-\caption{Logarithmic number densities for objects in \wise\, and \twomass\, color-color space, binned in 0.1 dex on each axis.
+%
+%\begin{figure}
+%\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
+%\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
+%\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
+%\caption{Logarithmic number densities for objects in \wise\, and \twomass\, color-color space, binned in 0.1 dex on each axis.
 \emph{Left:} The combined AGB sample matched to ALLWISE.
 \emph{Right:} The combined contaminant sample.\label{fig:distros}}
-\end{figure}
+%\end{figure}

### 2015-08-10 paper1/paper1_intro.tex
Finished section 2

@@
 episodes with different observational characteristics: the early \agb\, phase (E-\agb)  and the thermally-pulsing \agb\, phase (TP).
-During the thermally-pulsing phase, \agb\, stars produce substantial dust-driven stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4}$ $M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in oxides (SiO, Al$_2$O$_3$, etc.) or carbon-rich molecules (SiC, AmC, etc.) The dominant chemical composition is highly dependent upon the metallicity of the host galaxy \citep{2005A&A...434..691M}.
+During the thermally-pulsing phase, \agb\, stars produce substantial dust-driven stellar winds \citep[$10^{-7} < \dot{M} < 10^{-4}$ $M_\odot$ yr$^{-1}$,][]{2002A&A...391.1053O} rich in oxides (SiO, Al$_2$O$_3$, etc.) or carbon-rich molecules (SiC, amC, etc.).
+The dominant chemical composition is highly dependent upon the metallicity of the host galaxy \citep{2005A&A...434..691M}.
 High-metallicity galaxies like  the Milky Way have a substantial population of oxygen-rich AGB stars \citep{1985A&A...152L...1H}, whereas low-metallicity galaxies such as the Magellanic clouds are dominated by carbon-rich \agb\ stars \citep{2011AJ....142..103B}.
@@
 copiously in the near- and mid-infrared (NIR \& MIR, respectively). Together with high bolometric luminosity
-(10$^3$--10$^4$ \Mo), this redistribution of the output radiation to the infrared wavelength makes \agb\ stars
+(10$^3$--10$^4$ \Mo, \citealt{2001ApJ...552..787K}), this redistribution of the output radiation to the infrared wavelength makes \agb\ stars
 excellent disk probe when infrared survey data are available. Indeed, they were detected all the way to the 

### 2015-08-10 paper1/paper1_data.tex
Slight correction

@@
 		\hline
-		Original & 46,385 & 122,550 & 19,184 & 232,845 & 67,128 & 105,631 & 290 & 25,254 & $\mathbf{619,267}$ \\
+		Original & 46,262 & 122,550 & 19,184 & 232,845 & 67,128 & 105,631 & 290 & 25,254 & $\mathbf{619,267}$ \\
 		WISE match & 43,209 & 6,902 & 3,098 & 36,539 & 10,359 & 75,543 & 274 & 25,254 & $\mathbf{201,178}$ \\

### 2015-08-27 paper1/paper1_criteria.tex
Another correction
Finished section 4

@@
 \label{sec:criteria}
-\subsection{Overview}
-{\color{red}Some words that introduce the section.
-What do we do here?
-Calculate completeness and contamination for various color cuts.
-Re-emphasize that the extragalactic sources are not in the region of the LMC, so we estimate their numbers (and thus the contamination fraction) by scaling their number density per square degree up by the space occupied by the LMC.
-Then we add some more color cuts to reject YSOs.
-We should also do something with O-rich and C-rich criteria from Nikutta et al 2014, however it does use W4 which we neglect.}
+In producing our candidate sample, we must consider our likely physical sources of contamination, and how removing those sources reduces the completeness of our criteria.
+In Section~\ref{sec:comp_cont} we consider sample completeness and contamination, and use the sample sets from Section~\ref{sec:merged} to draw simple criteria in WISE-2MASS color-color space.
+YSOs present a special case of IR contaminants, as they are fairly well embedded in the same regions as AGB stars.
+We consider the special case of YSOs in Section~\ref{sec:kill_ysos}, developing finer criteria to remove these objects while still preserving as much of the AGB sample as possible.
 
-\subsection{Completeness and Contamination of Initial Samples}
+\subsection{Completeness and Contamination of Initial Samples}\label{sec:comp_cont}
 Sample completeness $\eta$ is defined as
@@
 where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
-Our goal is to get the lowest AGB sample ($\le$1\% contamination) while still retaining a large number of AGB stars.
+Our goal is to get the lowest AGB sample ($\le$2\% contamination) while still retaining a large number of AGB stars.
 
@@
 \begin{eqnarray}
-(J - K_s) &>& 1.1\\
-(W2 - W3) &<&  2.2
+&(J - K_s) &>\;\;1.1\label{eq:criteria1}\\
+-0.1\;<&(W2 - W3) &<\;\; 2.5\label{eq:criteria2}
 \end{eqnarray}
 \noindent Figure~\ref{fig:colorcuts} shows completeness distributions for the OGLE-2MASS-WISE AGB sample, and corresponding contamination distributions for each contaminant sample.
-Tables~\ref{tab:criteria_completeness} \&~\ref{tab:criteria_contamination} show the completeness and contamination fractions for each color criterion.
+Table~\ref{tab:criteria_completeness_contamination} shows the completeness and contamination fractions for each color criterion.
+We include the lower limit in ($W2-W3$) because, although there's a higher penalty on completeness than reduction in contamination, we want to exclude the spike of locus stars with ($W2-W3$) $< -0.1$.
 
 \begin{figure}[h]
-\includegraphics[width=3.3in]{figs/completeness_contamination_jkcut.pdf}\includegraphics[width=3.3in]{figs/completeness_contamination_w23cut.pdf}
+\includegraphics[width=3.3in]{figs/completeness_contamination_jkcut.pdf}\includegraphics[width=3.3in]{figs/completeness_contamination_allcuts.pdf}
 \caption{\emph{Left:} Cumulative completeness and contamination distributions before $(J-K_s)$ color cut.
@@
 	\scalebox{0.85}{		
-		\begin{tabular}{l r r r}
-		\hline\hline
-		Criteria & Completness \\
-		\hline
-		$(J-K_s) > 1.1$ & 89.62\%\\
-		$(W2-W3) < 2.2$ & 82.40\%\\
-		\hline\hline
-		\end{tabular}}    
-	\caption{Sample completeness with respect to each successive cut on WISE-2MASS color.\label{tab:criteria_completeness}}
-	\end{center}
-
-\end{table}
-\begin{table}[h]
-	\begin{center}
-	\scalebox{0.85}{		
 		\begin{tabular}{l r r r r r r r r}
 		\hline\hline
-		Critiera & QSO & AGN & LRG & SF & SB & Locus \\
+		Critiera & QSO & AGN & LRG & SF & SB & Locus & \textbf{AGB Completeness}\\
 		\hline
-		$(J-K_s) > 1.1$ & 0.18\% & 0.09 \%& 0.04\% & 0.18\% & 0.31\% & 2.26\%\\
-		$(W2-W3) < 2.2$ & 0\% & 0\%& 0.01\% & 0\% & 0\% & 0.83\%\\
+		$(J-K_s) > 1.1$ & 0.34\% & 0.18 \%& 0.09\% & 0.34\% & 0.60\% & 4.48\% &90.97\%\\
+		$(W2-W3) < 2.2$ & 0.03\% & 0\%& 0.04\% & 0.03\% & 0.02\% & 1.60\% &89.43\% \\
+		$(W2-W3) > -0.1$ & 0.03\% & 0\%& 0.04\% & 0.03\% & 0.02\% & 1.59\% &88.83\% \\
+
 		\hline\hline
@@
 		}    
-	\caption{Sample contamination with respect to each successive cut on WISE-2MASS color.\label{tab:criteria_contamination}}
+	\caption{Sample contamination and completeness (\emph{bold}) with respect to each successive cut on WISE-2MASS color.\label{tab:criteria_completeness_contamination}}
 	\end{center}
@@
 
-\subsection{Accounting for YSOs}
-{\color{red}Talk about how you reject YSOs}
+\subsection{Accounting for YSOs}\label{sec:kill_ysos}
+YSOs often directly overlap with AGB stars in WISE color-color space, so we treat these objects separately from our  extragalactic contaminants.
+\cite{2011ApJS..196....4R} produced a catalog of YSOs in the Taurus Molecular Cloud to bring the total number of YSOs in that region up to 290.
+We use these and the AGB candidate criteria from Section~\ref{sec:comp_cont} to further refine our criteria and increase the purity of our sample.
+
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/taurus_field_YSOs.pdf}
+\caption{YSOs (\emph{red x}) in the Taurus Molecular Cloud \citep{2011ApJS..196....4R}.\label{fig:yso_field}}
+\end{figure}
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/ysos_taurus_color.pdf}
+\caption{YSOs (\emph{red x}) and AGB candidates (\emph{black o}) in the Taurus Molecular Cloud.\label{fig:ysos_agbs_taurus}}
+\end{figure}
+\begin{figure}[h]
+\includegraphics[width=6in]{figs/remove_ysos.pdf}
+\caption{YSOs (\emph{red x}) and AGB candidates (\emph{black o}) in the Taurus Molecular Cloud.\label{fig:ysos_lmc}}
+\end{figure}
+
+Figure~\ref{fig:yso_field} shows these 290 YSOs in the field of the Taurus Molecular Cloud, overplotted on WISE objects in that field.
+As shown in Table~\ref{tab:reductions}, after photometric and quality cuts we are left with 184 YSOs, and only 74 after applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2}.
+In contrast, applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2} as well as photometric criteria from Section~\ref{sec:merged} to the WISE field in the Taurus Molecular Cloud yields 6,995 AGB candidates.
+Figure~\ref{fig:ysos_agbs_taurus} shows the degree of entanglement in color-color space for YSOs and AGB candidates.
+We note, however, that because these are galactic AGB candidates, and the Milky Way is deficient in C-rich AGB stars, Figure~\ref{fig:ysos_agbs_taurus} misses the degree to which YSOs contaminate the C-rich sequence of AGB stars.
+To clarify this, we compare Taurus YSOs to AGB stars from the LMC (Figure~\ref{fig:ysos_lmc}).
+Based on their positions in color-color space, we add three more constraints to remove YSOs from our AGB candidate sample:
+
+\begin{align}
+(W1-W2) &< 0.75(W2-W3) - 0.33 & \text{for } (W2-W3) > 1.075\\
+(W1-W2) &> -(W2-W3) + 1.5 & \text{for }(W2-W3) > 1.075\\
+(W1-W2) &> 0.2 & \text{for }(W2-W3) > 1.3
+\end{align}
+
+Collectively, these criteria preserve the C-rich AGB sequence, retain a relatively high completeness fraction for the LMC objects (82.72\%), and reduce the contamination in the Taurus Molecular Cloud due to YSOs from 1.91\% to 0.4\%.
 

### 2015-08-27 paper1/paper1_data.tex
Finished section 3

@@
 Our principal data source is the merged WISE-2MASS catalog. In order to derive
-WISE-2MASS color-based selection criteria, we use %several training samples
-%described below: a heterogeneous sample of bright \agb\ stars listed in 
-%SIMBAD database and more homogneous fainter samples selected using 
-%optical time-domain surveys OGLE-III and MACHO. The latter samples 
-%include \agb\ stars from the LMC and SMC and are also used to 
-%calibrate color-absolute magnitude relations that underlie our distance
-%estimates. In order to assess what infrared populations could contribute to
-%contamination of selected \agb\ candidates, we utilize a number of SDSS
-%catalogs, also described below. We conclude this section by describing
-%how these catalogs auxiliary catalogs were  positionally merged with the 
-%WISE-2MASS catalog and summarize the object counts before and after quality control in Table 1 
-a homogenous sample of AGB stars from the OGLE-III Catalog of Variable Stars supplying objects in the Magellanic Clouds. This sample serves to calibrate color-color and color-absolute magnitude
+WISE-2MASS color-based selection criteria, we use a homogenous sample of AGB stars from the OGLE-III Catalog of Variable Stars supplying objects in the Magellanic Clouds.
+This sample serves to calibrate color-color and color-absolute magnitude
 relations that underlie our distance estimates. In order to assess what infrared populations
@@
 We conclude this section by describing how these auxiliary catalogs were positionally merged
-with the WISE-2MASS catalog, summarize the object counts before and after quality control (Table~\ref{tab:reductions}), and summarize the effects of WISE-2MASS color-color cuts (Tables~\ref{tab:criteria_completeness} \& ~\ref{tab:criteria_contamination}) on the AGB and contaminant samples.
+with the WISE-2MASS catalog, summarize the object counts before and after quality control (Table~\ref{tab:reductions}), and summarize the effects of WISE-2MASS color-color cuts (Table~\ref{tab:criteria_completeness_contamination}) on the AGB and contaminant samples.
 
@@
 The OGLE-III Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall OGLE-III dataset, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars, with typically 700-900 data points per star in the $I$-band and about 50 data points per star in $V$-band.
-With high-precision photometry \citep[$~0.01$ mag][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
+With high-precision photometry \citep[$\sim0.01$ mag,][]{2007AcA....57..201S}, these observations saturate at $I = 12.5$ mag, and are limited at the faint end at $I = 20.5$ mag \citep{2001AcA....51..303Z}.
 The variability of these stars covers a wide range of periodicity ($4 < P < 2000$ days) and amplitudes ($0.005 < A_I < 5.7$ mag).
@@
 Mira variables and Semi-Regular Variables (hereon SRVs) are identified using $I$-band amplitudes and $P-W_I$, where $W_I$ is the reddening-free Wesenheit index \citep{2005AcA....55..331S}:
+
 \begin{eqnarray}
@@
 
-The resulting locus sample, as well as the original \cite{2014MNRAS.440.3430D} bounds and locus are showin in Fig~\ref{fig:locus}. Note that our sample goes beyond the limits of the stated locus.
+The resulting locus sample, as well as the original \cite{2014MNRAS.440.3430D} bounds and locus are shown in in Fig~\ref{fig:locus}. Note that our sample goes beyond the limits of the stated locus.
 
@@
 Note that we deliberately neglect $W4$ because, after applying the 5$\sigma$ faint limits as well as {\tt cc\_flag}=0, our sample is too heavily reduced.
+For similar reasons, we only apply the {\tt cc\_flag} criteria to [$W1/W2$], noting that clean flags in these two bands should be sufficient for source detection considering the extended NEOWISE mission contains extra observations for these wavelengh regimes.
 The resulting populations from each reduction step can be found in Table~\ref{tab:reductions}.
@@
 	\scalebox{0.85}{
-
 		\begin{tabular}{l r r r r r r r r r r}
@@
 		 $W1$-3 SNR $> 3$ & 18,995 & 5,279 & 2,876 & 34,285 & 9,673 & 1,079 & 213 & 3,056 & $\mathbf{75,456}$\\
+		 cc\_flag = 0 & 9,239 & 5,029 &  2,741 &  32,805 &  9,219 &  1,038 & 184 &  3,056 & $\mathbf{63,308}$\\
 		\hline\hline

### 2015-11-13 paper1/paper1_criteria.tex
Everything to the end of section 3

@@
 In producing our candidate sample, we must consider our likely physical sources of contamination, and how removing those sources reduces the completeness of our criteria.
-In Section~\ref{sec:comp_cont} we consider sample completeness and contamination, and use the sample sets from Section~\ref{sec:merged} to draw simple criteria in WISE-2MASS color-color space.
-YSOs present a special case of IR contaminants, as they are fairly well embedded in the same regions as AGB stars.
+Using the sample sets from Section~\ref{sec:merged}, we consider sample completeness and contamination in Section~\ref{sec:comp_cont} and devise simple criteria in WISE-2MASS color-color space.
+YSOs and RGBs present a special case of IR contaminants, as they are fairly well embedded in the same color-color regions as AGB stars.
 We consider the special case of YSOs in Section~\ref{sec:kill_ysos}, developing finer criteria to remove these objects while still preserving as much of the AGB sample as possible.
+We do the same with RGBs in Section~\ref{sec:kill_rgbs}.
 
@@
 where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$.
-Our goal is to get the lowest AGB sample ($\le$2\% contamination) while still retaining a large number of AGB stars.
+Our primary goal is to get an acceptably low contamination of the AGB sample ($\le$2\% contamination) while still retaining a large number of AGB stars.
 
@@
 &(J - K_s) &>\;\;1.1\label{eq:criteria1}\\
--0.1\;<&(W2 - W3) &<\;\; 2.5\label{eq:criteria2}
+0.2\;<&(W2 - W3) &<\;\; 2.5\label{eq:criteria2}
 \end{eqnarray}
-\noindent Figure~\ref{fig:colorcuts} shows completeness distributions for the OGLE-2MASS-WISE AGB sample, and corresponding contamination distributions for each contaminant sample.
+\noindent Figure~\ref{fig:colorcuts} shows completeness distributions for the OGLE-2MASS-WISE AGB sample, and corresponding contamination distributions for the contaminant samples.
+YSOs are omitted as they were not objects that could be projected to the LMC.
 Table~\ref{tab:criteria_completeness_contamination} shows the completeness and contamination fractions for each color criterion.
-We include the lower limit in ($W2-W3$) because, although there's a higher penalty on completeness than reduction in contamination, we want to exclude the spike of locus stars with ($W2-W3$) $< -0.1$.
+We include the lower limit in ($W2-W3$) because, although there's a higher penalty on completeness and a slight increase in contamination, we want to exclude the spike of locus stars with ($W2-W3$) $< 0.2$.
 
@@
 \caption{\emph{Left:} Cumulative completeness and contamination distributions before $(J-K_s)$ color cut.
-\emph{Right:} Cumulative completeness and contamination after $(J-K_s)$ cut and before $(W2-W3)$ cuts.
-Colors are the same as Figure~\ref{fig:wisecontaminants}.
+\emph{Right:} Cumulative completeness and contamination after $(J-K_s)$ and $(W2-W3)$ cuts.
 Thick black line in top panels is OGLE-2MASS-WISE AGB completeness as a function of color.
+The bottom panels show the contamination distributions with respect to color for stellar locus stars (\emph{orange, dashed}), RGB stars (\emph{red, dotted}), LRGs (\emph{purple, dashed}), QSOs (\emph{red, dashed}), AGN (\emph{green, dashed}), starforming galaxies (\emph{black, dashed}), and starburst galaxies (\emph{blue, dashed}).
 Bin width for both panels is 0.1 dex.
+Shaded regions represent the color-color space affected by equations~\ref{eq:criteria1} \& \ref{eq:criteria2}.
 \label{fig:colorcuts}}
@@
 		\hline\hline
-		Critiera & QSO & AGN & LRG & SF & SB & Locus & \textbf{AGB Completeness}\\
+		Critiera & QSO & AGN & LRG & SF & SB & Locus & RGB & \textbf{AGB Completeness}\\
 		\hline
-		$(J-K_s) > 1.1$ & 0.34\% & 0.18 \%& 0.09\% & 0.34\% & 0.60\% & 4.48\% &90.97\%\\
-		$(W2-W3) < 2.2$ & 0.03\% & 0\%& 0.04\% & 0.03\% & 0.02\% & 1.60\% &89.43\% \\
-		$(W2-W3) > -0.1$ & 0.03\% & 0\%& 0.04\% & 0.03\% & 0.02\% & 1.59\% &88.83\% \\
+		$(J-K_s) > 1.1$ & 0.25\% & 0.13 \%& 0.08\% & 0.56\% & 0.33\% & 4.86\% & 9.65\% &91.29\%\\
+		$(W2-W3) < 2.5$ & 0.03\% & $<0.01$\%& 0.04\% & 0.05\% & 0.01\% & 1.73\% & 7.41\% &90.03\% \\
+		$(W2-W3) > 0.2$ & 0.03\% & $<0.01$\%& 0.05\% & 0.06\% & 0.02\% & 1.83\% & 7.94\% &83.13\% \\
 
@@
 Figure~\ref{fig:lmc_completeness} shows the distribution of recovered AGB stars on the sky as well as in color-color space.
-We reserve an accounting of YSOs for the next section, as there is significant overlap between YSOs and AGBs, and they are not evenly distributed across the sky.
+We reserve a full accounting of YSOs and RGBs for the next subsections, as there is significant overlap between YSOs, RGBs, and AGBs, and they are not evenly distributed across the sky.
 
@@
 \includegraphics[width=6in]{figs/ogle_completeness_map.png}
-\caption{Map of completeness fractions for AGB stars in the LMC \label{fig:lmc_completeness}}
+\caption{Completeness fraction distributions for AGB stars in the LMC.
+\emph{Top:} OGLE-III AGB completeness fractions with position.
+Bin size is 0.2 deg on each axis.
+\emph{Bottom:} completeness  fractions in color-color space.
+Dashed lines represent the primary AGB candidate criteria in equations~\ref{eq:criteria1} \& \ref{eq:criteria2}.
+Bin size is 0.1 dex on each axis.
+\label{fig:lmc_completeness}}
 \end{figure}
@@
 \begin{figure}[h]
-\includegraphics[width=6in]{figs/taurus_field_YSOs.pdf}
-\caption{YSOs (\emph{red x}) in the Taurus Molecular Cloud \citep{2011ApJS..196....4R}.\label{fig:yso_field}}
+\includegraphics[width=4in]{figs/taurus_field_YSOs.pdf}
+\caption{YSOs (\emph{red x}) in the Taurus Molecular Cloud (\emph{background}), sampled from \citep{2011ApJS..196....4R}.\label{fig:yso_field}}
 \end{figure}
 \begin{figure}[h]
-\includegraphics[width=6in]{figs/ysos_taurus_color.pdf}
-\caption{YSOs (\emph{red x}) and AGB candidates (\emph{black o}) in the Taurus Molecular Cloud.\label{fig:ysos_agbs_taurus}}
-\end{figure}
-\begin{figure}[h]
-\includegraphics[width=6in]{figs/remove_ysos.pdf}
-\caption{YSOs (\emph{red x}) and AGB candidates (\emph{black o}) in the Taurus Molecular Cloud.\label{fig:ysos_lmc}}
+\includegraphics[width=6.5in]{figs/ysos_taurus_color.pdf}
+\caption{YSOs (\emph{red x}) and AGB candidates (\emph{black o}) in the Taurus Molecular Cloud. \label{fig:ysos_agbs_taurus}}
 \end{figure}
@@
 Figure~\ref{fig:yso_field} shows these 290 YSOs in the field of the Taurus Molecular Cloud, overplotted on WISE objects in that field.
-As shown in Table~\ref{tab:reductions}, after photometric and quality cuts we are left with 184 YSOs, and only 74 after applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2}.
-In contrast, applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2} as well as photometric criteria from Section~\ref{sec:merged} to the WISE field in the Taurus Molecular Cloud yields 6,995 AGB candidates.
+As shown in Table~\ref{tab:reductions}, after photometric and quality cuts we are left with 174 YSOs, and only 68 after applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2}.
+In contrast, applying equations~\ref{eq:criteria1} \& \ref{eq:criteria2} as well as photometric criteria from Section~\ref{sec:merged} to the WISE field in the Taurus Molecular Cloud yields 2,195 AGB candidates in that field.
 Figure~\ref{fig:ysos_agbs_taurus} shows the degree of entanglement in color-color space for YSOs and AGB candidates.
-We note, however, that because these are galactic AGB candidates, and the Milky Way is deficient in C-rich AGB stars, Figure~\ref{fig:ysos_agbs_taurus} misses the degree to which YSOs contaminate the C-rich sequence of AGB stars.
-To clarify this, we compare Taurus YSOs to AGB stars from the LMC (Figure~\ref{fig:ysos_lmc}).
+However, we note that because these are galactic AGB candidates, and the Milky Way is deficient in C-rich AGB stars [XXX CITATION HERE XXX], Figure~\ref{fig:ysos_agbs_taurus} misses the degree to which YSOs contaminate the C-rich sequence of AGB stars.
+To clarify, we compare Taurus YSOs to AGB stars from the LMC (Figure~\ref{fig:ysos_lmc}).
 Based on their positions in color-color space, we add three more constraints to remove YSOs from our AGB candidate sample:
@@
 \begin{align}
-(W1-W2) &< 0.75(W2-W3) - 0.33 & \text{for } (W2-W3) > 1.075\\
-(W1-W2) &> -(W2-W3) + 1.5 & \text{for }(W2-W3) > 1.075\\
-(W1-W2) &> 0.2 & \text{for }(W2-W3) > 1.3
+(W1-W2) &< 0.75(W2-W3) - 0.33 & \text{for } (W2-W3) > 1.075\label{eq:yso_criteria1}\\
+(W1-W2) &> -(W2-W3) + 1.5 & \text{for }(W2-W3) > 1.075\label{eq:yso_criteria2}\\
+(W1-W2) &> 0.2 & \text{for }(W2-W3) > 1.3\label{eq:yso_criteria3}
 \end{align}
 
-Collectively, these criteria preserve the C-rich AGB sequence, retain a relatively high completeness fraction for the LMC objects (82.72\%), and reduce the contamination in the Taurus Molecular Cloud due to YSOs from 1.91\% to 0.4\%.
+Collectively, these criteria preserve the C-rich AGB sequence, retain a relatively high completeness fraction for the LMC objects (82.34\%), and reduce the sample contamination in the Taurus Molecular Cloud due to YSOs from 1.91\% to 0.63\%.
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/remove_ysos.pdf}
+\caption{YSOs (\emph{red x}) and OGLE-III AGBs (\emph{black o}) in the LMC before (\emph{left}) and after (\emph{right}) the application of YSO exclusion criteria in equations \ref{eq:yso_criteria1}-\ref{eq:yso_criteria3}.
+Dashed green lines outline the bounds for the YSO exclusion criteria, whose area is shaded in black.\label{fig:ysos_lmc}}
+\end{figure}
+
+\subsection{Removing RGBs from WISE Color-Color Space}\label{sec:kill_rgbs}
+From [cite that Sosynski paper] we see that in the LMC RGB stars overlap with AGBs in color-color and period-luminosity space, and are very difficult to disentangle from AGB stars.
+However, they are differentiated by the color-corrected magnitudes of TRGB stars and confidence in a narrow-enough distance distribution that these magnitudes can serve a similar purpose as an absolute magnitude--reference magnitudes where the TRGB can be identified and cut.
+Since we don't have the luxury of known absolute magnitudes or distances for AGB candidates in the Milky Way, we instead use the OGLE-III RGB sample to estimate where they would be in color-color space.
+
+We begin by convolving the color-color criteria of equations~\ref{eq:criteria1} \& \ref{eq:criteria2} with the YSO exclusion critieria in \ref{eq:yso_criteria1}-\ref{eq:yso_criteria3}, yielding 603 LMC RGB stars.
+Even with this grand reduction in the population of RGB stars in our sample, we're still left with 7.97\% contamination after excluding YSO stars.
+This calls for a specific removal of RGB stars from the sample, so we sought to find where they are most concentrated and easiest removed without adversely affecting sample completeness.
+
+Figure~\ref{fig:rgbs_lmc} shows the distribution of the resulting RGB star sample with respect to the AGB star sample in the LMC in four colors. 
+We find that RGB stars overlap with AGB stars in a narrow sequence with the overwhelming majority falling under $J-K_s < 1.35$.
+In order to reduce the contamination from these objects in our candidate sample, we impose the following cut:
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/agbs_rgbs_color.pdf}
+\caption{OGLE-III RGBs (\emph{red x}) and OGLE-III AGBs (\emph{black dot}) in the LMC. 
+\label{fig:rgbs_lmc}}
+\end{figure}
+
+\begin{align}
+(J-K_s) > 1.35 \text{ for } (W2-W3) > 0.8.\label{eq:rgb_criteria}
+\end{align}
+
+\noindent This new criterion results in an LMC AGB star completeness of 71.14\%, while reducing the RGB star contamination down to 1.05\%.
+The resulting distribution is seen in Figure~\ref{fig:rgbs_after_cut}.
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/remove_rgbs.pdf}
+\caption{OGLE-III RGBs (\emph{red x}) and OGLE-III AGBs (\emph{black dot}) in the LMC before (\emph{left}) and after (\emph{right}) the application of equation~\ref{eq:rgb_criteria}. 
+\label{fig:rgbs_after_cut}}
+\end{figure}
+
+\subsection{Post-processing Filter}
+Using the criteria in equations~\ref{eq:criteria1}-\ref{eq:rgb_criteria}, we selected AGB candidates from the ALLWISE-2MASS database.
+We recovered 1,383,366 candidates, an almost full order of magnitude over the number estimated by [XXX CITE JACKSON, IVEZIC, AND KNAPP 2002XXX] for the entire Milky Way.
+Suspecting that we hadn't fully accounted for the contribution to the candidate sample from naked stars, we reconcile this discrepancy by adding one final cut motivated by Figure 4 [XXX CITE NIKUTTA ET AL 2014 XXX].
+In that figure, stars from SDSS with $W1 < 15.8$ are plotted in $W1$ vs $(W1-W2)$, showing a median $(W1-W2)$ of -0.04 dex.
+With that in mind, we remove objects with $(W1-W2) < -0.04$, retaining a final  population of 306,793 candidate AGB stars.
 

### 2015-11-13 paper1/paper1_data.tex
Everything to the end of section 3

@@
 relations that underlie our distance estimates. In order to assess what infrared populations
-could contribute to contamination of selected \agb\ candidates, we utilize
+could contribute to contamination of selected \agb\ candidates, we utilize stellar and 
 extragalactic sources from SDSS data release 7 pulled from the NYU Value-added Galaxy Catalog,
-the SDSS unofficial Luminous Red Galaxy sample,
-and young stellar objects identified in \wise\ data, also described below. 
+the SDSS unofficial Luminous Red Galaxy sample, young stellar objects identified in \wise\ data, and red giant branch stars from the OGLE-III Catalog of Variable Stars, also described below. 
 We conclude this section by describing how these auxiliary catalogs were positionally merged
@@
 
-\subsection{AGB stars from the OGLE-III Catalog of Variable Stars}\label{sec:ogle}
+\subsection{Variable Giants from the OGLE-III Catalog of Variable Stars}\label{sec:ogle}
 The OGLE-III Catalog of Variable Stars (CVS) \citep{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S} is a subset of the overall OGLE-III dataset, containing roughly 10 years of observations in the $V$- and $I$-bands of over 120,000 variable stars, with typically 700-900 data points per star in the $I$-band and about 50 data points per star in $V$-band.
@@
 
-The OGLE-III CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with long-period variables (LPVs) classified by object type, evolutionary status, and spectral type.
-Object type is based chiefly on variability and luminosity.
+The OGLE-III CVS is accessible through an online database\footnote{\url{http://ogledb.astrouw.edu.pl/~ogle/CVS/}}, with long-period variables (LPVs) classified by object type, evolutionary status , and spectral type.
+\subsubsection{AGB Stars}
+Stars brighter than the tip of the red giant branch (RGB) ($K_s < 12.05$ mag in LMC; \citealt{2004AcA....54..129S, 2007AcA....57..201S}) are classified as AGBs.
+Object types (OSARG, Mira, SRV) are based chiefly on variability and luminosity.
 OGLE Small Amplitude Red Giants (hereon OSARGs) are weakly-variable ($0.005 < A_I < 0.13$ mag), with relatively short periods \citep[$10 < P < 100$ days;][]{2004AcA....54..129S}.
@@
 
-As red giant branch stars (RGBs) can contaminate the same log($P)-L$ space as AGB stars, evolutionary type is distinguished in two ways.
-Stars above the tip of the RGB ($K_s = 12.05$ mag in LMC; \citealt{2004AcA....54..129S, 2007AcA....57..201S}) are classified as AGBs.
-Below that threshold, a narrow sequence of stars appears to share the same slope and intercept with AGB stars above the TRGB.
-Additionally they share similar primary-secondary period ratios as known AGB stars \citep{2005AcA....55..331S}.
-However, as there is still some contamination by RGBs using period ratios, we reject all OGLE objects below the TRGB.
-
 For individual log$(P)-L$ sequences, spectral type (O-rich or C-rich) is easily seen as separations in log$(P)-L$ space as well as visible-NIR color-color space.
@@
 
-From the initial OGLE-III sample of 52,976 LPVs, we retain 43,262 after matching to ALLWISE within 3", and ensuring only 1 match to 2MASS as well as all objects needing to be in the LMC.
-The $I$-band median of OGLE-III LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag.
-Out of the entire sample only 13 objects (Miras) fall beyond the 20.5 mag 5$\sigma$ faint limit, and 41 beyond the saturation limit.
-As such, we expect this to effectively represent the complete sample of oscillating AGB stars in the LMC.
+From the initial OGLE-III LMC sample of 46,467 LPVs we retain 44,013 after matching to ALLWISE within 3", and 43,247 after ensuring only 1 match to 2MASS.
+The $I$-band median of OGLE-III LPVs is $\approx14.68$ mag with a standard deviation of 0.72 mag, appearing well above the OGLE $I$-band 5$\sigma$ faint limit.
+Out of the entire sample only 13 objects (Miras) fall below the 5$\sigma$ faint limit, and 41 above the saturation limit.
+Because so few stars even approach the faint limit of OGLE, we expect the OGLE-2MASS catalog of AGBs to effectively represent the complete sample of AGB stars in the LMC down to $I < 20.5$, oscillating with amplitudes $> 0.005$ mag.
 
-Out of the 43,262 objects matched to ALLWISE within 3", 1,633 are Mira variables, 30,768 are OSARGs, and 10,861 are SRVs.
-The right side of Figure~\ref{fig:oglepmag} shows color-color plots for OGLE-2MASS colors, with a line at $(J-K_s) = 1.4$ drawn to separate O-rich and C-rich Miras.
+Out of the 43,247 objects matched to ALLWISE and 2MASS within 3", 1,633 are Mira variables, 30,756 are OSARGs, and 10,858 are SRVs.
+%The right side of Figure~\ref{fig:oglepmag} shows color-color plots for OGLE-2MASS colors, with a line at $(J-K_s) = 1.4$ drawn to separate O-rich and C-rich Miras.
 Out of the total of 457 O-rich Miras, 36 (7.9\%) have $(J-K_s) > 1.4$.
 Conversely, out of the 1,176 C-rich total Miras, 1,138 (96.8\%) have $(J-K_s) > 1.4$.
-The distribution of these objects on the sky is summarized in Figure~\ref{fig:oglegalplot}.
+The distributions of these objects in log$(P)-L$ and visible-NIR color-color space are summarized in Figure~\ref{fig:oglegalplot}.
+
+\subsubsection{Red Giant Branch Stars}
+As RGBs can contaminate the same log($P)-L$ space and color-color space as AGB stars, we seek to use RGBs identified by the OGLE-III CVS to quantify their contamination of our AGB sample.
+Stars brighter than the tip of the RGB ($K_s < 12.05$ mag in LMC; \citealt{2004AcA....54..129S, 2007AcA....57..201S}) are classified as AGBs.
+Below that threshold, narrow sequences of variable stars appear to share similar slopes and intercept with AGB stars above the TRGB.
+\cite{2004AcA....54..129S} classifies these objects as RGBs.
+The OGLE-III CVS LMC sample initially contains 45,528 RGB stars.
+When matched to ALLWISE-2MASS catalog within 3" we retain 40,764 RGBs.
 
@@
 \subsection{WISE+2MASS Stellar Locus}
-The color-color space occupied by AGB stars showing significant photospheric emission is heavily populated by naked stars from the main sequence.
+The color-color space occupied by AGB stars showing significant photospheric emission in WISE filters is heavily populated by naked stars from the main sequence \citep{2014MNRAS.442.3361N}.
 By sheer number, these stars would drown out sources that could be AGB stars with thin circumstellar envelopes.
 To address this, we extract the stellar locus from ALLWISE in the direction of the LMC ($276.5^\circ < l < 284^\circ$, $-38.2^\circ < b < -28^\circ$).
-We require that every object have 1 2MASS association, [$W1/W2/W3$] signal-to-noise $>$ 1, detections in 2MASS J, H, and K,  and confusion \& contamination flags set to 0 for [$W1/W2/W3$].
+We require that every object have one 2MASS association, [$W1/W2/W3$] signal-to-noise $>$ 1, detections in 2MASS $J$, $H$, and $K_s$,  and confusion \& contamination flags set to 0 for [$W1/W2/W3$].
 We then adapt the color-color criteria from \cite{2014MNRAS.440.3430D}.
@@
 (\emph{black dots}) objects from the stellar locus in the direction of the LMC.
-(\emph{blue line}) WISE-2MASS color-color stellar locus derived from \cite{2014MNRAS.440.3430D}.
+(\emph{green line}) WISE-2MASS color-color stellar locus derived from \cite{2014MNRAS.440.3430D}.
 (\emph{red line}) 3$\sigma$ boundaries for WISE-2MASS color-color locus.
@@
 
-The final OGLE-2MASS-WISE AGB sample is produced from the obejct sample in section~\ref{sec:ogle}.
+The final OGLE-2MASS-WISE AGB sample is produced from the object sample in section~\ref{sec:ogle}.
 We apply the point source saturation limits of [$W1/W2/W3/K_s$] $<$ [$2.0/1.5/-3.0/8.5$].
@@
 Note that we deliberately neglect $W4$ because, after applying the 5$\sigma$ faint limits as well as {\tt cc\_flag}=0, our sample is too heavily reduced.
-For similar reasons, we only apply the {\tt cc\_flag} criteria to [$W1/W2$], noting that clean flags in these two bands should be sufficient for source detection considering the extended NEOWISE mission contains extra observations for these wavelengh regimes.
+%For similar reasons, we only apply the {\tt cc\_flag} criteria to [$W1/W2$], noting that clean flags in these two bands should be sufficient for source detection considering the extended NEOWISE mission contains extra observations for these wavelengh regimes.
 The resulting populations from each reduction step can be found in Table~\ref{tab:reductions}.
@@
 		\hline\hline
-		Population & OGLE & QSO & AGN & SF & SB & LRGs & YSOs & Locus & \textbf{Total} \\ 
+		Population & OGLE & QSO & AGN & SF & SB & LRGs & YSOs & Locus & RGBs & \textbf{Total} \\ 
 		\hline
-		Original & 46,262 & 122,550 & 19,184 & 232,845 & 67,128 & 105,631 & 290 & 25,254 & $\mathbf{619,267}$ \\
-		WISE match & 43,209 & 6,902 & 3,098 & 36,539 & 10,359 & 75,543 & 274 & 25,254 & $\mathbf{201,178}$ \\
-		$W1$-3, $K_s$ saturation limit& 43,201 & 6,900 & 3,098 & 36,538 & 10,358 & 75,527 & 214 & 25,156 & $\mathbf{200,992}$ \\
-		 $W1$-2 faint limit & 43,200 & 6,891 & 3,098 & 36,537 & 10,358 & 75,527 & 214 & 24,382 & $\mathbf{200,207}$  \\
-		 $W1$-3 faint limit & 19,358 & 6,728 & 3,005 & 35,678 & 10,105 & 1,184 & 214 & 3,154 & $\mathbf{79,426}$  \\
-		 $W1$-3, $K_s$ faint limit & 19,358 & 5,283 & 2,878 & 34,299 & 9,676 & 1,167 & 213 & 3,070 & $\mathbf{75,944}$ \\
-		 $W1$-3 SNR $> 3$ & 18,995 & 5,279 & 2,876 & 34,285 & 9,673 & 1,079 & 213 & 3,056 & $\mathbf{75,456}$\\
-		 cc\_flag = 0 & 9,239 & 5,029 &  2,741 &  32,805 &  9,219 &  1,038 & 184 &  3,056 & $\mathbf{63,308}$\\
+		Original & 46,467 & 122,550 & 19,184 & 232,845 & 67,128 & 105,631 & 290 & 25,254 & 45,528 & $\mathbf{664,877}$ \\
+		WISE match & 43,247 & 6,902 & 3,098 & 36,539 & 10,359 & 75,543 & 274 & 25,254 & 40,764 & $\mathbf{241,980}$ \\
+		$W1$-3, $K_s$ sat. limit& 43,201 & 6,900 & 3,098 & 36,538 & 10,358 & 75,527 & 214 & 25,156 & 40,763 & $\mathbf{241,755}$ \\
+		 $W1$-2 faint limit & 43,200 & 6,891 & 3,098 & 36,537 & 10,358 & 75,527 & 214 & 24,382 & 40,760 & $\mathbf{240,967}$  \\
+		 $W1$-3 faint limit & 19,358 & 6,728 & 3,005 & 35,678 & 10,105 & 1,184 & 214 & 3,154 & 5,170 & $\mathbf{84,596}$  \\
+		 $W1$-3, $K_s$ faint limit & 19,358 & 5,283 & 2,878 & 34,299 & 9,676 & 1,167 & 213 & 3,070 & 5,170 & $\mathbf{81,114}$ \\
+		 $W1$-3 SNR $> 3$ & 18,995 & 5,279 & 2,876 & 34,285 & 9,673 & 1,079 & 213 & 3,056 & 4,727 & $\mathbf{80,183}$\\
+		 cc\_flag = 0 & 8,454 & 5,029 &  2,741 &  32,805 &  9,219 &  1,038 & 174 &  3,056 & 1,599 & $\mathbf{64,125}$\\
 		\hline\hline

### 2016-02-02 paper1/paper1_criteria.tex
Updated section 4 with figures and some starter text. Much more in the commented text

@@
 We recovered 1,383,366 candidates, an almost full order of magnitude over the number estimated by [XXX CITE JACKSON, IVEZIC, AND KNAPP 2002XXX] for the entire Milky Way.
-Suspecting that we hadn't fully accounted for the contribution to the candidate sample from naked stars, we reconcile this discrepancy by adding one final cut motivated by Figure 4 [XXX CITE NIKUTTA ET AL 2014 XXX].
+Suspecting that we hadn't fully accounted for the contribution to the candidate sample from naked stars, we reconcile this discrepancy by adding one final cut motivated by Figure 4 of [XXX CITE NIKUTTA ET AL 2014 XXX].
 In that figure, stars from SDSS with $W1 < 15.8$ are plotted in $W1$ vs $(W1-W2)$, showing a median $(W1-W2)$ of -0.04 dex.

### 2016-02-02 paper1/paper1_distribution.tex
Updated section 4 with figures and some starter text. Much more in the commented text

@@
 \label{sec:distribution}
-Put words here
+%Introduce the section.
+What's the overall gist?
+What are the high-level details that define this section?
+You show the full candidate distribution throughout the galaxy as well as in color-color space.
+You use OGLE AGB stars to define O- and C-rich AGB star color-magnitude trends as a basis for calculating the distances to AGB stars
+Having obtained our candidate sample, we validate by observing the spatial distribution of candidate AGB stars throughout the sky.
+We begin with the distribution of candidates in latitude and longitude, as well as their distributions in notable color-color relationships.
+Because their NIR brightness is dictated chiefly by the temperatures and compositions of their substantial dusty circumstellar shells, which themselves occupy a narrow range of potential temperatures, we can use their color information and narrow magnitude distribution to produce a color-magnitude relationship for AGB stars and estimate their distances.
+
+We set that stage using known chemically-classified AGB stars from OGLE in the magellanic clouds and find that narrow color-magnitude relationship.
+We then extend that relationship to our candidate sample, and continue on to produce a three-dimensional map of the galaxy in AGB stars.
+We discuss the methods involved in that production, the limitations of those methods, and the resulting galactic density distribution of our candidate AGB sample.
+
+\subsection{The Full Candidate Distribution}
+%Talk about the full distribution in galactic latitude and longitude space.
+Talk about how the distribution in color-color space is different depending on the region of the galaxy being explored.
+Where are the majority of stars located in physical space?
+Color-color space?
+Any trends worth noting outside of hard boundaries imposed by color cuts?
+This does not have to be a significantly long section, this is just a description of the data.
+We divide the candidate sample into 6 regions based on galactic position and investigate the resulting galactic and color-color distributions.
+These distributions can be seen in figures~\ref{fig:color_map_candidates0} through ~\ref{fig:color_map_candidates5}.
+The positional breakdown of the spatial regions are in Table~\ref{tab:region_table}:
+
+\begin{table}[h]
+	\begin{center}
+	\scalebox{0.85}{		
+		\begin{tabular}{c l r r}
+		\hline\hline
+		Region & Description & Count\\
+		\hline
+		1 & $\lvert\text{gal } b \rvert > 5$, $r_\text{GC} > 20$ & 22,280\\
+		2 & $\lvert\text{gal } b \rvert < 5$, $r_\text{GC} > 20$, $\lvert\text{gal } l \rvert < 90$ & 165,886\\
+		3 & $\lvert\text{gal } b \rvert < 5$, $r_\text{GC} > 20$, $\lvert\text{gal } l \rvert > 90$ & 52,715\\
+		4 & $r_\text{GC} < 20$, $r_\text{GC} > 10$ & 37,364\\
+		5 & $r_\text{GC} < 10$, $r_\text{GC} > 3$ & 18,856\\
+		6 & $r_\text{GC} < 3$ & 2,861\\
+		\hline\hline
+		\end{tabular}
+		}
+	\caption{$r_\text{GC}$ is the radius from the Galactic center in degrees. \label{tab:region_table}}
+	\end{center}
+\end{table}
+
+What we find is that, regardless of location, Galactic AGB candidates primarily populate the region of W1-W2 vs W2-W3 occupied by O-rich OGLE AGB stars with W1-W2 $<$ 0.4.
+Additionally the inner 5 kpc contains most of the candidate AGB stars, with 240,881 candidates outside of 20$^\circ$ from the Galactic  center.
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region0.pdf}
+\caption{\label{fig:color_map_candidates0}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region1.pdf}
+\caption{\label{fig:color_map_candidates1}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region2.pdf}
+\caption{\label{fig:color_map_candidates2}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region3.pdf}
+\caption{\label{fig:color_map_candidates3}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region4.pdf}
+\caption{\label{fig:color_map_candidates4}}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=6.5in]{figs/color_and_map_candidates_region5.pdf}
+\caption{\label{fig:color_map_candidates5}}
+\end{figure}
+
+\subsection{O-rich and C-rich AGB Candidates}
+%This section should be somewhat substantial, as you did a lot to get from the previous one through to the end of this one.
+You used the known AGB stars from OGLE matched to WISE in addition to some help from Nikutta et al 2014 to define O-rich and C-rich boundaries for AGB stars in color-color space.
+Talk about how these boundaries reduced your sample.
+Also talk about how these selections of stars were used to estimate  color-magnitude relationships for AGB stars with the LMC as reference, and find some way to justify the 2 magnitude offset that makes the results approach believability.
+Then talk about how distances were calculated.
+It was an iterative process that used a test distance to calculate the expected dust column along the line of sight until the sum of distance modulus and extinction coefficient matched apparent minus absolute magnitude.
+What were the steps in distance modulus?
+How did you get the extinction coefficient from the color excess that's actually output from galfastdust?
+What's the source for the galfastdust estimation procedure?
+And what are the results for the distance measurement?
+
+\subsection{Galactic Number Density}
+%This is the section where we show that what we've done is not complete bullshit, though some fecal flecks may remain. The big fig is going to be that plot of |Z| vs. R
